### 2020-10-19 manuscript.Rmd
Initialize manuscript styling

---
documentclass: jss
classoption: nojss
citation_package: natbib
bibliography: references.bib
author:
  - name: Devin Incerti
    affiliation: Genentech
    # use this syntax to add text on several lines
    address: |
      | South San Francisco
    email: \email{incerti.devin@gene.com}
  - name: Jeroen Jansen
    affiliation: |
      | University of California, San Francisco
    address: |
      | San Francisco
    email: \email{jeroen.jansen@ucsf.edu}
title:
  plain:     "hesim: Health Economic Simulation Modeling and Decision Analysis"
  formatted: "\\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis"
abstract: >
  The abstract of the article.
keywords:
  formatted: [health economic evaluation, cost-effectiveness analysis, simulation, modeling, "\\proglang{R}"]
  plain: [health economic evaluation, cost-effectiveness analysis, simulation, modeling, R]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
knitr::opts_chunk$set(eval = TRUE)
```

# Introduction

### 2020-10-21 manuscript.Rmd
First draft of introduction

@@
 abstract: >
-  The abstract of the article.
+  To do.
 keywords:
@@
 # Introduction
+Health technology assessment (HTA) is a systematic approach for comparing competing health technologies to inform the efficient use of health care resources.
+Publicly funded health systems such as those in Australia, Canada, and the United Kingdom, among others, use HTA to help maximize health gains for the population given a fixed budget.
+HTA is also commonly used by private payers to guide coverage decisions and the adoption of new technologies \citep{trosman2011health}.
+Such assessments usually rely heavily on cost-effectiveness analysis (CEA) so that decisions can be made on the basis of a formal evaluation of costs and effects \citep{dakin2015influence}.
+CEA is made feasible by development of health economic models that simulate costs and effects over relevant time horizons using the totality of available evidence.
+
+The most commonly used economic models are Markov state transition models (STMs) that simulate transitions between mutually exclusive health states.
+When the Markov assumption holds so that transition probabilities are either constant over time or depend only on model time, a cohort-level model can be used \citep{briggs1998introduction}.
+If the Markov assumption is relaxed---e.g., with a semi-Markov model so that transition probabilities depend on time in an intermediate health state---then individual-level models are required in most cases \citep{brennan2006taxonomy, fiocco2008reduced}.
+Individual-level models afford considerably more flexibility and allow patient history to be tracked over time.
+
+Decisions informed by economic models and CEA are subject to uncertainty.
+One source of decision uncertainty stems from uncertainty in the underlying model parameters.
+Parameter uncertainty is typically quantified using probabilistic sensitivity analysis (PSA), which involves randomly sampling the model parameters from suitable probability distribution and simulating the model for each sampled parameter set \citep{claxton2005probabilistic}.
+When combined with individual-level simulation, PSA can take an appreciable amount of time to run \citep{ohagan2007monte}.
+
+Estimation of model parameters should ideally be performed using statistical models that are aligned with the structure of the economic model.
+For instance, when patient-level data is available, a multi-state model can be used to parameterize all possible transitions in a STM while accounting for censoring and competing risks \citep{williams2017cost}.
+Similarly, in oncology, partitioned survival models (PSMs) can be parameterized from estimates of progression-free survival (PFS) and overall survival (OS).
+In other cases, parameters might be combined from disparate sources, such as within a single Bayesian model \citep{baio2012bayesian}.
+When the clinical evidence base is not limited to a single study, a formal evidence synthesis, such as a network-meta analysis (NMA), might even be performed \citep{dias2018network}.
+
+Despite their computational demands and foundations in statistics, health economic models have historically been developed with specialized commercial software (e.g., \proglang{TreeAge}) or more commonly with a spreadsheet (almost always Microsoft \proglang{Excel}).
+The limitations of such software relative to programming languages like \proglang{R} have been increasingly emphasized in the literature \citep{baio2017simple, incerti2019r, jalal2017overview}.
+It is therefore no surprise that a number of related \proglang{R} packages have recently been developed, such as \pkg{BCEA} \citep{baio2017bayesian}, \pkg{SAVI} \citep{strong2014estimating}, \pkg{survHE} \citep{baio2020survhe}, and \pkg{heemod} \citep{filipovic2017markov}.
+Still, of the available packages, only \pkg{heemod} provides a general purpose framework for developing simulation models and it is limited to cohort Markov models.
+
+`hesim` is an \proglang{R} package that advances the functionality and performance of the existing software.
+Multiple model types are supported including cohort discrete time state transition models (DTSTMs), N-state PSMs, and individual-level continuous time state transition models (CTSTMs), encompassing both Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
+To maximize flexibility and faciliate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R} or by inputting parameters obtained from external sources.
+So that individual-level simulation and PSA can be run quickly, \pkg{Rcpp} and \pkg{data.table} are heavily utilized.
+After simulating costs and quality-adjusted life-years (QALYs) from a PSA, decision analysis can be performed within a cost-effectiveness framework.

### 2020-10-23 manuscript.Rnw
Switch from R Markdown to Sweave

--- 2020-10-21 manuscript.Rmd
@@
----
-documentclass: jss
-classoption: nojss
-citation_package: natbib
-bibliography: references.bib
-author:
-  - name: Devin Incerti
-    affiliation: Genentech
-    # use this syntax to add text on several lines
-    address: |
-      | South San Francisco
-    email: \email{incerti.devin@gene.com}
-  - name: Jeroen Jansen
-    affiliation: |
-      | University of California, San Francisco
-    address: |
-      | San Francisco
-    email: \email{jeroen.jansen@ucsf.edu}
-title:
-  plain:     "hesim: Health Economic Simulation Modeling and Decision Analysis"
-  formatted: "\\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis"
-abstract: >
+\documentclass[article, nojss]{jss}
+
+%% -- LaTeX packages and custom commands ---------------------------------------
+
+%% recommended packages
+\usepackage{thumbpdf,lmodern}
+
+%% new custom commands
+\newcommand{\class}[1]{`\code{#1}'}
+\newcommand{\fct}[1]{\code{#1()}}
+
+%% For Sweave-based articles about R packages:
+%% need no \usepackage{Sweave}
+\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
+<<preliminaries, echo=FALSE, results=hide>>=
+options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+@
+
+%% -- Article metainformation (author, title, ...) -----------------------------
+
+%% - \author{} with primary affiliation
+%% - \Plainauthor{} without affiliations
+%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
+%% - \AND starts a new line, \And does not.
+\author{Devin Incerti\\Genentech
+   \And Jeroen Jansen\\University of California, San Francisco}
+\Plainauthor{Devin Incerti, Jeroen Jansen}
+
+%% - \title{} in title case
+%% - \Plaintitle{} without LaTeX markup (if any)
+%% - \Shorttitle{} with LaTeX markup (if any), used as running title
+\title{\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis}
+\Plaintitle{hesim: Health Economic Simulation Modeling and Decision Analysis}
+%\Shorttitle{A Short Demo Article in \proglang{R}}
+
+%% - \Abstract{} almost as usual
+\Abstract{
   To do.
-keywords:
-  formatted: [health economic evaluation, cost-effectiveness analysis, simulation, modeling, "\\proglang{R}"]
-  plain: [health economic evaluation, cost-effectiveness analysis, simulation, modeling, R]
-preamble: >
-  \usepackage{amsmath}
-output: rticles::jss_article
----
+}
 
-```{r, setup, include=FALSE}
-options(prompt = 'R> ', continue = '+ ')
-knitr::opts_chunk$set(eval = TRUE)
-```
+%% - \Keywords{} with LaTeX markup, at least one required
+%% - \Plainkeywords{} without LaTeX markup (if necessary)
+%% - Should be comma-separated and in sentence case.
+\Keywords{health economic evaluation, cost-effectiveness analysis, simulation, modeling, \proglang{R}}
+\Plainkeywords{health economic evaluation, cost-effectiveness analysis, simulation, modeling, R}
 
-# Introduction
+%% - \Address{} of at least one author
+%% - May contain multiple affiliations for each author
+%%   (in extra lines, separated by \emph{and}\\).
+%% - May contain multiple authors for the same affiliation
+%%   (in the same first line, separated by comma).
+\Address{
+  Devin Incerti\\
+  Genentech\\
+  South San Francisco\\
+  E-mail: \email{incerti.devin@gene.com}\\
+
+  Jeroen Jansen\\
+  University of California, San Francisco\\
+  San Francisco\\
+  E-mail: \email{jeroen.jansen@ucsf.edu}\\
+}
+
+\begin{document}
+
+%% -- Introduction -------------------------------------------------------------
+
+%% - In principle "as usual".
+%% - But should typically have some discussion of both _software_ and _methods_.
+%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
+%% - If such markup is in (sub)section titles, a plain text version has to be
+%%   added as well.
+%% - All software mentioned should be properly \cite-d.
+%% - All abbreviations should be introduced.
+%% - Unless the expansions of abbreviations are proper names (like "Journal
+%%   of Statistical Software" above) they should be in sentence case (like
+%%   "generalized linear models" below).
+
+\section{Introduction} \label{sec:intro}
 Health technology assessment (HTA) is a systematic approach for comparing competing health technologies to inform the efficient use of health care resources.
@@
 
-`hesim` is an \proglang{R} package that advances the functionality and performance of the existing software.
+\pkg{hesim} is an \proglang{R} package that advances the functionality and performance of the existing software.
 Multiple model types are supported including cohort discrete time state transition models (DTSTMs), N-state PSMs, and individual-level continuous time state transition models (CTSTMs), encompassing both Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
@@
 After simulating costs and quality-adjusted life-years (QALYs) from a PSA, decision analysis can be performed within a cost-effectiveness framework.
+
+\section{Framework} \label{sec:framework}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width = 15cm,  keepaspectratio]{econ-eval-process-hesim.png}
+\caption{Economic modeling process}\label{fig:econ-eval-process-hesim}
+\end{figure}
+
+%% -- Bibliography -------------------------------------------------------------
+%% - References need to be provided in a .bib BibTeX database.
+%% - All references should be made with \cite, \citet, \citep, \citealp etc.
+%%   (and never hard-coded). See the FAQ for details.
+%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
+%% - Titles in the .bib should be in title case.
+%% - DOIs should be included where available.
+\bibliography{references}
+
+%% -- Appendix (if any) --------------------------------------------------------
+%% - After the bibliography with page break.
+%% - With proper section titles and _not_ just "Appendix".
+
+%% -----------------------------------------------------------------------------
+
+\end{document}

### 2020-10-23 manuscript.Rnw
First draft of abstract

@@
 \Abstract{
-  To do.
+Health economic models simulate the costs and effects of health technologies for use in health technology assessment (HTA) to inform efficient use of scarce resources.
+Models have historically been developed using spreadsheet software (e.g., Microsoft \proglang{Excel}) and while use of \proglang{R} is growing, general purpose modeling software is still limited.
+\pkg{hesim} helps fill this gap by facilitating parameterization, simulation, and analysis of economic models in an integrated manner.
+Supported model types include cohort discrete time state transition models (cDTSTMs), individual continuous time state transition models (iCTSTMs), and partitioned survival models (PSMs), encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
+A modular design based on \pkg{R6} and \proglang{S3} classes allows users to combine separate submodels for disease progression, costs, and quality-adjusted life years (QALYs) in a flexible way.
+Probabilistic sensitivity analysis (PSA) is used to propagate uncertainty in model parameters to model outputs.
+Simulation code is written in \proglang{C++} so complex simulations such as those combining PSA and individual simulation can be run much more quickly than previously possible.
+Decision analysis within a cost-effectiveness framework is performed using simulated costs and QALYs from a PSA.
 }
@@
 \section{Framework} \label{sec:framework}
+As shown in Figure~\ref{fig:econ-eval-process-hesim}, a typical analysis proceeds in a 3 steps:
 
@@
 \centering
-\includegraphics[width = 15cm,  keepaspectratio]{econ-eval-process-hesim.png}
+\includegraphics[width = 15cm]{econ-eval-process-hesim.png}
 \caption{Economic modeling process}\label{fig:econ-eval-process-hesim}

### 2020-10-25 manuscript.Rnw
First draft of cDTSTM mathematical description

@@
 
+%% -- Model taxonomy -----------------------------------------------------------
+\section{Model taxonomy} \label{sec:model-taxonomy}
+STMs simulate transitions between mutually exclusive health states.
+A common assumption is that the STM is a Markov model, meaning that transitions to the next health state can only depend on the present health state.
+In a time homogeneous Markov model transition probabilities are constant over time, whereas in a time inhomogeneous model they can depend on time since the start of the model.
+A semi-Markov model relaxes the Markov assumption and allows transitions to depend on time since entering an intermediate state.
+
+Markov and semi-Markov models can be formulated in either continuous or discrete time, at either the cohort or individual level.
+In health economics, Markov cohort models tend to be formulated in discrete time (i.e., as cDTSTMs), although state probabilities can be computed in continuous time models using the Aalen-Johansen estimator \citep{aalen1978empirical}.
+While tunnel states can be used to approximate a semi-Markov model using a cohort approach, they can only be simulated in a general fashion using individual level models.
+Discrete time individual simulation is possible, but we use a continuous time models (i.e., iCTSTMs) because they do not require specification of model cycles and can be run considerably faster.
+
+PSMs are specialized models that can be parameterized using survival curves and are especially useful in oncology where PFS and OS are commonly reported.
+They are ``area under the curve'' models, although they can also be formulate as STMs by using the survival curves to construct transition probabilities.
+
+\subsection{Cohort Discrete time state transition models} \label{sec:cDTSTMs}
+cDTSTMs simulate the probability that a cohort of patients is in each of $H$ health states over time.
+Time is measured at discrete times with each time point known as a model cycle.
+A $1 \times H$ state vector that stores the probability of being in each health state at time $t$ is written as $v_t = [v_{1t}, v_{2t}, \ldots, v_{Ht}]$ where $\sum_i v_{ht} = 1$.
+A transition probability matrix is denoted by $p_t$ where the $(r,s)$th element represents a transition from state $r$ to state $s$ between times $t$ and $t+1$.
+The state vector at time $t+1$ for each of $T$ model cycles can be computed as,
+
+\begin{equation} \label{eqn:markov-sim}
+v_{t+1}= v_t p_t, \quad t = 0,\ldots, T.
+\end{equation}
+
+Costs and QALYs are computed by assigning (potentially time-varying) values to each health state.
+Utility, a measure of preference for a health state that normally ranges from $0$ (dead) to $1$ (perfect health), is used when computing QALYs \citep{torrance1986measurement}.
+Assuming model time is in years, state values for costs are estimated by annualizing costs.
+In a Markov model, state values, like transition probabilities, can depend on time since the start of the model but not on time since entering an intermediate health state.
+
+Expected values are computed by integrating the "weighted" probability of being in each state, where weights are a function of the discount factor and state values.
+That is, for a time horizon $T$, discounted costs and QALYs in health state h are computed as,
+
+\begin{equation} \label{eqn:cohort-costs-qalys}
+\int_0^{T} z_h(t) e^{-rt} P_h(t) dt,
+\end{equation}
+
+where $z_h(t)$ is the predicted cost or utility value at time $t$, $r$ is the discount rate, and $P_h(t)$ is the probability of being in a given health state.
+
+In a discrete time, the integral is approximated with
+
+\begin{equation} \label{eqn:riemann-sum}
+\sum_{j = 1}^{T}f(t_j^{*})\Delta t_j
+\end{equation}
+
+where $\Delta t_j = t_j - t_{j-1}$, $t_j^{*}\in[t_{j-1},t_j]$, and $f(t_j^{*})= z_h(t_j^{*}) e^{-rt_j^{*}} P_h(t_j^{*})$.
+Three methods can be used to estimate $f(t_j^{*})$.
+First, a left Riemann sum uses values at the start of each time interval $f(t_j^{*}) = f(t_{j-1})$.
+Second, a right Riemann sum uses values at the end of each time interval $f(t_j^{*}) = f(t_j)$.
+Finally, the trapezoid rule averages values at the start and end of each interval $f(t_j^{*}) = \frac{1}{2}\Delta t_j[f(t_{j-1}) + f(t_j)]$.
+
+\subsection{Individual Continuous time state transition models} \label{sec:cDTSTMs}
+
+\subsection{Partitioned survival models} \label{sec:PSMs}
+
+%% -- Framework ----------------------------------------------------------------
 \section{Framework} \label{sec:framework}

### 2020-10-25 manuscript.Rnw
Disease progression theory for iCTSTM

@@
 \usepackage{thumbpdf,lmodern}
+
+%% other packages
+\usepackage{algorithm}
 
@@
 
-\subsection{Cohort Discrete time state transition models} \label{sec:cDTSTMs}
+\subsection{Cohort discrete time state transition models} \label{sec:cDTSTMs}
 cDTSTMs simulate the probability that a cohort of patients is in each of $H$ health states over time.
 Time is measured at discrete times with each time point known as a model cycle.
-A $1 \times H$ state vector that stores the probability of being in each health state at time $t$ is written as $v_t = [v_{1t}, v_{2t}, \ldots, v_{Ht}]$ where $\sum_i v_{ht} = 1$.
-A transition probability matrix is denoted by $p_t$ where the $(r,s)$th element represents a transition from state $r$ to state $s$ between times $t$ and $t+1$.
+A $1 \times H$ state vector that stores the probability of being in each health state at time $t$ is written as $v_t = (v_{1t}, v_{2t}, \ldots, v_{Ht})$ where $\sum_i v_{ht} = 1$.
+A transition probability matrix is denoted by $P_t$ where the $(r,s)$th element represents a transition from state $r$ to state $s$ between times $t$ and $t+1$.
 The state vector at time $t+1$ for each of $T$ model cycles can be computed as,
@@
 \begin{equation} \label{eqn:markov-sim}
-v_{t+1}= v_t p_t, \quad t = 0,\ldots, T.
+v_{t+1}= v_t P_t, \quad t = 0,\ldots, T.
 \end{equation}
@@
 
-\subsection{Individual Continuous time state transition models} \label{sec:cDTSTMs}
+\subsection{Individual continuous time state transition models} \label{sec:iCTSTMs}
+iCTSTMs simulate individual trajectories between health states in continuous time using random number generation.
+Trajectories are simulated for multiple patients and costs and QALYs are computed by averaging across the simulated patients.
+A reasonably large number of patients must be simulated to ensure that expected values are stable \citep{ohagan2007monte}.
+
+Simulation is performed using a parametric and flexible parametric multi-state modeling framework based on the \pkg{flexsurv} package \citep{jackson2016flexsurv}.
+Let $t_j$ denote the time a patient enters state $r$ and $t_{j+1}$ denote the time a patient starting in state $r$ transitions to state $s$.
+Time-to-event $t^*$ for the $r \rightarrow s$ has probability density function,
+
+\begin{equation}
+f_{rs}(t^*|\theta(z), t_j), \quad t^*\geq 0,
+\end{equation}
+
+where parameters $\theta = (\theta_1, \theta_2, \ldots, \theta_p)$ may depend on covariates $z_p$ through the link function $g(\theta_p) = z_p^T \gamma$ where $\gamma$ are a vector of regression coefficients.
+In a time inhomogeneous Markov model, time $t^*=t_{j+1}$ is conditional on not experiencing event $s$ until time $t_j$ (i.e., it is left-truncated at time $t_j$).
+In a semi-Markov model, time-to-event is expressed in terms of $t^* = t_{j+1}-t_j$ and a patient enters state $s$ at time $t_j + t^*$.
+
+A survival distribution is specified for each permitted transition in the multi-state model.
+A trajectory through the model can then be simulated as described in Algorithm~\ref{alg:sim-ictstm}.
+While the algorithm repeats until a patient dies, it can also be stopped at a specified time $t$ or when a patient reaches a maximum age.
+In the latter scenario, death is assume to occur at the maximum age.
+
+\begin{algorithm}
+\caption{Simulation of individual continuous time state transition model}
+\label{alg:sim-ictstm}
+\begin{enumerate}
+\item Let $r$ be the starting state and starting time be $t_j = 0$. The number of permitted transitions from state $r$ is given by $n_r$.
+\item Simulate times $\mathcal{T} = \{t_{1, j+1}, t_{2, j+1}, \ldots t_{n_r, j+1}\}$ to each of the $n_r$ permitted transitions.
+\item Set the time of the transition $t_{j+1}$ equal to the minimum simulated time in $\mathcal{T}$ and the next state $s$ to the state with the minimum simulated time.
+\item Set $r=s$ and $t_j = t_{j+1}$. If the patient is still alive, repeat the previous steps until death.
+\end{enumerate}
+\end{algorithm}
 

### 2020-10-25 manuscript.Rnw
Add cost + QALYs to iCTSTM theory section

@@
 
+Costs and QALYs are computed using the continuous time present value given a flow of state values, which change as patients transition between health states or as costs vary as a function of time.
+The state values can be partitioned into $M$ time intervals indexed by $m = 1,\ldots, M$ where interval $m$ contains times $t$ such that $t_m\leq t \leq t_{m+1}$ and values for state $h$ are equal to $z_{hm}$ during interval $m$.
+$z_{hm}$ will equal zero during time intervals in which a patient is not in state $h$.
+Discounted costs and QALYs for health state $h$ are then given by,
+
+\begin{equation}
+\sum_{m = 1}^M \int_{t_m}^{t_m+1} z_{hm}e^{-rt}dt = \sum_{m = 1}^M z_{hm} \left(\frac{e^{-r{t_{m}}} - e^{-r{t_{m+1}}}}{r}\right),
+\end{equation}
+
+where $r > 0$ is again the discount rate. If $r = 0$, then the present value simplifies to $\sum_{m = 1}^M z_{hm}(t_{m+1} - t_{m})$.
+
+Note that while state values in cohort models can depend on time since the start of the model, state values in individual-level models can depend on either time since the start of the model or time since entering the most recent health state.
+Individual-level models consequently not only afford more flexibility than cohort models when simulating disease progression, but when simulating costs and/or QALYs as well.
+
 \subsection{Partitioned survival models} \label{sec:PSMs}

### 2020-10-27 manuscript.Rnw
Some updates to iCTSTM theory section

@@
 Time is measured at discrete times with each time point known as a model cycle.
-A $1 \times H$ state vector that stores the probability of being in each health state at time $t$ is written as $v_t = (v_{1t}, v_{2t}, \ldots, v_{Ht})$ where $\sum_i v_{ht} = 1$.
+A $1 \times H$ state vector that stores the probability of being in each health state at time $t$ is written as $x_t = (x_{1t}, x_{2t}, \ldots, x_{Ht})$ where $\sum_i x_{ht} = 1$.
 A transition probability matrix is denoted by $P_t$ where the $(r,s)$th element represents a transition from state $r$ to state $s$ between times $t$ and $t+1$.
-The state vector at time $t+1$ for each of $T$ model cycles can be computed as,
+The state vector at time $t+1$ for each of $T$ model cycles is given by,
 
 \begin{equation} \label{eqn:markov-sim}
-v_{t+1}= v_t P_t, \quad t = 0,\ldots, T.
+x_{t+1}^T= x_t^T P_t, \quad t = 0,\ldots, T.
 \end{equation}
@@
 \subsection{Individual continuous time state transition models} \label{sec:iCTSTMs}
-iCTSTMs simulate individual trajectories between health states in continuous time using random number generation.
+iCTSTMs simulate individual trajectories between health states using random number generation.
 Trajectories are simulated for multiple patients and costs and QALYs are computed by averaging across the simulated patients.
@@
 
-Simulation is performed using a parametric and flexible parametric multi-state modeling framework based on the \pkg{flexsurv} package \citep{jackson2016flexsurv}.
-Let $t_j$ denote the time a patient enters state $r$ and $t_{j+1}$ denote the time a patient starting in state $r$ transitions to state $s$.
-Time-to-event $t^*$ for the $r \rightarrow s$ has probability density function,
+In continuous time, a patient is in state $X(t)$ at time $t$.
+State transitions are modeled using a multi-state modeling framework \citep{putter2007tutorial} where the probability of a transition from state $r$ to state $s$ is governed by the hazard function,
+
+\begin{equation}
+h_{rs}(t) = \lim_{\Delta t\to 0} \frac{\Prob(X(t + \Delta t) = s | X(t) = r)}{\Delta t}.
+\end{equation}
+
+Simulated disease progression is characterized by $J$ distinct jumps between health states $D = \{(t_0, X(t_0)), (t_1, X(t_1)), \ldots (t_J, X(t_J))\}$ with a patient remaining in a health state from time $t_j$ until transitioning to the next state at $t_{j+1}$.
+Jumps between health states are simulated using parametric and flexible parametric survival models as implemented in the \pkg{flexsurv} package \citep{jackson2016flexsurv}.
+Specifically, if a patient enters state $r$ at time $t_j$, then a probability density function for the time-to-event $t^*$ for the $r \rightarrow s$ transition is,
 
@@
 
-where parameters $\theta = (\theta_1, \theta_2, \ldots, \theta_p)$ may depend on covariates $z_p$ through the link function $g(\theta_p) = z_p^T \gamma$ where $\gamma$ are a vector of regression coefficients.
+where parameters $\theta = (\theta_1, \theta_2, \ldots, \theta_p)$ may depend on covariates $z_p$ through the link function $g(\theta_p) = z_p^T \gamma$ and $\gamma$ is a vector of regression coefficients.
 In a time inhomogeneous Markov model, time $t^*=t_{j+1}$ is conditional on not experiencing event $s$ until time $t_j$ (i.e., it is left-truncated at time $t_j$).
@@
 \begin{enumerate}
-\item Let $r$ be the starting state and starting time be $t_j = 0$. The number of permitted transitions from state $r$ is given by $n_r$.
+\item Let $r$ be the state entered at time $t_j$. The number of permitted transitions from state $r$ is given by $n_r$. If $j=0$, then $t_j = 0$. 
 \item Simulate times $\mathcal{T} = \{t_{1, j+1}, t_{2, j+1}, \ldots t_{n_r, j+1}\}$ to each of the $n_r$ permitted transitions.

### 2020-10-28 manuscript.Rnw
First draft of partitioned survival modeling theory section

@@
 \subsection{Partitioned survival models} \label{sec:PSMs}
+PSMs are conceptually similar to state transition models in that they are characterized by mutually exclusive health states.
+They differ, however, in that state probabilities are not computed via matrix multiplication or individual simulation, but from a set of non-mutually exclusive survival curves \citep{glasziou1990quality, woods2018nice}.
+Each survival curve represents time to transitioning to that state or to a more severe health state.
+
+In $N$-state model, $N-1$ non-mutually exclusive survival curves are required.
+The cumulative survival function, $S_n(t)$, represents the probability that a patient survives to health state $n$ or to a lower indexed state beyond time $t$.
+The probability that a patient is in health state 1 is $S_1(t)$.
+State membership in health states $2,\ldots, N-1$ is computed as $S_{n}(t) - S_{n-1}(t)$.
+Finally, the probability of being in the final health state $n$ (i.e., the death state) is $1-S_{N-1}(t)$, or one minus overall survival function.
+
+Survival functions are estimated by fitting parametric or flexible parametric survival models as described in Section~\ref{sec:iCTSTMs}.
+The $n$th fitted survival model with density $f_n(t)$ has cumulative density function $F_n(t)$, survivor function $1 - F_n(T)$, cumulative hazard $H_n(t) = -\log S_n(t)$, and hazard $h_n(t) = f_n(t)/S_n(t)$.
+State probabilities for each health states can be computed for an arbitrarily fine grid of time points to produce the health state vector $x_t$ for each time $t$ in the grid.
+Costs and QALYs are then computed in the same manner as in the cohort model described in Section~\ref{sec:cDTSTMs}.
 

### 2020-10-29 manuscript.Rnw
A start on the coding framework section

@@
 \usepackage{algorithm}
+\usepackage{multirow}
+\usepackage{caption} 
+\captionsetup[table]{skip=5pt}
 
@@
 Supported model types include cohort discrete time state transition models (cDTSTMs), individual continuous time state transition models (iCTSTMs), and partitioned survival models (PSMs), encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
-A modular design based on \pkg{R6} and \proglang{S3} classes allows users to combine separate submodels for disease progression, costs, and quality-adjusted life years (QALYs) in a flexible way.
+A modular design based on \pkg{R6} and \code{S3} classes allows users to combine separate submodels for disease progression, costs, and quality-adjusted life years (QALYs) in a flexible way.
 Probabilistic sensitivity analysis (PSA) is used to propagate uncertainty in model parameters to model outputs.
@@
 %% - Should be comma-separated and in sentence case.
-\Keywords{health economic evaluation, cost-effectiveness analysis, simulation, modeling, \proglang{R}}
-\Plainkeywords{health economic evaluation, cost-effectiveness analysis, simulation, modeling, R}
+\Keywords{health economic evaluation, cost-effectiveness analysis, simulation, multi-state models, \proglang{R}}
+\Plainkeywords{health economic evaluation, cost-effectiveness analysis, simulation, multi-state models, R}
 
@@
 \subsection{Partitioned survival models} \label{sec:PSMs}
-PSMs are conceptually similar to state transition models in that they are characterized by mutually exclusive health states.
+PSMs are conceptually similar to STMs in that they are characterized by mutually exclusive health states.
 They differ, however, in that state probabilities are not computed via matrix multiplication or individual simulation, but from a set of non-mutually exclusive survival curves \citep{glasziou1990quality, woods2018nice}.
@@
 \section{Framework} \label{sec:framework}
-As shown in Figure~\ref{fig:econ-eval-process-hesim}, a typical analysis proceeds in a 3 steps:
+The \pkg{hesim} coding framework aims to follow three main tenets.
+First, it should be possible to seamlessly integrate parameter estimation with simulation.
+Second, modular code show allow for models to be built in a flexible manner.
+Finally, code should be computationally efficient.
+
+Each economic model consists of separate statistical models of disease progression, costs, and utilities.
+All analyses proceed in 4 steps.
+The first step is to setup the model by defining the model structure, target population, and treatment strategies of interest.
+This is implemented in \pkg{hesim} with the \code{hesim_data} class as will be illustrated in the examples below.
+
+The actual analysis is performed in the next three steps as shown in Figure~\ref{fig:econ-eval-process-hesim}.
+First, a statistical models for disease progression, costs, and utilities are parameterized using ``estimation'' datasets.
+Next, the statistical models are combined to construct an economic model.
+For a given model structure, disease progression, QALYs, and costs are simulated from ``input data'', based on the target population and treatment strategies of interest.
+Finally, the simulated outcomes are used to perform decision analysis, currently limited to CEA.
 
@@
 \end{figure}
+
+The three economic models described in Section~\ref{sec:model-taxonomy} are implemented as the \pkg{R6} classes \code{CohortDtstm}, \code{IndivCtstm}, and \code{Psm}.
+Methods are provided within each class for simulating disease progression, costs, and QALYs.
+These methods are made possible by separate \pkg{R6} classes for disease progression, costs, and utility, which comprise the full economic model.
+Below, we describe this modular design in more detail.
+
+\subsection{Parameterization} \label{sec:parameterization}
+\subsubsection{Disease progression}
+The statistical model used to parameterize the disease model component of an economic model varies by the type of economic model.
+For example, multinomial logistic regressions can be used to parameterize a cDTSTM, a set of N-1 independent survival models are used to parameterize an N-state partitioned survival model, and multi-state models can be used to parameterize an iCTSTM (Table~\ref{tbl:parameterize-disease-model}).
+\begin{table} 
+\caption{Parameterization of disease models}\label{tbl:parameterize-disease-model}
+\footnotesize
+\begin{tabular}{l l l l}
+\hline
+Economic model & Statistical model & Parameter object & Model fit object\\
+\hline
+\multirow{2}{*}{\code{CohortDtstm}}& Custom & \code{tparams_transprobs} & \code{define_model()}\\
+ & Multinomial logistic regressions & \code{params_mlogit} & \code{multinom_list}\\
+\multirow{2}{*}{\code{IndivCtstm}} & Multi-state model (joint likelihood) & \code{params_surv} & \code{flexsurv::flexsurvreg}\\
+& Multi-state model (transition-specific) & \code{params_surv_list} & \code{flexsurvreg_list}\\
+\code{Psm} & Independent survival models & \code{param_surv_list} & \code{flexsurvreg_list}\\
+\hline
+\end{tabular}
+\end{table}
 

### 2020-11-02 manuscript.Rnw
Updates to the coding framework section

@@
 \section{Framework} \label{sec:framework}
-The \pkg{hesim} coding framework aims to follow three main tenets.
-First, it should be possible to seamlessly integrate parameter estimation with simulation.
-Second, modular code show allow for models to be built in a flexible manner.
-Finally, code should be computationally efficient.
+Economic models consists of separate submodels of disease progression, costs, and utilities.
+Model development proceed in 4 steps.
+The first step is to setup the model by defining the model structure, target population, and treatment strategies of interest.
+This is implemented with the \code{hesim_data} class as will be illustrated in the examples below.
 
-Each economic model consists of separate statistical models of disease progression, costs, and utilities.
-All analyses proceed in 4 steps.
-The first step is to setup the model by defining the model structure, target population, and treatment strategies of interest.
-This is implemented in \pkg{hesim} with the \code{hesim_data} class as will be illustrated in the examples below.
-
-The actual analysis is performed in the next three steps as shown in Figure~\ref{fig:econ-eval-process-hesim}.
-First, a statistical models for disease progression, costs, and utilities are parameterized using ``estimation'' datasets.
-Next, the statistical models are combined to construct an economic model.
+The analysis is performed in the next three steps as shown in Figure~\ref{fig:econ-eval-process-hesim}.
+First, models for disease progression, costs, and utilities are parameterized using ``estimation'' datasets.
+Next, the submodels are combined to construct an economic model.
 For a given model structure, disease progression, QALYs, and costs are simulated from ``input data'', based on the target population and treatment strategies of interest.
@@
 Methods are provided within each class for simulating disease progression, costs, and QALYs.
-These methods are made possible by separate \pkg{R6} classes for disease progression, costs, and utility, which comprise the full economic model.
+These methods are made possible by separate \pkg{R6} classes for the disease, costs, and utility submodels.
 Below, we describe this modular design in more detail.
@@
 \subsection{Parameterization} \label{sec:parameterization}
+Each submodel contains fields for the model parameters and the input data.
+Parameters can either be created from a statistical model fit using \proglang{R} (e.g., a survival model) or input directly by the user.
+There are two types of parameter objects, standard parameter objects prefixed by \code{``params''} and transformed parameter objects prefixed by \code{``tparams''}.
+The former contain the underlying parameters of a statistical model while the latter contain parameters more immediate to prediction that have been transformed as function of the input data.
+The coefficients from a Weibull regression model are an example of a parameter object and the predicted shape and scale parameters of the Weibull distribution are an example of a transformed parameter object.
+
+Although economic models are ideally parameterized using explicit statistical models, it is not always feasible.
+Parameters are often taken from a variety of sources or from summary level data.
+In these cases, it can be convenient to build models using mathematical expressions.
+This can be done when building cDTSTMs using \code{defime_model()}.
+
 \subsubsection{Disease progression}
-The statistical model used to parameterize the disease model component of an economic model varies by the type of economic model.
+The statistical model used to parameterize the disease model depends on the type of economic model.
 For example, multinomial logistic regressions can be used to parameterize a cDTSTM, a set of N-1 independent survival models are used to parameterize an N-state partitioned survival model, and multi-state models can be used to parameterize an iCTSTM (Table~\ref{tbl:parameterize-disease-model}).
@@
 \footnotesize
-\begin{tabular}{l l l l}
+\begin{tabular*}{\textwidth}{l l l l}
 \hline
@@
 \hline
-\end{tabular}
+\end{tabular*}
 \end{table}
+
+The parameters of a survival model are stored in a \code{params_surv} object and a \code{params_surv_list} can be used to store the parameters of multiple survival models.
+The latter is useful for storing the parameters of a mutli-state model or the independent survival models required for a PSM.
+The parameters of a multinomial logistic regression are stored in a \code{params_mlogit} and can be created by fitting a model with \code{nnet::multinom()}.
 

### 2020-11-07 manuscript.Rnw
Update 'Framework' section

@@
 Markov and semi-Markov models can be formulated in either continuous or discrete time, at either the cohort or individual level.
-In health economics, Markov cohort models tend to be formulated in discrete time (i.e., as cDTSTMs), although state probabilities can be computed in continuous time models using the Aalen-Johansen estimator \citep{aalen1978empirical}.
+In health economics, Markov cohort models tend to be formulated in discrete time (i.e., as cDTSTMs), although state probabilities can be computed in continuous time models using the Aalen-Johansen estimator \citep{aalen1978empirical} or the Kolmogorov forward equation \citep{cox1977theory}.
 While tunnel states can be used to approximate a semi-Markov model using a cohort approach, they can only be simulated in a general fashion using individual level models.
@@
 \section{Framework} \label{sec:framework}
-Economic models consists of separate submodels of disease progression, costs, and utilities.
+Economic models consist of a disease model, a utility model, and a set of cost models for each cost category.
 Model development proceed in 4 steps.
-The first step is to setup the model by defining the model structure, target population, and treatment strategies of interest.
-This is implemented with the \code{hesim_data} class as will be illustrated in the examples below.
+The first step is to setup the model by defining the model structure, target population, and treatment strategies of interest, and storing the relevant information in a \code{hesim_data} object.
 
 The analysis is performed in the next three steps as shown in Figure~\ref{fig:econ-eval-process-hesim}.
-First, models for disease progression, costs, and utilities are parameterized using ``estimation'' datasets.
+First, the disease progression, costs, and utility submodels are parameterized using ``estimation'' datasets.
 Next, the submodels are combined to construct an economic model.
-For a given model structure, disease progression, QALYs, and costs are simulated from ``input data'', based on the target population and treatment strategies of interest.
+The economic model is then used to simulate, disease progression, QALYs, and costs as a function of ``input data''.
+The input data always contains variables describing the target population and treatment strategies of interest, but can also contain variables related to time to facilitate use of time-varying covariates or parameters.
 Finally, the simulated outcomes are used to perform decision analysis, currently limited to CEA.
+All models are simulated using PSA so that decision uncertainty can be represented.
 
@@
 The three economic models described in Section~\ref{sec:model-taxonomy} are implemented as the \pkg{R6} classes \code{CohortDtstm}, \code{IndivCtstm}, and \code{Psm}.
-Methods are provided within each class for simulating disease progression, costs, and QALYs.
-These methods are made possible by separate \pkg{R6} classes for the disease, costs, and utility submodels.
-Below, we describe this modular design in more detail.
+Each class contains methods for simulating disease progression, QALYs, and costs These methods are made possible by separate \pkg{R6} classes for the disease, utility, and cost submodels.
 
@@
 Each submodel contains fields for the model parameters and the input data.
-Parameters can either be created from a statistical model fit using \proglang{R} (e.g., a survival model) or input directly by the user.
+Parameters can either be created from a statistical model fit using \proglang{R} or input directly by the user.
 There are two types of parameter objects, standard parameter objects prefixed by \code{``params''} and transformed parameter objects prefixed by \code{``tparams''}.
@@
 For example, multinomial logistic regressions can be used to parameterize a cDTSTM, a set of N-1 independent survival models are used to parameterize an N-state partitioned survival model, and multi-state models can be used to parameterize an iCTSTM (Table~\ref{tbl:parameterize-disease-model}).
-\begin{table} 
+
+\begin{table} [h]
 \caption{Parameterization of disease models}\label{tbl:parameterize-disease-model}
@@
 \hline
-Economic model & Statistical model & Parameter object & Model fit object\\
+Economic model & Statistical model & Parameter object & Model object\\
 \hline
@@
 The parameters of a survival model are stored in a \code{params_surv} object and a \code{params_surv_list} can be used to store the parameters of multiple survival models.
-The latter is useful for storing the parameters of a mutli-state model or the independent survival models required for a PSM.
+The latter is useful for storing the parameters of a multi-state model or the independent survival models required for a PSM.
 The parameters of a multinomial logistic regression are stored in a \code{params_mlogit} and can be created by fitting a model with \code{nnet::multinom()}.
+
+Although disease models are ideally parameterized by fitting a single statistical model, this is often infeasible.
+Parameters may be taken from multiple sources, frequently using summary level data.
+In these cases it can be more convenient to define transition probability matrices using mathematical expressions relating parameters and input data to transformed parameters.
+Transition probabilities for cDTSTMs can be constructed in this manner using \code{define_model()}.
+
+\subsubsection{Costs and utility} 
+State values (i.e., costs and utilities) do not depend on the choice of disease model.
+They can currently either be modeled using a linear model or using predicted means.
+The latter is an example of a transformed parameter object since the predicted means are parameters that are presumably a function of the underlying parameters of a statistical model and input data.
+
+\begin{table} [h]
+\caption{Parameterization of state value models}\label{tbl:parameterize-disease-model}
+\footnotesize
+\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l l l l}
+\hline
+Statistical model & Parameter object & Model object\\
+\hline
+\multirow{2}{*}{Predicted means} & \code{tparams_mean} & \code{stateval_tbl}\\
+ & \code{tparams_mean} & \code{define_model()}\\
+Linear model & \code{params_lm} & \code{stats::lm}\\
+\hline
+\end{tabular*}
+\end{table}
+
+The most straightforward way to construct state values is with a \code{stateval_tbl}, which is a special object used to assign values to health states that can vary across PSA samples, treatment strategies, patients, and/or time intervals.
+State values can be specified either as moments (i.e., mean and standard error) or parameters (e.g., shape and scale of gamma distribution) of a probability distribution, or by pre-simulating values from a suitable probability distribution (e.g., from a Bayesian model).
+Like transition probabilities, state value can also be constructed with \code{define_model}.
+
+\subsection{Simulation} \label{sec:simulation}
+The utility and cost models are always \pkg{R6} objects of class \code{StateVals}, whereas the disease models vary by economic model (Table~\ref{tbl:submodels}).
+The disease model is used to simulate survival curves in a PSM and health state transitions in a cDTSTM and iCTSTM.
+
+\begin{table} [h]
+\caption{Submodels comprising an economic model}\label{tbl:submodels}
+\footnotesize
+\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l l l l}
+\hline
+Economic model & Disease model & Utility model & Cost model(s)\\
+\hline
+\code{CohortDtstm} & \code{CohortDtstmTrans} & \code{StateVals} & \code{StateVals}\\
+\code{Psm} & \code{PsmCurves} & \code{StateVals} & \code{StateVals}\\
+\code{IndivCtstm} & \code{IndivCtstmTrans} & \code{StateVals} & \code{StateVals}\\
+\hline
+\end{tabular*}
+\end{table}
+
+The disease and state value models are most easily instantiated using \code{S3} generic functions prefixed by ``\code{create}'' from parameter, transformed parameter, or statistical model objects.
+An \code{IndivCtstmTrans} can, for instance, be created from a \code{flexsurvreg_list} or \code{params_surv_list} object with \code{create_IndivCtstmTrans()}.
+Similarly, a \code{StateVals} object can, for example, be created from a \code{stateval_tbl} object with \code{create_StateVals()}.
+The complete economic model is instantiated by combining the submodels using the \pkg{R6} constructor method \code{$new()}.
+
+Each economic model contains methods for simulating disease progression, QALYs, and costs (Table~\ref{tbl:simulate-outcomes}).
+The cost and utility models always simulate costs and QALYs from the simulated  progression of disease with the methods \code{$sim_qalys()} and \code{$sim_costs()}, respectively.
+Methods for simulating disease progression differ slightly since the simulation approaches differ as described in Section~\ref{sec:model-taxonomy}.
+In an N-state PSM $n-1$ survival curves are generated and in a iCTSTM a trajectory through state transitions is simulated for multiple patients via individual simulation.
+All models can simulate state probabilities over time.
+
+\begin{table} [h]
+\caption{Methods for simulating outcomes from economic models}\label{tbl:simulate-outcomes}
+\footnotesize
+\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l l l l}
+\hline
+Economic model & Disease progression  & QALYs & Costs \\
+\hline
+\code{CohortDtstm} & \code{sim_stateprobs()} & \code{sim_qalys()} & \code{sim_costs()}\\
+\code{Psm} & \code{sim_survival()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
+\code{IndivCtstm} & \code{sim_disease()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
+\hline
+\end{tabular*}
+\end{table}
 

### 2020-11-08 manuscript.Rnw
Add all sections (even if no content yet)

@@
 Multiple model types are supported including cohort discrete time state transition models (DTSTMs), N-state PSMs, and individual-level continuous time state transition models (CTSTMs), encompassing both Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
-To maximize flexibility and faciliate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R} or by inputting parameters obtained from external sources.
+To maximize flexibility and facilitate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R} or by inputting parameters obtained from external sources.
 So that individual-level simulation and PSA can be run quickly, \pkg{Rcpp} and \pkg{data.table} are heavily utilized.
 After simulating costs and quality-adjusted life-years (QALYs) from a PSA, decision analysis can be performed within a cost-effectiveness framework.
+
+The remainder of this article is organized as follows.
+Section~\ref{sec:model-taxonomy} describes the economic models supported by \pkg{hesim}.
+An overview of the coding framework is provided in Section~\ref{sec:framework}.
+The next two sections contain example analyses.
+Section~\ref{sec:modeling-aggregate} shows how both cohort and individual models can be used in the common scenario where only aggregate level data is available.
+Section~\ref{sec:modeling-patient} moves to a case where richer patient level data is available and shows how partitioned survival analysis and multi-state modeling can be employed.
+The cost-effectiveness framework is described in Section~\ref{sec:cea} along with worked examples.
+Section~\ref{sec:discussion} makes comparisons to other software and discusses possible extensions.
+Finally, Section~\ref{sec:conclusion} concludes.
 
@@
 
+%% -- Models with aggregate level data -----------------------------------------
+\section{Economic modeling with aggregate level data} \label{sec:modeling-aggregate}
+
+\subsection{Cohort model}
+
+\subsection{Individual model}
+
+%% -- Models with patient level data -------------------------------------------
+\section{Economic modeling with patient level data} \label{sec:modeling-patient}
+
+\subsection{Partitioned survival analysis}
+
+\subsection{Multi-state model}
+
+%% -- Cost-effectiveness analysis ----------------------------------------------
+\section{Cost-effectiveness analysis} \label{sec:cea}
+
+%% -- Discussion ---------------------------------------------------------------
+\section{Discussion} \label{sec:discussion}
+
+%% -- Conclusion ---------------------------------------------------------------
+\section{Conclusion} \label{sec:conclusion}
+
 %% -- Bibliography -------------------------------------------------------------

### 2020-11-10 manuscript.Rnw
Replace 3-step process figure with LaTeX tikz figure

@@
 \usepackage{multirow}
+\usepackage{tikz}
+\usetikzlibrary{positioning}
 \usepackage{caption} 
@@
 The input data always contains variables describing the target population and treatment strategies of interest, but can also contain variables related to time to facilitate use of time-varying covariates or parameters.
-Finally, the simulated outcomes are used to perform decision analysis, currently limited to CEA.
+Finally, the simulated outcomes are used to perform decision analysis within a cost-effectiveness framework.
+While other approaches such as multi-criteria decision analysis (MCDA) could, in principle, be used as well, only CEA is currently supported.
 All models are simulated using PSA so that decision uncertainty can be represented.
@@
 \centering
-\includegraphics[width = 15cm]{econ-eval-process-hesim.png}
+\begin{tikzpicture}
+% Parameterization
+\node[draw, minimum height=4cm, minimum width=3.5cm, font=\footnotesize, text depth = 3.5cm] (A) at (-5,0) {\textbf{1. Parameterization}};
+\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (-5, .8){Disease model};
+\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (-5, -.2){Utility model};
+\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (-5, -1.2){Cost model(s)};
+
+% Simulation
+\node[draw, minimum height=4cm, minimum width=3.5cm, font=\footnotesize, text depth = 3.5cm] (B) at (0,0) {\textbf{2. Simulation}};
+\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (0, .8){Disease model};
+\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (0, -.2){Utility model};
+\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (0, -1.2){Cost model(s)};
+
+% Decision analysis
+\node[draw, minimum height=4cm, minimum width=3.5cm, font=\footnotesize, text depth = 3.5cm] (C) at (5,0) {\textbf{3. Decision analysis}};
+\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (5, .3){CEA};
+\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (5, -.7){MCDA};
+
+% Data
+\node[draw, above = .5 of A, minimum width=3cm, font=\footnotesize] (D) {Estimation data};
+\node[draw, above = .5 of B, minimum width=3cm, font=\footnotesize] (E) {Input data};
+
+% Arrows
+\draw [thick, ->] (A.east) -- (B.west);
+\draw [thick, ->] (B.east) -- (C.west);
+\draw [thick, ->] (D.south) -- (A.north);
+\draw [thick, ->] (E.south) -- (B.north);
+
+\end{tikzpicture}
 \caption{Economic modeling process}\label{fig:econ-eval-process-hesim}
@@
 The three economic models described in Section~\ref{sec:model-taxonomy} are implemented as the \pkg{R6} classes \code{CohortDtstm}, \code{IndivCtstm}, and \code{Psm}.
-Each class contains methods for simulating disease progression, QALYs, and costs These methods are made possible by separate \pkg{R6} classes for the disease, utility, and cost submodels.
+Each class contains methods for simulating disease progression, QALYs, and costs.
+These methods are made possible by separate \pkg{R6} classes for the disease, utility, and cost submodels.
+The remainder of this section describes the framework for parameterization and simulation.
+The cost-effectiveness framework is described in Section~\ref{sec:cea}.
 

### 2020-11-10 manuscript.Rnw
Add figure for 4-state disease progression model

@@
 
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}
+% States
+\node[draw, minimum height=1cm, minimum width=2cm, font=\footnotesize] (S1) {\textbf{State 1}};
+\node[draw, minimum height=1cm, minimum width=2cm, font=\footnotesize] (S2) [right = of S1] {\textbf{State 2}};
+\node[draw, minimum height=1cm, minimum width=2cm, font=\footnotesize] (S3) [right = of S2] {\textbf{State 3}};
+\node[draw, minimum height=1cm, minimum width=2cm, font=\footnotesize] (D) [below = of S2] {\textbf{Death}};
+
+% Arrows
+\draw [thick, <->] (S1.east) -- (S2.west);
+\draw [thick, ->] (S2.east) -- (S3.west);
+\draw [thick, ->] (S1.south) -- (D.west);
+\draw [thick, ->] (S2.south) -- (D.north);
+\draw [thick, ->] (S3.south) -- (D.east);
+
+\end{tikzpicture}
+\caption{Four state disease progression model}\label{fig:econ-eval-process-hesim}
+\end{figure}
+
 \subsection{Cohort model}

### 2020-11-17 manuscript.Rnw
Start on first example model

@@
 %% need no \usepackage{Sweave}
-\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
+\SweaveOpts{engine = R, eps = FALSE, keep.source = TRUE, eval = TRUE}
 <<preliminaries, echo=FALSE, results=hide>>=
@@
 \section{Economic modeling with aggregate level data} \label{sec:modeling-aggregate}
+Health economic models are commonly developed using aggregate level data taken from multiple sources.
+In these cases it is helpful to be able input model parameters rather than fitting a statistical model.
+We illustrate with the 4 state disease progression model displayed in Figure~\ref{fig:four-state-model}.
+Patients advance between successively more severe disease states.
+It is possible to move from State 2 back to State 1 but recovery is not possible from State 3.
+Death can occur from any state.
 
@@
 \end{tikzpicture}
-\caption{Four state disease progression model}\label{fig:econ-eval-process-hesim}
+\caption{Four state disease progression model}\label{fig:four-state-model}
 \end{figure}
 
+There are 3 treatment strategies: standard of care (SOC) and two new interventions (New 1 and New 2).
+For illustration, we will consider a target population approximated by two representative patients of age 60 and 70.
+The 60-year old patient represents $1/3$ of all patients with the disease while the 70-year old patient represents $2/3$.
+The \code{hesim_data} class is used to store this data.
+
+<<>>=
+library("hesim")
+library("data.table")
+strategies <- data.table(
+  strategy_id = 1:3,
+  strategy_name = c("SOC", "New 1", "New 2")
+)
+patients <- data.table(
+  patient_id = 1:2,
+  grp_id = 1:2,
+  age = c(60, 70),
+  patient_wt = c(1/3, 2/3)
+)
+hesim_dat <- hesim_data(
+  strategies = strategies,
+  patients = patients
+)
+print(hesim_dat)
+@
+
 \subsection{Cohort model}
+\subsubsection{Parameterization}
+Model parameters related to transition probabilities are obtained from a hypothetical analysis of a natural history cohort.
+If data is not collected at unevenly spaced intervals, then transitions can be described with transition intensity matrices, $Q$, by fitting a continuous time multi-state model designed for panel data (e.g., with the \proglang{R} package \pkg{msm}) \citep{jackson2011multi}.
+We assume that regression coefficients and a covariance matrix for $Q$ are available from such an analysis.
+
+<<q_coef, echo = FALSE>>=
+trans <- c("1->2", "1->4", "2->1", "2->3", "3->4")
+q_coef_int <- log(c(.9, .2, .5, 1.1, 1.5))
+q_coef_time <- rep(log(1.2), 5)
+q_coef_age <- rep(log(1.05), 5)
+q_coef_soc <- data.table(
+  trans = rep(trans, 3),
+  var = rep(c("intercept", "time[2,Inf)", "age"), each = 5),
+  loghr = c(q_coef_int, q_coef_time, q_coef_age)
+)
+@
+
+<<q_vcov, echo = FALSE>>=
+n_vars <- nrow(q_coef_soc)
+A <- matrix(runif(n_vars^2) * 2 - 1, ncol = 15)/30
+q_vcov_soc <- t(A) %*% A
+@
+
+<<q_vcov_print>>=
+q_vcov_soc[1:4, 1:4]
+@
+
+If many studies are available, then relative treatment effects can be estimated by means of a network meta-analysis.
+We assume that an NMA was performed to estimate the relative risk of transitions from state 1 to both state 2 and death.
+Treatment strategy 1 was the reference arm.
+Estimation is performed in terms of the log relative risk since it is approximately normally distributed.
+
+The utility associated with each state is assumed to follow a beta distribution, which is bounded between 0 and 1.
+The beta distribution is defined in terms of 2 shape parameters, but can be derived from the mean and standard error using the method of moments.
+Point estimates for utility in states 1, 2, and 3 are 0.9, 0.75, and 0.5, respectively.
+Standard errors in the same three states are 0.02, 0.03, and 0.05.
+Utility is always assumed to be 0 in death states in \pkg{hesim}.
+
+<<params>>=
+params <- list(
+  q_coef_soc = q_coef_soc,
+  q_vcov_soc = q_vcov_soc,
+  lrr_est = c(log(.80), log(.71)),
+  lrr_se = c(log(.06), log(.08)),
+  u_mean = c(.9, .75, .05),
+  u_se = c(.02, .03, .05)
+)
+@
 

### 2020-11-18 manuscript.Rnw
More updates to parameterization of first example model

@@
 Parameters are often taken from a variety of sources or from summary level data.
-In these cases, it can be convenient to build models using mathematical expressions.
+In these cases, it can be convenient to build models with mathematical expressions usin non-standard evaluation.
 This can be done when building cDTSTMs using \code{defime_model()}.
@@
 \subsubsection{Parameterization}
-Model parameters related to transition probabilities are obtained from a hypothetical analysis of a natural history cohort.
-If data is not collected at unevenly spaced intervals, then transitions can be described with transition intensity matrices, $Q$, by fitting a continuous time multi-state model designed for panel data (e.g., with the \proglang{R} package \pkg{msm}) \citep{jackson2011multi}.
+Model parameters related to transition probabilities for SOC are obtained from a hypothetical analysis of a historical cohort.
+If data is collected at unevenly spaced intervals, then transitions can be described with transition intensity matrices, $Q$, by fitting a continuous time multi-state model designed for panel data (e.g., with the \proglang{R} package \pkg{msm}) \citep{jackson2011multi}.
 We assume that regression coefficients and a covariance matrix for $Q$ are available from such an analysis.
@@
 
-<<q_vcov, echo = FALSE>>=
+<<q_coef_soc_print>>=
+q_coef_soc
+@
+
+<<q_vcov_soc, echo = FALSE>>=
 n_vars <- nrow(q_coef_soc)
@@
 
-<<q_vcov_print>>=
+<<q_vcov_soc_print>>=
 q_vcov_soc[1:4, 1:4]
@@
 
-If many studies are available, then relative treatment effects can be estimated by means of a network meta-analysis.
-We assume that an NMA was performed to estimate the relative risk of transitions from state 1 to both state 2 and death.
-Treatment strategy 1 was the reference arm.
+If many studies are available, then relative treatment effects can be estimated by means of a network meta-analysis (NMA).
+We assume that a NMA was performed to estimate the relative risk of transitions from State 1 to both State 2 and Death.
+The SOC was the reference arm.
 Estimation is performed in terms of the log relative risk since it is approximately normally distributed.
@@
 The beta distribution is defined in terms of 2 shape parameters, but can be derived from the mean and standard error using the method of moments.
-Point estimates for utility in states 1, 2, and 3 are 0.9, 0.75, and 0.5, respectively.
+Point estimates for utility in State 1, State 2, and State 3 are 0.9, 0.75, and 0.5, respectively.
 Standard errors in the same three states are 0.02, 0.03, and 0.05.
-Utility is always assumed to be 0 in death states in \pkg{hesim}.
+Utility (and costs) are always assumed to be 0 in the death state.
+
+We assume that two categories of costs are measured: treatment and medical.
+The former are the costs of the three treatment strategies and are assumed fixed.
+The SOC is the cheapest option at \$2,000 per annum while New 1 and New 2 cost \$10,000 and \$12,000 per year, respectively.
+
+While there is no uniform way to define medical costs, they often consist of costs accrued from hospitalizations and inpatient visits.
+They are assumed to follow a gamma distribution given that they are usually heavily right skewed.
+Like the beta distribution, the parameters of the gamma distribution are derived from the mean and standard error using the method of moments.
+Mean medical costs are \$2,000, \$4,000, and \$15,000 in State 1, State 2, and State 3, respectively.
+We assume the variance  is unavailable and set the standard error in each state equal to the mean.
 
@@
   lrr_est = c(log(.80), log(.71)),
-  lrr_se = c(log(.06), log(.08)),
-  u_mean = c(.9, .75, .05),
-  u_se = c(.02, .03, .05)
+  lrr_se = c(.06, .08),
+  u_mean = c(s1 = .9, s2 = .75, s3 = .05),
+  u_se = c(s1 = .02, s2 = .03, s3 = .05),
+  c_medical = c(s1 = 2000, s2 = 4000, s3 = 15000),
+  c_tx = c(soc = 2000, new1 = 10000, new2 = 12000)
 )
+@
+
+To perform a PSA, it is necessary to randomly draw the parameters from suitable probability distributions.
+These distributions can be defined using \code{define_rng()}.
+While any random number generation function can be used, a number of commonly used distributions are specifically designed for use within a \code{define_rng()} block (see \code{?rng_distributions}).
+The random number generation expressions can either be directly evaluated with \code{eval_rng()} or passed to \code{define_model()} and evaluated during construction of the economic model.
+We will take the latter approach here.
+
+<<define_rng>>=
+rng_def <- define_rng({
+  list( # Parameters to return
+    q_coef = multi_normal_rng(q_coef_soc$loghr, q_vcov_soc),
+    lrr = lognormal_rng(lrr_est, lrr_se),
+    u = beta_rng(mean = u_mean, sd = u_se),
+    c_tx = fixed(c_tx),
+    c_medical = gamma_rng(mean = c_medical, sd = c_medical)
+  )
+}, n = 1000)
+@
+
+<<eval_rng, eval = FALSE, echo = FALSE>>=
+eval_rng(rng_def, params = params)
 @

### 2020-11-28 manuscript.Rnw
First draft of cohort model with aggregate data

@@
 options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+set.seed(101)
 @
@@
 The next two sections contain example analyses.
-Section~\ref{sec:modeling-aggregate} shows how both cohort and individual models can be used in the common scenario where only aggregate level data is available.
-Section~\ref{sec:modeling-patient} moves to a case where richer patient level data is available and shows how partitioned survival analysis and multi-state modeling can be employed.
+Section~\ref{sec:modeling-patient} shows how both partitioned survival and multi-state models can be used in an ideal case where rich patient level data is available.
+Section~\ref{sec:modeling-aggregate} then moves to the common scenario where only aggregate level data is available shows how both cohort and individual models can be employed.
 The cost-effectiveness framework is described in Section~\ref{sec:cea} along with worked examples.
@@
 
+%% -- Models with patient level data -------------------------------------------
+\section{Economic modeling with patient level data} \label{sec:modeling-patient}
+
+\subsection{Partitioned survival analysis}
+
+\subsection{Multi-state model}
+
 %% -- Models with aggregate level data -----------------------------------------
@@
 
+The four health states and five possible transitions are characterized below.
+
+<<q_coef, echo = FALSE>>=
+library("data.table")
+states <- data.table(
+  state_id = c(1, 2, 3, 4),
+  state_name = c(paste0("State ", 1:3), "Death")
+)
+
+tmat <- rbind(
+  c(NA, 1, NA, 2),
+  c(3, NA, 4, NA),
+  c(NA, NA, NA, 5),
+  c(NA, NA, NA, NA)
+)
+colnames(tmat) <- rownames(tmat) <- states$state_name
+n_trans <- max(tmat)
+print(tmat)
+@
+
 There are 3 treatment strategies: standard of care (SOC) and two new interventions (New 1 and New 2).
@@
 The 60-year old patient represents $1/3$ of all patients with the disease while the 70-year old patient represents $2/3$.
-The \code{hesim_data} class is used to store this data.
+The \code{hesim_data} class is again used to store this data as well as the number of non-death health states.
 
-<<>>=
+<<hesim_data>>=
 library("hesim")
-library("data.table")
 strategies <- data.table(
   strategy_id = 1:3,
-  strategy_name = c("SOC", "New 1", "New 2")
+  strategy_name = c("SOC", "New 1", "New 2"),
+  soc = c(1, 0, 0),
+  new1 = c(0, 1, 0),
+  new2 = c(0, 0, 1)
 )
+
 patients <- data.table(
@@
 )
+
 hesim_dat <- hesim_data(
   strategies = strategies,
-  patients = patients
+  patients = patients,
+  states = states[state_name != "Death"]
 )
@@
 \subsubsection{Parameterization}
-Model parameters related to transition probabilities for SOC are obtained from a hypothetical analysis of a historical cohort.
-If data is collected at unevenly spaced intervals, then transitions can be described with transition intensity matrices, $Q$, by fitting a continuous time multi-state model designed for panel data (e.g., with the \proglang{R} package \pkg{msm}) \citep{jackson2011multi}.
-We assume that regression coefficients and a covariance matrix for $Q$ are available from such an analysis.
+The core of a cohort DTSTM is the disease model which is determined by the transition matrices $P_t$.
+In this example, model parameters related to the transition probabilities for SOC are obtained from a hypothetical analysis of a historical cohort.
+If data is collected at unevenly spaced intervals, then transitions can be described with transition intensity matrices, $Q_t$, by fitting a continuous time multi-state model designed for panel data (e.g., with the \proglang{R} package \pkg{msm}) \citep{jackson2011multi}.
+We assume that regression coefficients and a covariance matrix are available from such an analysis and that a piecewise exponential model was fit so that intensities change after year 2.
 
@@
 trans <- c("1->2", "1->4", "2->1", "2->3", "3->4")
+trans_short <- c("12", "14", "21", "23", "34")
+var_short <- c("int", "time", "age") 
+var_long <- c("intercept", "time[2,Inf)", "age") 
 q_coef_int <- log(c(.9, .2, .5, 1.1, 1.5))
 q_coef_time <- rep(log(1.2), 5)
-q_coef_age <- rep(log(1.05), 5)
+q_coef_age <- rep(log(1.003), 5)
 q_coef_soc <- data.table(
   trans = rep(trans, 3),
-  var = rep(c("intercept", "time[2,Inf)", "age"), each = 5),
+  var = rep(var_long, each = 5),
+  name = paste0(rep(var_short, each = 5), "_", rep(trans_short, 3)),
   loghr = c(q_coef_int, q_coef_time, q_coef_age)
@@
 n_vars <- nrow(q_coef_soc)
-A <- matrix(runif(n_vars^2) * 2 - 1, ncol = 15)/30
+A <- matrix(runif(n_vars^2) * 2 - 1, ncol = 15)/60
 q_vcov_soc <- t(A) %*% A
@@
   q_vcov_soc = q_vcov_soc,
-  lrr_est = c(log(.80), log(.71)),
-  lrr_se = c(.06, .08),
+  lrr_12_est = c(soc = log(1), new1 = log(.80), new2 = log(.71)),
+  lrr_12_se = c(soc = 0, new1 = .03, new2 = .04),
+  lrr_14_est = c(soc = log(1), new1 = log(.90), new2 = log(.85)),
+  lrr_14_se = c(soc = 0, new1 = .02, new2 = .03),
   u_mean = c(s1 = .9, s2 = .75, s3 = .05),
@@
 While any random number generation function can be used, a number of commonly used distributions are specifically designed for use within a \code{define_rng()} block (see \code{?rng_distributions}).
-The random number generation expressions can either be directly evaluated with \code{eval_rng()} or passed to \code{define_model()} and evaluated during construction of the economic model.
-We will take the latter approach here.
+The built in distributions do not require specification of the number of PSA iterations (since it is defined within the block) and automatically perform certain operations (e.g., sampling from a gamma distribution from means and standard errors by usin methods of moments to derive the underlying shape and scale parameters).
+We include utility for illustrative purposes but omit the cost parameters since they can be sampled more directly with \code{stateva_tbl()}.
 
@@
   list( # Parameters to return
-    q_coef = multi_normal_rng(q_coef_soc$loghr, q_vcov_soc),
-    lrr = lognormal_rng(lrr_est, lrr_se),
-    u = beta_rng(mean = u_mean, sd = u_se),
-    c_tx = fixed(c_tx),
-    c_medical = gamma_rng(mean = c_medical, sd = c_medical)
+    q_coef_soc = multi_normal_rng(q_coef_soc$loghr, q_vcov_soc,
+                                  names = q_coef_soc$name),
+    rr_12 = lognormal_rng(lrr_12_est, lrr_12_se),
+    rr_14 = lognormal_rng(lrr_14_est, lrr_14_se),
+    u = beta_rng(mean = u_mean, sd = u_se)
   )
@@
 
-<<eval_rng, eval = FALSE, echo = FALSE>>=
-eval_rng(rng_def, params = params)
+The random number generation expressions can are evaluated with \code{eval_rng()}.
+
+<<eval_rng>>=
+params_rng <- eval_rng(rng_def, params = params)
+@
+
+The transition intensity matrices and relative risks can be combined to derive the relevant transition probability matrices, stored as a \code{tparams_transprobs} object.
+A \code{tparams_transprobs} is simply a 3-dimensional array of transition probability matrices with a unique predicted matrix for each PSA sample, treatment strategy, representative patient, and (optionally) time interval.
+As in the prior example, the underlying input data can be constructed with \code{expand}.
+
+<<transmod_data>>=
+transmod_data <- expand(hesim_dat, by = c("strategies", "patients"),
+                        times = 2)
+head(transmod_data)
+@
+
+A complete dataset of ID variables that also incorporate the PSA iterations is constructed with \code{tpmatrix_id()}.
+The objective is to construct a transition probability for each row in this dataset.
+The user has complete flexibility to do this.
+
+<<tpmat_id>>=
+tpmat_id <- tpmatrix_id(transmod_data, rng_def$n)
+head(tpmat_id)
+@
+
+In this example we do this by first using the coefficients and input data to to predict transition intensity matrices for SOC (by age group and time interval).
+This is done by separately predicting intensities for all feasible transitions.
+There are $n=4$ observations for SOC and $B=1000$ random draws of the coefficients for the PSA.
+A transition from state $r$ to state $s$ is modeled as $log(q_{rs}) = X\beta^T$ where $X$ is an $n \times 3$ design matrix and $\beta$ is a $B \times 3$ sample of the coefficients.
+The resulting $n \times B$ matrix can be conveniently flattened into a vector of length $nB$ with the \code{c()} function so that predicted values are stacked by parameter sample as in \code{tpmatrix_id()}.
+
+<<Q_soc>>=
+predict_q_soc <- function(data, beta){
+  data_soc <- data[strategy_name == "SOC"]
+  x <- model.matrix(~factor(time_start) + age, data = data_soc)
+  n_trans <- 5
+  qmat_soc <- matrix(NA, nrow = nrow(beta) * nrow(x), ncol = n_trans)
+  colnames(qmat_soc) <- c("q_12", "q_14", "q_21", "q_23", "q34")
+  for (i in 1:n_trans) {
+     beta_i <- as.matrix(beta)[, seq(i, ncol(beta), n_trans)]
+     qmat_soc[, i] <- exp(c(x %*% t(beta_i)))
+  }
+  return(qmat_soc)
+}
+q_soc <- predict_q_soc(transmod_data, params_rng$q_coef_soc)
+head(q_soc)
+@
+
+The predicted transition intensities are, in turn, use to form the complete transition intensity matrix.
+Specifically, \code{qmatrix()} forms an array of transition intensity matrices where each row sums to $0$ so that $q_{rr} = \sum_{s \neq r} -q_{rs}$.
+
+<<Q_soc>>=
+qmat_soc <- qmatrix(q_soc, tmat)
+qmat_soc[,, 1]
+@
+
+Transition probability matrices are constructed from the transition intensity matrices using the matrix exponential \citep{jackson2011multi}.
+In continuous time, $P(t) = \exp(tQ)$ so a transition probability matrix with model cycles of length $t$ years is given by $\exp(Q_t)$.
+In this example we will assume model cycles are 1/2 of a year.
+
+<<P_soc>>=
+pmat_soc <- expmat(qmat_soc, t = 1/2)
+pmat_soc[,, 1]
+@
+
+Transition probability matrices can be derived for treatment strategies New 1 and New 2 using the relative risks. Relative risks are predicted for each row in the \code{tmat_id} dataset above.
+
+<<rr>>=
+x_rr <- as.matrix(transmod_data[, .(soc, new1, new2)])
+rr_12 <- c(x_rr %*% t(params_rng$rr_12))
+rr_14 <- c(x_rr %*% t(params_rng$rr_14))
+rr <- cbind(rr_12, rr_14)
+@
+
+Given reference transition probability matrices (i.e., those for SOC), we can apply the relative risks for the other treatment strategies using the \code{apply_rr()} function.
+An array of transition probability matrices is returned with one matrix for each row in \code{rr}.
+
+<<apply_rr>>=
+pmat <- apply_rr(pmat_soc, rr = rr,
+                 index = list(c(1, 2), c(1, 4)))
+@
+
+Finally, the array of transition probability matrices is combined with the relevant ID variables indexing each array slice to create the \code{tparams_transprobs} object.
+
+<<apply_rr>>=
+tprobs <- tparams_transprobs(pmat, tpmat_id)
+@
+
+The utility and cost parameters are much more straightforward to store. As in the previous example we consider two cost categories and store parameter values in a \code{stateval_tbl}.
+
+<<tx_costs>>=
+txcost_tbl <- stateval_tbl(
+  data.table(
+    strategy_id = 1:3,
+    est = c(soc = 2000, new1 = 10000, new2 = 12000)
+  ),
+  dist = "fixed",
+  hesim_data = hesim_dat
+)
+head(txcost_tbl)
+@
+
+<<tx_costs>>=
+medcost_tbl <- stateval_tbl(
+  data.table(
+    state_id = 1:3,
+    mean = c(s1 = 2000, s2 = 4000, s3 = 15000),
+    se = c(s1 = 2000, s2 = 4000, s3 = 15000)
+  ),
+  dist = "gamma",
+  hesim_data = hesim_dat
+)
+head(medcost_tbl)
+@
+
+Although we could do the same for utility, we illustrate an alternative approach in which the distribution of utility values has been previously drawn from a suitable probability distribution.
+While not necessary in this case, this feature might be useful in settings where samples have already been drawn from the posterior distribution of a Bayesian model.
+
+<<utility>>=
+utility_tbl <- stateval_tbl(
+  data.table(
+    state_id = rep(c(1, 2, 3), each = rng_def$n),
+    sample = rep(1:rng_def$n, nrow(hesim_dat$states)),
+    value = c(as.matrix(params_rng$u))
+  ),
+  dist = "custom",
+  hesim_data = hesim_dat
+)
+head(utility_tbl)
+@
+
+\subsubsection{Simulation}
+The simulation proceeds the same way as in the prior examples.
+We first create the transition, cost, and utility models, here specifying that model cycles are 6 months long since the model is in discrete time.
+
+<<transmod>>=
+transmod <- CohortDtstmTrans$new(params = tprobs, cycle_length = 1/2)
+@
+
+<<stvalmods>>=
+# Utility
+utilitymod <- create_StateVals(utility_tbl)
+
+# Costs
+txcostmod <- create_StateVals(txcost_tbl, n = rng_def$n)
+medcostmod <- create_StateVals(medcost_tbl, n = rng_def$n)
+costmods <- list(Treatment = txcostmod,
+                 Medical = medcostmod)
+@
+
+<<econmod>>=
+econmod <- CohortDtstm$new(trans_model = transmod,
+                           utility_model = utilitymod,
+                           cost_models = costmods)
+@
+
+State probabilities are simulated for 45 years, or 90 model cycles.
+
+<<sim_stateprobs>>=
+econmod$sim_stateprobs(n_cycles = 45 * 2)
+@
+
+Costs and QALYs are again simulated with 3 percent discount rates.
+
+<<sim_costs_qalys>>=
+econmod$sim_costs(dr = .03)
+econmod$sim_qalys(dr = .03)
+@
+
+Costs and QALYs are first summarized by subgroup.
+
+<<icer_grp>>=
+ce_sim_grp <- econmod$summarize(by_grp = TRUE)
+cea_pw_out <- cea_pw(ce_sim_grp, comparator = 1, dr_costs = .03, dr_qalys = .03)
+icer_tbl(cea_pw_out, output = "data.table")
+@
+
+They can also be aggregated across groups using the weighted specified in the \code{hesim_data} object above.
+
+<<icer>>=
+ce_sim <- econmod$summarize()
+cea_pw_out <- cea_pw(ce_sim, comparator = 1, dr_costs = .03, dr_qalys = .03)
+icer_tbl(cea_pw_out, output = "data.table")
 @
@@
 \subsection{Individual model}
-
-%% -- Models with patient level data -------------------------------------------
-\section{Economic modeling with patient level data} \label{sec:modeling-patient}
-
-\subsection{Partitioned survival analysis}
-
-\subsection{Multi-state model}
 

### 2020-12-15 manuscript.Rnw
Fully working code + results

--- 2020-11-17 manuscript.Rnw
@@
 %% need no \usepackage{Sweave}
-\SweaveOpts{engine = R, eps = FALSE, keep.source = TRUE, eval = TRUE}
-<<preliminaries, echo=FALSE, results=hide>>=
+
+<<include=FALSE>>=
+library(knitr)
+knitr::render_sweave()
+opts_chunk$set(
+  engine="R", tidy = FALSE, eval = TRUE, cache = TRUE,
+  fig.height = 5, fig.width = 7
+)
+@
+
+<<preliminaries, echo=FALSE, results="hide">>=
 options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+ggplot2::theme_set(ggplot2::theme_bw())
+set.seed(102)
 @
@@
 
-%% -- Models with aggregate level data -----------------------------------------
-\section{Economic modeling with aggregate level data} \label{sec:modeling-aggregate}
-Health economic models are commonly developed using aggregate level data taken from multiple sources.
-In these cases it is helpful to be able input model parameters rather than fitting a statistical model.
-We illustrate with the 4 state disease progression model displayed in Figure~\ref{fig:four-state-model}.
-Patients advance between successively more severe disease states.
-It is possible to move from State 2 back to State 1 but recovery is not possible from State 3.
-Death can occur from any state.
+%% -- Illustrative example -----------------------------------------------------
+\section{Illustrative example} \label{sec:illustrative-example}
 
@@
 % States
-\node[draw, minimum height=1cm, minimum width=2cm, font=\footnotesize] (S1) {\textbf{State 1}};
-\node[draw, minimum height=1cm, minimum width=2cm, font=\footnotesize] (S2) [right = of S1] {\textbf{State 2}};
-\node[draw, minimum height=1cm, minimum width=2cm, font=\footnotesize] (S3) [right = of S2] {\textbf{State 3}};
-\node[draw, minimum height=1cm, minimum width=2cm, font=\footnotesize] (D) [below = of S2] {\textbf{Death}};
+\node[draw, minimum height=1cm, minimum width=2.5cm, font=\footnotesize] (S) {\textbf{Stable}};
+\node[draw, minimum height=1cm, minimum width=2.5cm, font=\footnotesize] (P) [right = of S] {\textbf{Progression}};
+\node[draw, minimum height=1cm, minimum width=2.5cm, font=\footnotesize] (D) [right = of P] {\textbf{Death}};
 
 % Arrows
-\draw [thick, <->] (S1.east) -- (S2.west);
-\draw [thick, ->] (S2.east) -- (S3.west);
-\draw [thick, ->] (S1.south) -- (D.west);
-\draw [thick, ->] (S2.south) -- (D.north);
-\draw [thick, ->] (S3.south) -- (D.east);
+\draw [thick, ->] (S.east) -- (P.west);
+\draw [thick, ->] (P.east) -- (D.west);
+\draw [thick, ->] (S.north) to [bend left = 20] (D.north);
 
 \end{tikzpicture}
-\caption{Four state disease progression model}\label{fig:four-state-model}
+\caption{Three state model of disease progression in oncology}\label{fig:three-state-model}
 \end{figure}
 
-There are 3 treatment strategies: standard of care (SOC) and two new interventions (New 1 and New 2).
-For illustration, we will consider a target population approximated by two representative patients of age 60 and 70.
-The 60-year old patient represents $1/3$ of all patients with the disease while the 70-year old patient represents $2/3$.
-The \code{hesim_data} class is used to store this data.
+<<tmat>>=
+tmat <- rbind(
+  c(NA, 1, 2),
+  c(NA, NA, 3),
+  c(NA, NA, NA)
+)
+colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
+print(tmat)
+@
 
-<<>>=
+<<hesim_data>>=
 library("hesim")
 library("data.table")
-strategies <- data.table(
+n_patients <- 1000
+patients <- data.table(
+  patient_id = 1:n_patients,
+  age = rnorm(n_patients, mean = 45, sd = 7),
+  female = rbinom(n_patients, size = 1, prob = .51)
+)
+
+states <- data.table(
+  state_id = c(1, 2),
+  state_name = c("Stable", "Progression") # Non-death health states
+)
+n_states <- nrow(states)
+
+strategies <- data.frame(
   strategy_id = 1:3,
-  strategy_name = c("SOC", "New 1", "New 2")
+  strategy_name = c("SOC", "New 1", "New 2"),
+  soc = c(1, 0, 0),
+  new1 = c(0, 1, 0),
+  new2 = c(0, 0, 1)
 )
-patients <- data.table(
-  patient_id = 1:2,
-  grp_id = 1:2,
-  age = c(60, 70),
-  patient_wt = c(1/3, 2/3)
-)
+n_strategies <- nrow(strategies)
+
 hesim_dat <- hesim_data(
   strategies = strategies,
-  patients = patients
+  patients = patients,
+  states = states
 )
@@
 
-\subsection{Cohort model}
+%% -- Individual continuous time state transition model ------------------------
+\subsection{Individual continuous time state transition model}
+
 \subsubsection{Parameterization}
-Model parameters related to transition probabilities are obtained from a hypothetical analysis of a natural history cohort.
-If data is not collected at unevenly spaced intervals, then transitions can be described with transition intensity matrices, $Q$, by fitting a continuous time multi-state model designed for panel data (e.g., with the \proglang{R} package \pkg{msm}) \citep{jackson2011multi}.
-We assume that regression coefficients and a covariance matrix for $Q$ are available from such an analysis.
 
-<<q_coef, echo = FALSE>>=
-trans <- c("1->2", "1->4", "2->1", "2->3", "3->4")
-q_coef_int <- log(c(.9, .2, .5, 1.1, 1.5))
-q_coef_time <- rep(log(1.2), 5)
-q_coef_age <- rep(log(1.05), 5)
-q_coef_soc <- data.table(
-  trans = rep(trans, 3),
-  var = rep(c("intercept", "time[2,Inf)", "age"), each = 5),
-  loghr = c(q_coef_int, q_coef_time, q_coef_age)
+<<ictstm_data>>=
+onc3[patient_id %in% c(1, 2)]
+@
+
+<<ictstm_fits>>=
+library("flexsurv")
+n_trans <- max(tmat, na.rm = TRUE) 
+wei_fits <- vector(length = n_trans, mode = "list")
+f <- as.formula(Surv(time, status) ~ factor(strategy_name) + female + age)
+for (i in 1:length(wei_fits)){
+  if (i == 3) f <- update.formula(f, .~.-factor(strategy_name)) 
+  wei_fits[[i]] <- flexsurvreg(f, data = onc3, 
+                               subset = (transition_id == i),
+                               dist = "weibull")
+}
+wei_fits <- flexsurvreg_list(wei_fits)
+@
+
+<<ictstm_utility>>=
+utility_tbl <- stateval_tbl(
+  data.table(state_id = states$state_id,
+             mean = c(.8, .6),
+             se = c(0.02, .05)
+  ),
+  dist = "beta",
+  hesim_data = hesim_dat)
+print(utility_tbl)
+@
+
+<<ictstm_medcosts>>=
+medcost_tbl <- stateval_tbl(
+  data.table(state_id = states$state_id,
+             mean = c(2000, 9500),
+             se = c(2000, 9500)
+  ),
+  dist = "gamma",
+  hesim_data = hesim_dat)
+print(medcost_tbl)
+@
+
+<<ictstm_drugcost_setup, echo=FALSE, results="hide">>=
+n_times <- 2
+n_i <- n_states * n_times
+drugcost_dt <-  data.table(
+    strategy_id = rep(strategies$strategy_id, each = n_i),
+    state_id = rep(rep(states$state_id, each = n_times), n_strategies),
+    time_start = rep(c(0, 3/12), n_states * n_strategies),
+    est = c(rep(2000, n_i -1), 1500, 
+            rep(12000, n_i -1), 1000,
+            rep(15000, n_i - 1), 1200
+            ) 
 )
@@
 
-<<q_vcov, echo = FALSE>>=
-n_vars <- nrow(q_coef_soc)
-A <- matrix(runif(n_vars^2) * 2 - 1, ncol = 15)/30
-q_vcov_soc <- t(A) %*% A
-@
-
-<<q_vcov_print>>=
-q_vcov_soc[1:4, 1:4]
-@
-
-If many studies are available, then relative treatment effects can be estimated by means of a network meta-analysis.
-We assume that an NMA was performed to estimate the relative risk of transitions from state 1 to both state 2 and death.
-Treatment strategy 1 was the reference arm.
-Estimation is performed in terms of the log relative risk since it is approximately normally distributed.
-
-The utility associated with each state is assumed to follow a beta distribution, which is bounded between 0 and 1.
-The beta distribution is defined in terms of 2 shape parameters, but can be derived from the mean and standard error using the method of moments.
-Point estimates for utility in states 1, 2, and 3 are 0.9, 0.75, and 0.5, respectively.
-Standard errors in the same three states are 0.02, 0.03, and 0.05.
-Utility is always assumed to be 0 in death states in \pkg{hesim}.
-
-<<params>>=
-params <- list(
-  q_coef_soc = q_coef_soc,
-  q_vcov_soc = q_vcov_soc,
-  lrr_est = c(log(.80), log(.71)),
-  lrr_se = c(log(.06), log(.08)),
-  u_mean = c(.9, .75, .05),
-  u_se = c(.02, .03, .05)
+<<ictstm_drugcost>>=
+drugcost_tbl <- stateval_tbl(
+  drugcost_dt,
+  dist = "fixed",
+  hesim_data = hesim_dat
 )
@@
 
-\subsection{Individual model}
+\subsubsection{Simulation}
+<<ictstm_n_samples>>=
+n_samples <- 100
+@
 
-%% -- Models with patient level data -------------------------------------------
-\section{Economic modeling with patient level data} \label{sec:modeling-patient}
+<<ictstm_transmod_data>>=
+transmod_data <- expand(hesim_dat,
+                        by = c("strategies", "patients"))
+head(transmod_data)
+@
 
-\subsection{Partitioned survival analysis}
+<<ictstm_transmod>>=
+transmod <- create_IndivCtstmTrans(wei_fits, transmod_data,
+                                   trans_mat = tmat, n = n_samples,
+                                   clock = "reset",
+                                   start_age = patients$age)
+@
 
-\subsection{Multi-state model}
+<<ictstm_stvalmods>>=
+# Utility
+utilitymod <- create_StateVals(utility_tbl, n = n_samples)
+
+# Costs
+drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples,
+                                time_reset = TRUE)
+medcostmod <- create_StateVals(medcost_tbl, n = n_samples)
+costmods <- list(Drug = drugcostmod,
+                 Medical = medcostmod)
+@
+
+<<ictstm_econmod>>=
+ictstm <- IndivCtstm$new(trans_model = transmod,
+                         utility_model = utilitymod,
+                         cost_models = costmods)
+@
+
+<<ictstm_sim_disease>>=
+ictstm$sim_disease(max_age = 100)
+head(ictstm$disprog_)
+@
+
+<<ictstm_sim_stateprobs>>=
+ictstm$sim_stateprobs(t = seq(0, 30 , 1/12))
+head(ictstm$stateprobs_)
+@
+
+<<ictstm_summarize_stateprobs, echo=FALSE, results="hide">>=
+add_state_name <- function(x){
+  x[, state_name := factor(state_id,
+                           levels = 1:nrow(tmat),
+                           labels = colnames(tmat))]
+}
+
+add_strategy_name <- function(x) {
+  x[, strategy_name := factor(strategy_id, 
+                              labels = c("SOC", "New 1", "New 2"))]
+}
+
+summarize_stprobs <- function(stateprobs){
+  x <- stateprobs[, .(prob_mean = mean(prob)),
+                  by = c("strategy_id", "state_id", "t")]
+  add_state_name(x)
+  add_strategy_name(x)
+  return(x[, ])
+}
+@
+
+<<ictstm_plot_stateprobs, echo = FALSE, include = FALSE>>=
+library("ggplot2")
+stprobs <- summarize_stprobs(ictstm$stateprobs_)
+ggplot(stprobs, 
+       aes(x = t, y = prob_mean, col = strategy_name)) +
+  geom_line() + facet_wrap(~state_name) +
+  xlab("Years") + ylab("Probability in health state") +
+  scale_color_discrete(name = "Strategy") +
+  theme(legend.position = "bottom") 
+@
+
+\begin{figure}[t!]
+\centering
+\includegraphics{figure/ictstm_plot_stateprobs-1.pdf}
+\caption{Mean simulated state probabilities from the probabilistic sensitivity analysis} \label{fig:ictstm-stprobs}
+\end{figure}
+
+<<ictstm_sim_qalys>>=
+ictstm$sim_qalys(dr = c(0,.03))
+head(ictstm$qalys_)
+@
+
+<<ictstm_sim_costs>>=
+ictstm$sim_costs(dr = .03)
+head(ictstm$costs_)
+@
+
+<<ictstm_icer>>=
+ce_sim_ictstm <- ictstm$summarize()
+cea_pw_ictstm <- cea_pw(ce_sim_ictstm, comparator = 1,
+                        dr_qalys = .03, dr_costs = .03,
+                        k = seq(0, 25000, 500))
+icer_tbl(cea_pw_ictstm, colnames = strategies$strategy_name,
+         k = 100000)
+@
+
+%% -- Partitioned survival model -----------------------------------------------
+\subsection{Partitioned survival model}
+<<psm_hesim_data>>=
+hesim_dat$patients <- data.table(
+  patient_id = 1:4,
+  patient_wt = rep(1/4, 4),
+  age = c(55, 65, 55, 65),
+  female = c(1, 1, 0, 0)
+)
+@
+
+\subsubsection{Parameterization}
+<<psm_data>>=
+surv_est_data <- as_pfs_os(onc3, patient_vars = c("patient_id", "female", "age", 
+                                                  "strategy_name"))
+surv_est_data[patient_id %in% c(1, 2)]
+@
+
+<<psm_fits>>=
+fit_pfs_wei <- flexsurv::flexsurvreg(
+  Surv(pfs_time, pfs_status) ~ strategy_name + female,
+  data = surv_est_data,
+  dist = "weibull")
+
+fit_os_wei <- flexsurvreg(
+  Surv(os_time, os_status) ~ strategy_name + female,
+  data = surv_est_data,
+  dist = "weibull")
+
+psmfit_wei <- flexsurvreg_list(fit_pfs_wei, fit_os_wei)
+@
+
+<<psm_stateval_tbl>>=
+attr(utility_tbl, "patients") <- hesim_dat$patients
+attr(medcost_tbl, "patients") <- hesim_dat$patients
+@
+
+<<psm_drugcost>>=
+drugcost_tbl <- stateval_tbl(
+  data.table(strategy_id = strategies$strategy_id,
+             est = c(2000, 12000, 15000)),
+  dist = "fixed",
+  hesim_data = hesim_dat)
+print(drugcost_tbl)
+@
+
+\subsubsection{Simulation}
+
+<<psm_survmods_data>>=
+survmods_data <- expand(hesim_dat, by = c("strategies", "patients"))
+head(survmods_data)
+@
+
+<<psm_survmods>>=
+survmods <- create_PsmCurves(psmfit_wei,
+                             input_data = survmods_data,
+                             n = n_samples,
+                             bootstrap = FALSE, est_data = surv_est_data)
+@
+
+<<psm_stvalmods>>=
+utilitymod <- create_StateVals(utility_tbl, n = n_samples)
+drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples)
+medcostmod <- create_StateVals(medcost_tbl, n = n_samples)
+costmods <- list(Drug = drugcostmod, Medical = medcostmod)
+@
+
+<<psm_econmod>>=
+psm <- Psm$new(survival_models = survmods,
+               utility_model = utilitymod,
+               cost_models = costmods)
+@
+
+<<psm_sim_survival>>=
+times <- seq(0, 20, by = .1)
+psm$sim_survival(t = times)
+head(psm$survival_)
+@
+
+<<psm_plot_survival, echo = FALSE, include = FALSE>>=
+# Survival curve plotting data
+surv_means <- psm$survival_[, lapply(.SD, weighted.mean, w = patient_wt),
+                                by = c("strategy_id", "curve", "t"),
+                               .SDcols = "survival"]
+surv_mean_min <- surv_means[curve != 2][, curve := curve + 1]
+setnames(surv_mean_min, "survival", "survival_min")
+surv_means <- merge(surv_means, surv_mean_min,
+                    by = c("strategy_id", "curve", "t"),
+                    all.x = TRUE)
+surv_means[, survival_min := ifelse(curve == 1, 0, survival_min)]
+surv_means[, curve := factor(curve, labels = c("PFS", "OS"))]
+surv_means[, strategy_name := factor(strategy_id, labels = c("SOC", "New 1", "New 2"))]
+
+# Plot
+ggplot(surv_means, aes(x = t, y = survival, fill = curve)) +
+  geom_line(aes(col = curve)) +
+  facet_wrap(~strategy_name) +
+  geom_ribbon(aes(ymin = survival_min, ymax = survival), alpha = .5) +
+  xlab("Years") + ylab("Proportion surviving") +
+  scale_color_discrete(name = "Survival curve") +
+  guides(fill = FALSE) +
+  theme_bw() +
+  theme(legend.position = "bottom") 
+@
+
+\begin{figure}[t!]
+\centering
+\includegraphics{figure/psm_plot_survival-1.pdf}
+\caption{Mean progression free survival and overall survival, weighted by weight of each patient in the target population.
+Solid and dotted lines represent means and 95\% confidence intervals from the probabilistic sensitivity analysis, respectively.
+weighted} \label{fig:psm-survival}
+\end{figure}
+
+<<psm_sim_stateprobs>>=
+psm$sim_stateprobs()
+psm$stateprobs_[sample == 1 & patient_id == 1 & state_id == 2 & t == 12]
+@
+
+<<psm_costs_qalys>>=
+psm$sim_costs(dr = .03)
+psm$sim_qalys(dr = .03)
+@
+
+<<psm_icer>>=
+ce_sim_psm <- psm$summarize()
+cea_pw_psm <- cea_pw(ce_sim_psm, comparator = 1,
+                     dr_qalys = .03, dr_costs = .03,
+                     k = seq(0, 25000, 500))
+icer_tbl(cea_pw_psm, colnames = strategies$strategy_name,
+         k = 100000)
+@
+
+\subsection{Cohort disctete time state transition model}
+
+\subsubsection{Parameterization}
+
+<<cdtstm_fit_soc, warning = FALSE>>=
+library("msm")
+qinit <- matrix(0, nrow = 3, ncol = 3)
+qinit[1, 2] <-  0.28
+qinit[1, 3] <-  0.013
+qinit[2, 3] <-  0.10
+msm_fit <- msm(state_id ~ time, subject = patient_id, 
+               data = onc3p[strategy_name == "SOC"],
+               exacttimes = FALSE,
+               covariates = ~ age + female,
+               qmatrix = qinit, gen.inits = FALSE)
+@
+
+<<msm_exact, eval = FALSE>>=
+# msm with exact transition times
+tmat2 <- tmat; tmat2[is.na(tmat2)] <- 0
+msm_efit <- msm(state_id ~ time, subject = patient_id, 
+                data = onc3p[strategy_name == "SOC"],
+                qmatrix = tmat2, exacttimes = TRUE)
+
+# flexsurv with exponential model
+library("flexsurv")
+exp_fits <- vector(length = n_trans, mode = "list")
+for (i in 1:length(wei_fits)){
+  exp_fits[[i]] <- flexsurvreg(Surv(time, status) ~ 1, data = onc3, 
+                               subset = (transition_id == i & 
+                                         strategy_name == "SOC"),
+                               dist = "exponential")
+}
+
+# Compare
+msm_efit
+exp_fits[[1]]$res
+exp_fits[[2]]$res
+exp_fits[[3]]$res
+@
+
+<<cdtstm_params>>=
+params2 <- list(
+  lrr_12_est = c(soc = log(1), new1 = log(.80), new2 = log(.71)),
+  lrr_12_se = c(soc = 0, new1 = .03, new2 = .04),
+  lrr_13_est = c(soc = log(1), new1 = log(.90), new2 = log(.85)),
+  lrr_13_se = c(soc = 0, new1 = .02, new2 = .03),
+  u_mean = c(s1 = .9, s2 = .75, s3 = .05),
+  u_se = c(s1 = .02, s2 = .03, s3 = .05)
+)
+
+params2_def <- define_rng({
+  list(
+    rr_12 = lognormal_rng(lrr_12_est, lrr_12_se),
+    rr_13 = lognormal_rng(lrr_13_est, lrr_13_se),
+    u = beta_rng(mean = u_mean, sd = u_se)
+  )}, n = n_samples)
+
+params2_rng <- eval_rng(params2_def, params2)
+@
+
+<<qmat_soc>>=
+transmod_data <- survmods_data
+qmat_soc <- qmatrix(msm_fit, newdata = transmod_data[strategy_name == "SOC"],
+                    uncertainty = "normal", n = n_samples)
+qmat_soc[,, 1]
+@
+
+<<pmat_soc>>=
+cycle_len <- 1/4
+pmat_soc <- expmat(qmat_soc, t = cycle_len)
+pmat_soc[,, 1]
+@
+
+<<transprobs_id>>=
+tpmat_id <- tpmatrix_id(transmod_data, n_samples)
+head(tpmat_id)
+@
+
+<<xbeta>>=
+xbeta <- function(x, beta) c(x %*% t(beta))
+@
+
+<<rr>>=
+x_rr <- as.matrix(transmod_data[, .(soc, new1, new2)])
+rr <- cbind(xbeta(x_rr, params2_rng$rr_12),
+            xbeta(x_rr, params2_rng$rr_13))
+@
+
+<<apply_rr>>=
+pmat <- apply_rr(pmat_soc, rr = rr,
+                 index = list(c(1, 2), c(1, 3)))
+@
+
+<<tparams_transprobs>>=
+tprobs <- tparams_transprobs(pmat, tpmat_id)
+@
+
+\subsubsection{Simulation}
+<<dtstm_transmod>>=
+transmod <- CohortDtstmTrans$new(params = tprobs, cycle_length = cycle_len)
+@
+
+<<dtstm_econmod>>=
+cdtstm <- CohortDtstm$new(trans_model = transmod,
+                         utility_model = psm$utility_model,
+                         cost_models = psm$cost_models)
+@
+
+<<cdtstm_sim_stateprobs>>=
+cdtstm$sim_stateprobs(n_cycles = 30/cycle_len)
+@
+
+<<cdtstm_plot_stateprobs, echo = FALSE, include = FALSE>>=
+stprobs_tot <- cdtstm$stateprobs_[, lapply(.SD, weighted.mean, w = patient_wt),
+                                   by = c("sample", "strategy_id", "state_id", "t"),
+                                   .SDcols = "prob"]
+stprobs_summary <- summarize_stprobs(stprobs_tot)
+ggplot(stprobs_summary, aes(x = t, y = prob_mean)) +
+  geom_line(aes(col = strategy_name)) +
+  # geom_ribbon(aes(x = t, ymin = prob_lower, ymax = prob_upper,
+  #                 fill = strategy_name), alpha = .3) +
+  facet_wrap(~state_name) +
+  xlab("Year") + ylab("Probability") +
+  scale_fill_discrete("Strategy") + scale_color_discrete("Strategy")
+@
+
+<<cdtstm_sim_qalys_costs>>=
+cdtstm$sim_qalys(dr = .03)
+cdtstm$sim_costs(dr = .03)
+@
+
+<<cdtstm_icer>>=
+ce_sim_cdtstm <- cdtstm$summarize()
+cea_pw_cdtstm <- cea_pw(ce_sim_cdtstm, comparator = 1, 
+                        dr_costs = .03, dr_qalys = .03)
+icer_tbl(cea_pw_cdtstm, output = "data.table")
+@
 
@@
 \section{Cost-effectiveness analysis} \label{sec:cea}
+<<cea>>=
+wtp <- seq(0, 250000, 500) # Willingness to pay per QALY
+cea_ictstm <- cea(ce_sim_ictstm, dr_costs = .03, dr_qalys = .03, k = wtp)
+@
+
+<<ce_plane>>=
+library("scales")
+strategy_factor <- function (x) {
+  factor(x, levels = 1:3, labels = c("SOC", "New 1", "New 2"))
+}
+
+xlim <- ceiling(max(cea_pw_ictstm$delta[, ie]) * 1.2)
+ylim <- max(cea_pw_ictstm$delta[, ic]) * 1.2
+ggplot(cea_pw_ictstm$delta,
+       aes(x = ie, y = ic, col = strategy_factor(strategy_id))) +
+  geom_jitter(size = .5)  +
+  xlab("Incremental QALYs") +
+  ylab("Incremental cost") +
+  scale_y_continuous(limits = c(-ylim, ylim),
+                     labels = scales::dollar) +
+  scale_x_continuous(limits = c(-xlim, xlim), breaks = seq(-2, 2, .5)) +
+  theme(legend.position = "bottom") +
+  scale_colour_discrete(name = "Strategy") +
+  geom_abline(slope = 100000, linetype = "dashed") +
+  geom_hline(yintercept = 0) +
+  geom_vline(xintercept = 0)
+@
+
+<<ceac>>=
+ggplot(cea_ictstm$mce,
+       aes(x = k, y = prob, col = strategy_factor(strategy_id))) +
+  geom_line() +
+  xlab("Willingness to pay") +
+  ylab("Probability most cost-effective") +
+  scale_x_continuous(breaks = seq(0, max(wtp), length.out = 6),
+                     label = scales::dollar) +
+  theme(legend.position = "bottom")
+@
+
+<<evpi>>=
+ggplot(cea_ictstm$evpi, aes(x = k, y = evpi)) +
+  geom_line()  +
+  xlab("Willingness to pay") +
+  ylab("Expected value of perfect information") +
+  scale_x_continuous(breaks = seq(0, max(wtp), length.out = 6),
+                     label = scales::dollar) +
+  scale_y_continuous(label = scales::dollar) +
+  theme(legend.position = "bottom")
+@
 

### 2020-12-16 manuscript.Rnw
Some text updates

@@
 Multiple model types are supported including cohort discrete time state transition models (DTSTMs), N-state PSMs, and individual-level continuous time state transition models (CTSTMs), encompassing both Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
-To maximize flexibility and facilitate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R} or by inputting parameters obtained from external sources.
+To maximize flexibility and facilitate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R}, by inputting parameters obtained from external sources.
 So that individual-level simulation and PSA can be run quickly, \pkg{Rcpp} and \pkg{data.table} are heavily utilized.
@@
 An overview of the coding framework is provided in Section~\ref{sec:framework}.
-The next two sections contain example analyses.
-Section~\ref{sec:modeling-patient} shows how both partitioned survival and multi-state models can be used in an ideal case where rich patient level data is available.
-Section~\ref{sec:modeling-aggregate} then moves to the common scenario where only aggregate level data is available shows how both cohort and individual models can be employed.
-The cost-effectiveness framework is described in Section~\ref{sec:cea} along with worked examples.
+An illustrative example using a three state model of disease progression in oncology is provided in Section~\ref{sec:illustrative-example}.
+Analyses are conducted for each of the three supported model types, illustrating approaches suitable for both patient level and summary level data.
+The cost-effectiveness framework is described in Section~\ref{sec:cea} along with worked examples based on the output from the prior section.
 Section~\ref{sec:discussion} makes comparisons to other software and discusses possible extensions.
@@
 While tunnel states can be used to approximate a semi-Markov model using a cohort approach, they can only be simulated in a general fashion using individual level models.
-Discrete time individual simulation is possible, but we use a continuous time models (i.e., iCTSTMs) because they do not require specification of model cycles and can be run considerably faster.
+Discrete time individual simulation is possible, but we use continuous time models (i.e., iCTSTMs) because they do not require specification of model cycles and can be run considerably faster.
 
 PSMs are specialized models that can be parameterized using survival curves and are especially useful in oncology where PFS and OS are commonly reported.
-They are ``area under the curve'' models, although they can also be formulate as STMs by using the survival curves to construct transition probabilities.
+They are ``area under the curve'' models, although they can also be formulated as STMs by using the survival curves to construct transition probabilities.
 
@@
 Economic models consist of a disease model, a utility model, and a set of cost models for each cost category.
-Model development proceed in 4 steps.
+Model development proceeds in 4 steps.
 The first step is to setup the model by defining the model structure, target population, and treatment strategies of interest, and storing the relevant information in a \code{hesim_data} object.
@@
 The former contain the underlying parameters of a statistical model while the latter contain parameters more immediate to prediction that have been transformed as function of the input data.
-The coefficients from a Weibull regression model are an example of a parameter object and the predicted shape and scale parameters of the Weibull distribution are an example of a transformed parameter object.
-
-Although economic models are ideally parameterized using explicit statistical models, it is not always feasible.
-Parameters are often taken from a variety of sources or from summary level data.
-In these cases, it can be convenient to build models with mathematical expressions usin non-standard evaluation.
-This can be done when building cDTSTMs using \code{defime_model()}.
+The regression coefficients of a logistic regression are example of a parameter objects while the predicted probabilities are examples of a transformed parameter object.
+
+Transformed parameter objects are often useful when parameterizing a model using external sources.
+Two common examples are predicted probabilities for transition matrices and predicted means for assigning values to health states.
+The advantage of a transformed parameter object is that an explicit statistical model does not need to be specified which affords more flexibility to the user.
 
@@
 \hline
-\multirow{2}{*}{\code{CohortDtstm}}& Custom & \code{tparams_transprobs} & \code{define_model()}\\
+\code{CohortDtstm} & Custom & \code{tparams_transprobs} & \code{msm::msm} \\
  & Multinomial logistic regressions & \code{params_mlogit} & \code{multinom_list}\\
-\multirow{2}{*}{\code{IndivCtstm}} & Multi-state model (joint likelihood) & \code{params_surv} & \code{flexsurv::flexsurvreg}\\
+\code{IndivCtstm} & Multi-state model (joint likelihood) & \code{params_surv} & \code{flexsurv::flexsurvreg}\\
 & Multi-state model (transition-specific) & \code{params_surv_list} & \code{flexsurvreg_list}\\
@@
 
-Although disease models are ideally parameterized by fitting a single statistical model, this is often infeasible.
-Parameters may be taken from multiple sources, frequently using summary level data.
-In these cases it can be more convenient to define transition probability matrices using mathematical expressions relating parameters and input data to transformed parameters.
-Transition probabilities for cDTSTMs can be constructed in this manner using \code{define_model()}.
+\code{tparams_transprobs} objects contain transition probability matrices that have been predicted for each PSA sample, treatment strategy, and patient from the target population.
+Transition probabilities may also vary by time interval to allow for time inhomogeneous Markov models.
+Transition probabilities can be predicted from a fitted multi-state model using the \pkg{msm} package or constructed ``by hand'' in a custom manner.
 
 \subsubsection{Costs and utility} 
-State values (i.e., costs and utilities) do not depend on the choice of disease model.
-They can currently either be modeled using a linear model or using predicted means.
-The latter is an example of a transformed parameter object since the predicted means are parameters that are presumably a function of the underlying parameters of a statistical model and input data.
+State values (i.e., costs and utilities) do not depend on the choice of disease model. They can currently either be modeled using a linear model or using predicted means.
 
@@
 \hline
-\multirow{2}{*}{Predicted means} & \code{tparams_mean} & \code{stateval_tbl}\\
- & \code{tparams_mean} & \code{define_model()}\\
+Predicted means & \code{tparams_mean} & \code{stateval_tbl::lm}\\
 Linear model & \code{params_lm} & \code{stats::lm}\\
@@
 
-The most straightforward way to construct state values is with a \code{stateval_tbl}, which is a special object used to assign values to health states that can vary across PSA samples, treatment strategies, patients, and/or time intervals.
+The most straightforward way to construct state values is with a \code{stateval_tbl}, which is a special object used to assign values (i.e.
+predicted means) to health states that can vary across PSA samples, treatment strategies, patients, and/or time intervals.
 State values can be specified either as moments (i.e., mean and standard error) or parameters (e.g., shape and scale of gamma distribution) of a probability distribution, or by pre-simulating values from a suitable probability distribution (e.g., from a Bayesian model).
-Like transition probabilities, state value can also be constructed with \code{define_model}.
 
@@
 \section{Illustrative example} \label{sec:illustrative-example}
+We consider a three state model commonly used describe disease progression in oncology.
+As shown in Figure~\ref{fig:three-state-model}, the three health states are stable disease (i.e.
+not progressed), progressed, and death.
+Patients can transition to a more severe health state (stable $\rightarrow$ progressed, stable $\rightarrow$ death, and progression $\rightarrow$ death) but cannot recover to a less severe health state.
 
@@
 \end{figure}
+
+This model can be simulated using either of the model types described in Section~\ref{sec:model-taxonomy}. We demonstrate use of all three below, starting with the iCTSTM.
+
+%% -- Individual continuous time state transition model ------------------------
+\subsection{Individual continuous time state transition model}
+We setup the model by defining the model structure, target population, and treatment strategies of interest.
+The transitions are characterized by a matrix where each element denotes a transition from a row to a column.
+If a transition is not possible it is marked with \code{NA}; otherwise, an integer denotes the transition number.
 
@@
 @
+
+The target population, treatment strategies, and (non-death) health states are stored in a \code{hesim_data} object.
+There are 3 treatment strategies: standard of care (SOC) and two new interventions (New 1 and New 2).
+The target population consists of a heterogeneous population of 1000 patients, which is large enough to ensure that averages across patients in the individual simulation are reasonably stable.
 
@@
 )
-n_states <- nrow(states)
 
@@
 )
-n_strategies <- nrow(strategies)
 
@@
 
-%% -- Individual continuous time state transition model ------------------------
-\subsection{Individual continuous time state transition model}
-
 \subsubsection{Parameterization}
+In a STM, the disease model governs transitions between health states.
+For now, we assume that that patient level data with exact transition times is available so that a multi-state statistical model can be used to parameterize the disease model.
+
+Multi-state data consists of one row for each possible transition from a given health state where only one transition is observed and all others are censored.
+We use the simulated dataset \code{onc3} from the \pkg{hesim} package.
 
@@
 
-<<ictstm_fits>>=
+Multi-state models can be fit in one of two ways.
+First, a joint survival model with interaction terms for different transition can be estimated, which is useful if there are constraints in the parameters across
+transitions, such as coefficients that are assumed equal across transitions.
+Here, we will take a second approach that is computationally more efficient and fit a separate model for each transition.
+We illustrate with parametric Weibull models but multiple distributions should be compared in a real application \citep{williams2017cost}.
+The treatment effect is assumed to only influence transitions from stable disease since patients are assumed to switch to a second line treatment after progression.
+
+<<ictstm_fits, warning = FALSE>>=
 library("flexsurv")
@@
 n_times <- 2
+n_states <- nrow(states)
+n_strategies <- nrow(strategies)
 n_i <- n_states * n_times
@@
 
-\subsection{Cohort disctete time state transition model}
+\subsection{Cohort disctete time state transition model} \label{sec:cdtstm-example}
 

### 2020-12-17 manuscript.Rnw
First draft of iCTSTM example

@@
   engine="R", tidy = FALSE, eval = TRUE, cache = TRUE,
-  fig.height = 5, fig.width = 7
+  fig.height = 4, fig.width = 6
 )
@@
 Patients can transition to a more severe health state (stable $\rightarrow$ progressed, stable $\rightarrow$ death, and progression $\rightarrow$ death) but cannot recover to a less severe health state.
+We assume that patients remain on first line (1L) treatment until progression, at which time they switch to a second line (2L) treatment.
 
@@
 We illustrate with parametric Weibull models but multiple distributions should be compared in a real application \citep{williams2017cost}.
-The treatment effect is assumed to only influence transitions from stable disease since patients are assumed to switch to a second line treatment after progression.
+The treatment effect is assumed to only influence transitions from stable disease since patients are assumed to switch to a 2L treatment after progression.
 
@@
 
+Utility and cost values are stored in \code{stateval_tbl} objects.
+Utility is assumed to equal 0.8 with stable disease (standard error (SE) = 0.02) and 0.6 (SE = 0.05) with progressed disease.
+Utility values are assumed to follow a beta distribution, which can be parameterized either using its shape parameters or via the mean and standard error (in which case the method of moments is used to derive the shape parameters).
+
 <<ictstm_utility>>=
@@
 
+Both medical and drug costs are considered.
+Medical costs are assumed to follow a gamma distribution which is often appropriate as they tend to be right skewed.
+Like utility, the distribution is characterized by the mean (\$2,000 with stable disease, \$9,500 with progressed disease) and standard error (assumed equal to the mean) and the method of moments is used to derive the underlying shape and scale parameters.
+
 <<ictstm_medcosts>>=
@@
 
+Drug costs are assumed to be fixed (i.e., measured without uncertainty).
+Costs are \$2,000 with SOC, \$12,000 with New 1, and \$15,000 with New 2 when on 1L treatment.
+All patients are assumed to switch to the same treatment at 2L after progression, which costs \$1,500 for the first 3 months and \$1,200 thereafter.
+
 <<ictstm_drugcost_setup, echo=FALSE, results="hide">>=
@@
     time_start = rep(c(0, 3/12), n_states * n_strategies),
-    est = c(rep(2000, n_i -1), 1500, 
-            rep(12000, n_i -1), 1000,
-            rep(15000, n_i - 1), 1200
-            ) 
+    est = c(2000, 2000, 1500, 1200,
+           12000, 12000, 1500, 1200,
+           15000, 15000, 1500, 1200)
 ) 
@@
 )
+print(drugcost_tbl)
 @
@@
 \subsubsection{Simulation}
+Before constructing the economic model, we specify the number of parameters samples that will be drawn for each submodel for the PSA.
+
 <<ictstm_n_samples>>=
@@
 @
+
+The input data for the transition model consists of one row for each treatment strategy and patient and can be easily generated from a \code{hesim_data} object with the \code{expand} function.
 
@@
 @
+
+The transition model is then constructed as a function of the parameters (from the fitted Weibull models) and the input data.
+We must also specify the allowed transitions, the desired number of PSA samples (drawn from the multivariate normal distribution of the maximum likelihood estimate of the regression coefficients), the "clock" (recall that we used a clock reset approach when fitting the Weibull models), and the starting age of the patients.
+The starting age of the patients is useful because it allows us to specify a maximum age for patients during the simulation, ensuring that the simulated patients do not live unrealistically long lives.
 
@@
 @
+
+The utility and cost models can be easily constructed from the \code{stateval_tbl} objects with the \code{create_StateVals()} function, in which the (transformed) parameters are \code{tparams_mean} objects.
+The \code{hesim_data} object provides the information required to ensures that state values are constant within groups not specified within a particular \code{stateval_tbl}; for instance, since utility only varies by health state, it is assumed constant across patients, treatment strategies, and time.
+The models for each cost category are combined in a list.
 
@@
 
+The complete economic model is an constructed by combining the transition, utility, and cost submodels.
+
 <<ictstm_econmod>>=
@@
 
+Once the economic model has been constructed, it is straightforward to simulate outcomes.
+A trajectory through the multi-state model is simulated with the \code{$sim_disease()} method with patients assumed to live to a maximum age of 100.
+
 <<ictstm_sim_disease>>=
@@
 @
+
+State probabilities (averaged across patients in a given subgroup) can be constructed from the disease progression simulation output with the \code{$sim_stateprobs()} method.
+As shown in Figure~\ref{fig:ictstm-stprobs}, patients using the newer treatments remain in stable disease longer and have longer expected survival.
 
@@
 
+QALYs and costs are simulated from the disease progression simulation output using the utility and cost submodels with the \code{$sim_qalys()} and \code{$sim_costs()} methods.
+Outcomes are simulated for each discount rate specified with the \code{dr} argument.
+By default, \code{$sim_qalys()} will return life-years (i.e., a utility value of 1 for each year of life) in addition to QALYs.
+
 <<ictstm_sim_qalys>>=
@@
 @
+
+For a given subgroup, QALYs and costs are simulated for each PSA sample, treatment strategy, and health state.
+They are summarized (i.e., averaged) across PSA samples and treatment strategies (and subgroups) with the \code{$summarize()} method.
+
+<<ictstm_summarize>>=
+ce_sim_ictstm <- ictstm$summarize()
+head(ce_sim_ictstm$costs)
+@
+
+The output is a generic cost-effectiveness object (of class \code{ce}) that can be used to perform CEA.
+We demonstrate a formal CEA in Section~\ref{sec:cea}.
+Here, we make a simple pairwise comparisons between each of the new interventions (New 2 and New 1) and the comparator (SOC), and compute the incremental cost-effectiveness ratio (ICER).
 

### 2020-12-19 manuscript.Rnw
Rough draft of partitioned survival model illustration

@@
 An illustrative example using a three state model of disease progression in oncology is provided in Section~\ref{sec:illustrative-example}.
-Analyses are conducted for each of the three supported model types, illustrating approaches suitable for both patient level and summary level data.
+Analyses are conducted for each of the three supported model types, illustrating approaches suitable for both patient level and aggregate level data.
 The cost-effectiveness framework is described in Section~\ref{sec:cea} along with worked examples based on the output from the prior section.
@@
 Finally, Section~\ref{sec:conclusion} concludes.
+
+Analyses were performed using \pkg{hesim} version 0.5.0.
 
@@
 
-<<ictstm_fits, warning = FALSE>>=
+<<ictstm_fits, warning = FALSE, message = FALSE>>=
 library("flexsurv")
@@
 \subsection{Partitioned survival model}
+In the previous section multi-state models were fit to individual patient data.
+Partitioned survival analysis is an alternative approach that is frequently used for economic evaluations of therapies in oncology.
+Although there is increasing evidence that PSM are more biased than multi-state models \citep{woods2020partitioned}, PSMs are easier to parameterize with aggregate level data.
+Techniques for fitting multi-state models with aggregate data require more constraints and are less established \citep{price2011parameterization, pahuta2019technique, jansen2020multi}.
+PSMs, in contrast, consist of separate independent survival models that can be estimated with aggregate level data by reconstructing individual patient data from digitized survival curves \citep{guyot2012enhanced}.
+
+Since a PSM is a cohort model, the \code{hesim_data} object must be modified accordingly.
+We now specify 4 representative patients that comprise the cohort, each given an equal weight.
+We consider separate subgroups (identified with \code{grp_id}) for males and females.
+
 <<psm_hesim_data>>=
@@
   patient_id = 1:4,
+  grp_id = c(1, 1, 0, 0),
   patient_wt = rep(1/4, 4),
@@
 \subsubsection{Parameterization}
+A multi-state dataset can be converted into dataset with one row per patients and outcomes for PFS and OS using the function \code{as_pfs_os()}.
+
 <<psm_data>>=
-surv_est_data <- as_pfs_os(onc3, patient_vars = c("patient_id", "female", "age", 
+onc_pfs_os <- as_pfs_os(onc3, patient_vars = c("patient_id", "female", "age", 
                                                "strategy_name"))
-surv_est_data[patient_id %in% c(1, 2)]
-@
-
-<<psm_fits>>=
+onc_pfs_os[patient_id %in% c(1, 2)]
+@
+
+Survival models must be fit for both PFS and OS.
+To imitate a realistic scenario, we assume that patient level data is only available for SOC and that relative treatment effects are estimated with aggregate data (e.g., using a NMA).
+As such, we start by fitting Weibull models for the SOC.
+
+<<psm_fits_soc>>=
+fit_pfs_soc_wei <- flexsurv::flexsurvreg(
+  Surv(pfs_time, pfs_status) ~ female,
+  data = onc_pfs_os, subset = (strategy_name == "SOC"),
+  dist = "weibull")
+
+fit_os_soc_wei <- flexsurvreg(
+  Surv(os_time, os_status) ~  female,
+  data = onc_pfs_os, subset = (strategy_name == "SOC"),
+  dist = "weibull")
+
+psmfit_soc_wei <- flexsurvreg_list(pfs = fit_pfs_soc_wei, os = fit_os_soc_wei)
+@
+
+The relative treatment effects are available in terms of regression coefficients for the scale parameter on the log scale.
+The matrices \code{coef_pfs_wei} and \code{coef_os_wei} contain samples of the treatment effects for the PSA.
+We display the first 6 rows of the PFS coefficients.
+
+<<psm_fits, include = FALSE>>=
+# Create dummies
+onc_pfs_os[, new1 := ifelse(strategy_name == "New 1", 1, 0)]
+onc_pfs_os[, new2 := ifelse(strategy_name == "New 2", 1, 0)]
+
+# Fit models
 fit_pfs_wei <- flexsurv::flexsurvreg(
-  Surv(pfs_time, pfs_status) ~ strategy_name + female,
-  data = surv_est_data,
+  Surv(pfs_time, pfs_status) ~ female + new1 + new2,
+  data = onc_pfs_os,,
   dist = "weibull")
@@
 fit_os_wei <- flexsurvreg(
-  Surv(os_time, os_status) ~ strategy_name + female,
-  data = surv_est_data,
+  Surv(os_time, os_status) ~ female + new1 + new2,
+  data = onc_pfs_os,
   dist = "weibull")
 
-psmfit_wei <- flexsurvreg_list(fit_pfs_wei, fit_os_wei)
-@
+# Extract relative treatment effects
+params_pfs_wei <- create_params(fit_pfs_wei, n = n_samples)
+params_os_wei <- create_params(fit_os_wei, n = n_samples)
+coef_pfs_wei <- params_pfs_wei$coefs$scale[, c("new1", "new2")]
+coef_os_wei <- params_os_wei$coefs$scale[, c("new1", "new2")]
+@
+
+<<psm_coefs>>=
+head(coef_pfs_wei)
+@
+
+The parameters for utility and medical costs remain the same as in the previous section.
+We do, however, modify drug costs.
+Since we are no longer using an individual level model, costs cannot vary as a function of time in a health state.
+We now assume conservatively that drug costs remain at initial level when in the progression state.
 
@@
 drugcost_tbl <- stateval_tbl(
-  data.table(strategy_id = strategies$strategy_id,
-             est = c(2000, 12000, 15000)),
+  drugcost_dt[time_start ==  0][, time_start := NULL],
   dist = "fixed",
@@
 \subsubsection{Simulation}
+As with with the multi-state model, input data for the survival models is constructed with the \code{expand()} function.
 
@@
 
+Since we did not fit a survival regression model containing both the absolute and relative treatment effects, we must construct the survival parameter object manually.
+We obtain the parameters of the fitted Weibull regression models with the generic function \code{create_params()}, and then manually add the relative treatment effects to the coefficients for the scale parameters.
+(Note that if we had not fit a model for SOC, a survival parameter objects could have still been constructed with \code{params_surv()}.)
+
+<<psm_survmods_params>>=
+survmods_params <- create_params(psmfit_soc_wei, n = n_samples)
+survmods_params$pfs$coefs$scale <- cbind(survmods_params$pfs$coefs$scale,
+                                         coef_pfs_wei)
+survmods_params$os$coefs$scale <- cbind(survmods_params$os$coefs$scale,
+                                        coef_os_wei)
+@
+
+Survival curves for both PFS and OS are predicted using the \code{PsmCurves} class, which is setup using the input data and the parameters of the Weibull model.
+However, since we manually created the parameter object---meaning that the parameters were not generating using a formula interface---the input data must contain each term contained in the parameter object.
+We must consequently create an intercept column of ones.
+
 <<psm_survmods>>=
-survmods <- create_PsmCurves(psmfit_wei,
-                             input_data = survmods_data,
-                             n = n_samples,
-                             bootstrap = FALSE, est_data = surv_est_data)
-@
+# To remove with update of package
+colnames(survmods_params$pfs$coefs$shape)[1] <- "intercept"
+colnames(survmods_params$pfs$coefs$scale)[1] <- "intercept"
+colnames(survmods_params$os$coefs$shape)[1] <- "intercept"
+colnames(survmods_params$os$coefs$scale)[1] <- "intercept"
+
+# To keep
+survmods_data[, intercept := 1]
+survmods <- create_PsmCurves(survmods_params, 
+                             input_data = survmods_data)
+@
+
+We re-instantiate all of the utility and cost models.
+This is necessary because our patient population changed and we are now only making predictions for 4 representative patients (rather than 1,000 in the individual simulation).
+The \code{hesim_data} argument to \code{create_StateVals()} does this for us automatically.
 
@@
 
+Finally, the complete partitioned survial model is constructed using the survival, utility, and cost mdoels.
+
 <<psm_econmod>>=
@@
 @
+
+We first simulate survival curves by month for 20 years.
+State probabilities are easily computed from these survival curves with the \code{sim_stateprobs()} method.
+The area under the PFS curve is the proportion of patients in stable disease while the area between the PFS and OS curves is the proportion of patient whose disease has progressed.
 
@@
 
+QALYs and costs are simulated in using the same \code{sim_qalys()} and \code{sim_costs()} methods as in the previous section, as is the calculation of the ICER.
+
 <<psm_costs_qalys>>=
+psm$sim_qalys(dr = .03)
 psm$sim_costs(dr = .03)
-psm$sim_qalys(dr = .03)
 @
@@
 cea_pw_psm <- cea_pw(ce_sim_psm, comparator = 1,
-                     dr_qalys = .03, dr_costs = .03,
-                     k = seq(0, 25000, 500))
+                     dr_qalys = .03, dr_costs = .03)
 icer_tbl(cea_pw_psm, colnames = strategies$strategy_name,
@@
 
-<<msm_exact, eval = FALSE>>=
+<<msm_exact, eval = FALSE, message = FALSE>>=
 # msm with exact transition times
@@
 # flexsurv with exponential model
-library("flexsurv")
 exp_fits <- vector(length = n_trans, mode = "list")

### 2020-12-20 manuscript.Rnw
Complete draft of illustrative example

@@
 %% -- Individual continuous time state transition model ------------------------
-\subsection{Individual continuous time state transition model}
+\subsection{Individual continuous time state transition model} \label{sec:ictstm-example}
 We setup the model by defining the model structure, target population, and treatment strategies of interest.
@@
 %% -- Partitioned survival model -----------------------------------------------
-\subsection{Partitioned survival model}
+\subsection{Partitioned survival model} \label{sec:psm-example}
 In the previous section multi-state models were fit to individual patient data.
@@
 
-Finally, the complete partitioned survial model is constructed using the survival, utility, and cost mdoels.
+Finally, the complete partitioned survial model is constructed using the survival, utility, and cost models
 
@@
 
+%% -- Cohort discrete time state transition model ------------------------------
 \subsection{Cohort disctete time state transition model} \label{sec:cdtstm-example}
+The third model type supported by \pkg{hesim} is the cDTSTM.
+Although it is commonly used in health economics, the approaches taken in Section~\ref{sec:ictstm-example} and \ref{sec:psm-example} make better use of multi-state data if the time to event outcomes are continuously observed.
+However, in some cases panel data is available where observations are only collected at arbitrary times.
+This may, for instance, occur in routine practice oncology data when progression of disease is not continuously evaluated.
+To facilitate analysis of this type of data, we have converted the \code{onc3} dataset from above into panel form.
+
+<<cdtstm_data, warning = FALSE>>=
+head(onc3p)
+@
 
 \subsubsection{Parameterization}
-
-<<cdtstm_fit_soc, warning = FALSE>>=
+As in Section~\ref{sec:psm-example}, we consider a case where patient level data is available for SOC but aggregate level data is used to estimate the relative treatment effects.
+When exact transition times are not observed, the survival models estimated in Section~\ref{sec:ictstm-example} are inappropriate.
+The \pkg{msm} package, can, however, be used to estimate models from such data.
+The loss of information does come at a price as only exponential and piecewise exponential models are possible.
+We illustrate by fitting an exponential multi-state model for SOC using \pkg{msm}.
+(Note that if exact times were observed and we set the argument \code{exacttimes = TRUE}, then the parameter estimates would be the same as an exponential model fit using \code{flexsurvreg()}.)
+
+<<cdtstm_fit_soc, warning = FALSE, message = FALSE>>=
 library("msm")
 qinit <- matrix(0, nrow = 3, ncol = 3)
-qinit[1, 2] <-  0.28
-qinit[1, 3] <-  0.013
-qinit[2, 3] <-  0.10
+qinit[1, 2] <-  0.28; qinit[1, 3] <-  0.013; qinit[2, 3] <-  0.10
 msm_fit <- msm(state_id ~ time, subject = patient_id, 
@@
 
-<<msm_exact, eval = FALSE, message = FALSE>>=
+<<msm_exact, include = FALSE, message = FALSE>>=
 # msm with exact transition times
@@
 
+We assume that the relative treatment effects were estimated as relative risks.
+Either a posterior distribution from a Bayesian analysis or parameters of a parametric distribution can be used.
+We use the latter approach here and assume that the relative risk is approximately normally distributed on the log scale.
+\pkg{hesim} provides the \code{define_rng()} and \code{eval_rng()} functions for defining and evaluating random number generators for a PSA.
+While standard functions in \proglang{R} could be used to draw random deviates from probability distributions instead, these functions can be more convenient for health economic modeling.
+Notable advantages include only needing to specify the number of random samples once, being able to pass moments of distributions as arguments (as in \code{stateval_tbl()}), and guaranteeing a common output (either a vector or a two-dimensional tabular object, with dimensions determined by the number or random samples).
+
 <<cdtstm_params>>=
-params2 <- list(
+params_rr <- list(
   lrr_12_est = c(soc = log(1), new1 = log(.80), new2 = log(.71)),
@@
   lrr_13_est = c(soc = log(1), new1 = log(.90), new2 = log(.85)),
-  lrr_13_se = c(soc = 0, new1 = .02, new2 = .03),
-  u_mean = c(s1 = .9, s2 = .75, s3 = .05),
-  u_se = c(s1 = .02, s2 = .03, s3 = .05)
+  lrr_13_se = c(soc = 0, new1 = .02, new2 = .03)
 )
 
-params2_def <- define_rng({
+params_rr_def <- define_rng({
   list(
     rr_12 = lognormal_rng(lrr_12_est, lrr_12_se),
-    rr_13 = lognormal_rng(lrr_13_est, lrr_13_se),
-    u = beta_rng(mean = u_mean, sd = u_se)
+    rr_13 = lognormal_rng(lrr_13_est, lrr_13_se)
   )}, n = n_samples)
-
-params2_rng <- eval_rng(params2_def, params2)
-@
+params_rr_rng <- eval_rng(params_rr_def, params_rr)
+@
+
+In the prior two sections, we constructed the disease model using parameter (i.e., \code{params_surv}) objects.
+Here, we will use a transformed parameter (i.e., \code{tparams_transprobs}) object, which is a very flexible way to construct the transition probability matrices for a cDTSTM.
+A \code{tparams_transprobs} object requires a 3-dimensional array of transition matrices, sorted by PSA sample, treatment strategy, patient, and optionally (in a time-inhomogeneous model) time interval.
+
+We start by predicting transition intensity matrices for SOC from the exponential multi-state model using the same input data used for the survival models in the PSM example.
+A prediction is made for each representative patient and samples of the transition intensity matrices are drawn for the PSA using the asymptotic normal approximation.
+The array contains $B \times N_p$ matrices where $B$ is the number of PSA samples and $N_p$ is the number of patients.
 
@@
                     uncertainty = "normal", n = n_samples)
+dim(qmat_soc)
 qmat_soc[,, 1]
 @
+
+Transition probability matrices of arbitrary duration are derived from the transition intensity matrices with the matrix exponential. We use a cycle length of 3 months.
 
@@
 
+Identifiers for each matrix in the array can then be generated with \code{tpmatrix_id()}.
+
 <<transprobs_id>>=
@@
 
+Relative risks can be predicted for each row of the input data and each sample of the coefficients for the PSA using matrix multiplication.
+In particular, let $x$ be a $N \times k$ matrix where $k$ is the number of predictors and $\beta$ be a $B \times k$ matrix where each row contains a given sample for the PSA.
+The function \code{xbeta} produces a $N \times B$ matrix of predicted relative risks and then ``flattens'' the matrix to create a single vector of length $B \times N$.
+
 <<xbeta>>=
@@
 
+We use this function to predict distributions of relative risks (for the transitions stable $\rightarrow$ progression and stable $\rightarrow$ death) as a function of the three treatment strategies.
+
 <<rr>>=
 x_rr <- as.matrix(transmod_data[, .(soc, new1, new2)])
-rr <- cbind(xbeta(x_rr, params2_rng$rr_12),
-            xbeta(x_rr, params2_rng$rr_13))
-@
+rr <- cbind(xbeta(x_rr, params_rr_rng$rr_12),
+            xbeta(x_rr, params_rr_rng$rr_13))
+@
+
+Relative risks are applied to the elements (stable $\rightarrow$ progression (1, 2) and stable $\rightarrow$ death (1, 3)) of each transition probability matrix for SOC with \code{apply_rr()}.
+The result is an array of matrices for each PSA sample, treatment strategy,  and representative patient, as desired.
+The \code{tparams_transprobs} object can then be created by simply combining the transition probability matrices and the identification dataset.
 
@@
                  index = list(c(1, 2), c(1, 3)))
-@
-
-<<tparams_transprobs>>=
 tprobs <- tparams_transprobs(pmat, tpmat_id)
@@
 \subsubsection{Simulation}
+One the model has been parameterized, the simulation is straightforward. The transition model depends only on the transition probability matrices and the cycle length,
+
 <<dtstm_transmod>>=
@@
 @
+
+and the complete economic model is again instantiated from the transition, utility, and cost submodels.
 
@@
 @
+
+We simulate state probabilities for 120 cycles (i.e., 30 years).
 
@@
 
+QALYs and costs are again simulated using a 3 percent discount rate
+
 <<cdtstm_sim_qalys_costs>>=
@@
 @
+
+and an ICER is computed.
 
@@
                         dr_costs = .03, dr_qalys = .03)
-icer_tbl(cea_pw_cdtstm, output = "data.table")
+icer_tbl(cea_pw_cdtstm, output = "data.table", k = 100000)
 @

### 2020-12-21 manuscript.Rnw
Draft of CEA function

@@
 \captionsetup[table]{skip=5pt}
+\usepackage{amsmath}
+\DeclareMathOperator*{\argmax}{arg\,max}
 
@@
 \section{Cost-effectiveness analysis} \label{sec:cea}
+CEA is based on the net monetary benefit (NMB).
+For a given parameter set $\theta$, the NMB with treatment $j$ is computed as the difference between the monetized health gains from an intervention less costs, or,
+
+\begin{equation}
+NMB(j,\theta) = e_{j}(\theta)\cdot k- c_{j}(\theta),
+\end{equation}
+
+where $e_{j}$ and $c_{j}$ are measures of health outcomes (e.g.
+QALYs) and costs using treatment $j$ respectively, and $k$ is a decision makers willingness to pay (WTP) per unit of health outcomes.
+The optimal treatment is the one that maximizes the expected NMB.
+
+\begin{equation}
+j^{*} = \argmax_j E_{\theta} \left[NMB(j,\theta)\right].
+\end{equation}
+
+For a pairwise comparison, treatment $1$ is preferred to treatment $0$ if the expected incremental net monetary benefit (INMB) is positive; that is, if $E_\theta \left[INMB\right] > 0$ where the INMB is given by
+
+\begin{equation}
+INMB(\theta) = NMB(j = 1, \theta) - NMB(j = 0, \theta).
+\end{equation}
+
+Treatments can be compared in an equivalent manner using the incremental cost-effectiveness ratio (ICER).
+The most common case occurs when a new treatment is more effective and more costly so that treatment $1$ is preferred to treatment $0$ if the ICER is greater than the willingness to pay threshold $k$,
+
+\begin{equation}
+k > \frac{E_\theta[c_{1} - c_{0}]}{E_\theta[e_{1} - e_{0}]} = ICER.
+\end{equation}
+
+There are three additional cases.
+Treatment $1$ is considered to \emph{dominate} treatment $0$ if it is more effective and less costly.
+Treatment $1$ is \emph{dominated} by treatment $0$ if it is less effective and more costly.
+Finally, treatment $1$ is preferred to treatment $0$ if it is less effective and less costly when $k < ICER$.
+
+\subsection{Example}
+Uncertainty over $\theta$ is captured using the PSA.
+The simulated distribution of QALYs and costs for each treatment strategy can be produced with the \code{$summarize()} method.
+Such output was generated above for the iCTSTM.
+
+<<ce>>=
+ce_sim_ictstm
+@
+
+A CEA is performed from this output with the \code{cea()} and \code{cea_pw()} functions.
+\code{cea()} summarizes results by simultaneously accounting for each treatment strategy in the analysis, while \code{cea_pw()} (as used above) summarizes “pairwise” results in which each treatment is compared to a comparator.
+
 <<cea>>=
@@
 cea_ictstm <- cea(ce_sim_ictstm, dr_costs = .03, dr_qalys = .03, k = wtp)
-@
-
-<<ce_plane>>=
+cea_pw_ictstm <- cea_pw(ce_sim_ictstm, comparator = 1,
+                        dr_qalys = .03, dr_costs = .03,
+                        k = wtp)
+@
+
+A number of common measures can be used to summarize a CEA and represent decision uncertainty.
+\pkg{hesim} provides support for a number of them including the cost-effectiveness plane, cost-effectiveness acceptability curve (CEAC), cost-effectiveness acceptability frontier (CEAF), and expected value of perfect information (EVPI).
+
+\subsubsection{Cost-effectiveness plane}
+The cost-effectiveness plane plots the incremental effectiveness of a treatment strategy (relative to a comparator) against the incremental cost of the treatment strategy.
+The plot is useful because it demonstrates both the uncertainty and the magnitude of the estimates.
+Each point on the plot is from a particular draw from the PSA.
+Data for plotting a cost-effectiveness plane comes from the \code{delta} output generated from \code{cea_pw()}.
+
+<<ce_plane, fig.height = 4, fig.width = 6>>=
 library("scales")
@@
 
-<<ceac>>=
+\subsubsection{Cost-effectiveness acceptability curve}
+A useful summary measure for quantifying uncertainty is the CEAC, which can generated based on simultaneous comparisons of all treatment strategies or by comparing each treatment strategy to a single comparator.
+The former is the probability that each strategy is the most cost-effective and is available from the \code{mce} element produced by \code{cea()}.
+
+<<ceac_simultaneous>>=
 ggplot(cea_ictstm$mce,
@@
                      label = scales::dollar) +
+  scale_colour_discrete(name = "Strategy") +
   theme(legend.position = "bottom")
 @
+
+The relevant information for the latter is available from the \code{ceac} element produced by \code{cea_pw()}, and is the probability that each treatment is more cost-effective than the comparator.
+
+<<ceac_pw>>=
+ggplot(cea_pw_ictstm$ceac,
+        aes(x = k, y = prob, col = strategy_factor(strategy_id))) +
+  geom_line() +
+  xlab("Willingness to pay") +
+  ylab("Probability most cost-effective") +
+  scale_x_continuous(breaks = seq(0, max(wtp), length.out = 6),
+                     label = scales::dollar) +
+  theme(legend.position = "bottom") +
+  scale_colour_discrete(name = "Strategy")
+@
+
+\subsubsection{Cost-effectiveness acceptability frontier}
+One drawback of the CEAC is that the probability of being cost-effective cannot be used to determine the optimal treatment option.
+Instead, if the objective is to maximize health gain, then decisions should be based on the expected NMB.
+The CEAF, which plots the probability that the optimal treatment strategy (i.e., the strategy with the highest expected NMB) is cost-effective, is appropriate in this context.
+
+A CEAF curve can be created by using the \code{best} column from the \code{mce} element to subset to the treatment strategy with the highest expected NMB for each WTP value.
+
+<<ceaf>>=
+ggplot(cea_ictstm$mce[best == 1],
+       aes(x = k, y = prob, col = strategy_factor(strategy_id))) +
+  geom_line() +
+  xlab("Willingness to pay") +
+  ylab("Probability most cost-effective") +
+  scale_x_continuous(breaks = seq(0, max(wtp), length.out = 6),
+                     label = scales::dollar) +
+  theme(legend.position = "bottom") +
+  scale_colour_discrete(name = "Strategy")
+@
+
+\subsubsection{Expected value of perfect information}
+A limitation of the prior measures are that they ignore the magnitude of QALY or cost gains.
+A measure which combines the probability of being most effective with the magnitude of the expected NMB is the EVPI.
+Intuitively, the EVPI provides an estimate of the amount that a decision maker would be willing to pay to collect additional data and completely eliminate uncertainty.
+Mathematically, the EVPI is defined as the difference between the maximum expected NMB given perfect information and the maximum expected NMB given current information.
+In other words, we calculate the NMB for the optimal treatment strategy for each random draw of the parameters and compare that to the NMB for the treatment strategy that is optimal when averaging across all parameters,
+
+\begin{equation}
+EVPI = E_\theta \left[max_j NMB(j, \theta)\right] - max_j E_\theta \left [ NMB(j, \theta)\right].
+\end{equation}
+
+The \code{cea()} function performs the EVPI calculation across all simulation draws from the PSA and for a number of WTP values.
+A plot of the the EVPI is shown below.
+The kink represents the WTP value where the optimal strategy changes.
 

### 2020-12-21 manuscript.Rnw
Add formal figures for all plots

@@
 %% other packages
+\usepackage{graphicx}
+\usepackage{subcaption}
 \usepackage{algorithm}
@@
 
-<<ce_plane, fig.height = 4, fig.width = 6>>=
+<<ce_plane, fig.height = 4, fig.width = 6, fig.show="hide">>=
 library("scales")
@@
 
+\begin{figure}[t!]
+\centering
+\includegraphics{figure/ce_plane-1.pdf}
+\caption{Cost-effectiveness plane} \label{fig:ce-plane}
+\end{figure}
+
 \subsubsection{Cost-effectiveness acceptability curve}
@@
 
-<<ceac_simultaneous>>=
+<<ceac_simultaneous, fig.height = 4, fig.width = 6, fig.show = "hide">>=
 ggplot(cea_ictstm$mce,
@@
 
-<<ceac_pw>>=
+<<ceac_pw, fig.height = 4, fig.width = 6, fig.show = "hide">>=
 ggplot(cea_pw_ictstm$ceac,
@@
 
+\begin{figure}[t!]
+\centering
+\begin{subfigure}[b]{.8\textwidth}
+\includegraphics{figure/ceac_simultaneous-1.pdf}
+\caption{Simultaneous comparison} \label{fig:ceac-simultaneous}
+\end{subfigure}
+\begin{subfigure}[b]{.8\textwidth}
+\includegraphics{figure/ceac_pw-1.pdf}
+\caption{Pairwise comparison} \label{fig:ceac-pw}
+\end{subfigure}
+\caption{Cost-effectiveness acceptability curves}\label{fig:ceac}
+\end{figure}
+
 \subsubsection{Cost-effectiveness acceptability frontier}
@@
 
-<<ceaf>>=
+<<ceaf, fig.height = 4, fig.width = 6, fig.show = "hide">>=
 ggplot(cea_ictstm$mce[best == 1],
@@
 
+\begin{figure}[t!]
+\centering
+\includegraphics{figure/ceaf-1.pdf}
+\caption{Cost-effectiveness acceptability frontier} \label{fig:ceaf}
+\end{figure}
+
 \subsubsection{Expected value of perfect information}
@@
 
-<<evpi>>=
+<<evpi, fig.height = 4, fig.width = 6, fig.show = "hide">>=
 ggplot(cea_ictstm$evpi, aes(x = k, y = evpi)) +
@@
 
+\begin{figure}[t!]
+\centering
+\includegraphics{figure/evpi-1.pdf}
+\caption{Expected value of perfect information} \label{fig:evpi}
+\end{figure}
+
 %% -- Discussion ---------------------------------------------------------------

### 2020-12-23 manuscript.Rnw
Rough draft of discussion section

@@
 
-Markov and semi-Markov models can be formulated in either continuous or discrete time, at either the cohort or individual level.
+Markov and semi-Markov models can be formulated in either continuous or discrete time, at either the cohort or individual-level.
 In health economics, Markov cohort models tend to be formulated in discrete time (i.e., as cDTSTMs), although state probabilities can be computed in continuous time models using the Aalen-Johansen estimator \citep{aalen1978empirical} or the Kolmogorov forward equation \citep{cox1977theory}.
-While tunnel states can be used to approximate a semi-Markov model using a cohort approach, they can only be simulated in a general fashion using individual level models.
+While tunnel states can be used to approximate a semi-Markov model using a cohort approach, they can only be simulated in a general fashion using individual-level models.
 Discrete time individual simulation is possible, but we use continuous time models (i.e., iCTSTMs) because they do not require specification of model cycles and can be run considerably faster.
@@
 \subsection{Partitioned survival model} \label{sec:psm-example}
-In the previous section multi-state models were fit to individual patient data.
+In the previous section multi-state models were fit to individual patient data (IPD).
 Partitioned survival analysis is an alternative approach that is frequently used for economic evaluations of therapies in oncology.
@@
 Techniques for fitting multi-state models with aggregate data require more constraints and are less established \citep{price2011parameterization, pahuta2019technique, jansen2020multi}.
-PSMs, in contrast, consist of separate independent survival models that can be estimated with aggregate level data by reconstructing individual patient data from digitized survival curves \citep{guyot2012enhanced}.
+PSMs, in contrast, consist of separate independent survival models that can be estimated with aggregate level data by reconstructing IPD from digitized survival curves \citep{guyot2012enhanced}.
 
@@
 We do, however, modify drug costs.
-Since we are no longer using an individual level model, costs cannot vary as a function of time in a health state.
+Since we are no longer using an individual-level model, costs cannot vary as a function of time in a health state.
 We now assume conservatively that drug costs remain at initial level when in the progression state.
@@
 \subsubsection{Cost-effectiveness acceptability curve}
-A useful summary measure for quantifying uncertainty is the CEAC, which can generated based on simultaneous comparisons of all treatment strategies or by comparing each treatment strategy to a single comparator.
+A useful summary measure for quantifying uncertainty is the CEAC, which can be generated based on simultaneous comparisons of all treatment strategies or by comparing each treatment strategy to a single comparator.
 The former is the probability that each strategy is the most cost-effective and is available from the \code{mce} element produced by \code{cea()}.
@@
 \section{Discussion} \label{sec:discussion}
+This article provides a formal description of the economic models and cost-effectiveness framework used by the \pkg{hesim} package.
+An illustrative example is provided for the common three state model used to evaluate treatments in oncology.
+However, since \pkg{hesim} aims to be flexible, the example only covers a subset of the relevant modeling approaches and functionality.
+Additional examples as well as documentation for all functions and classes are consequently provided on the package website (\url{https://hesim-dev.github.io/hesim/}).
+
+One notable set of features not covered relates to the construction of transition probability matrices, the core of any cDTSTM.
+In Section~\ref{sec:cdtstm-example}, an approach was used that was tailored to a case in which transition intensity matrices were available for a reference treatment and relative treatment effects (i.e., relative risks) were available for the remaining competing interventions.
+To maximize efficiency, the matrix exponential was only applied where necessary (for SOC) and relative risks were applied to the transition probability matrices for SOC with the efficient \code{apply_rr()} function.
+
+Alternative approaches might be appropriate in other contexts.
+For example, if evenly spaced panel data is available, a multinomial logistic regressions can be used to predict transition probabilities.
+On the other hand, construction of the transition probability matrix and relative treatment effects may not be as straightforward.
+In these cases, it may be easier to use a \code{define_model()} block to construct transition probability matrices using nonstandard evaluation.
+When using this approach, each possible combination of the parameters and input data (treatment strategies and patients) is automatically constructed and the user must only write mathematical expressions that use the parameters and input data to create the transition probability matrices.
+The same \code{define_model()} block may also be used to construct utility and cost models, if desired.
+Example that leverage both multinomial logistic regession and \code{define_model()} are provided on the course website and the latter is used to reproduce two examples from the reference textbook by \citet{briggs2006decision}.
+
+In Section~\ref{sec:ictstm-example}, we fit a clock-reset multi-state model and simulated outcomes with an individual-level simulation.
+An iCTSTM can, however, also be developed without fitting a statistical model using \proglang{R} by creating a \code{params_surv} object for each possible transition.
+If IPD is available for the reference treatment, \code{create_params()} can be use to obtain the baseline \code{params_surv} object from the fitted multi-state model, and relative treatment effects (e.g., hazard ratios) could be added as additional columns to the coefficients stored in the \code{params_surv} object.
+A similar approach was used in the partitioned survival analysis example above (Section~\ref{sec:psm-example}).
+A re-analysis of an time-inhomogeneous Markov cohort model of total hip replacement from the \citet{briggs2006decision} textbook is provided on the package website to illustrate.
+
+A third additional feature worth mentioning is that a PSM can be extended to N-states.
+The 3-state oncology case might, for instance, be extended to account for adverse events or to multiple treatment lines.
+One related application is the quality-adjusted time without symptoms or toxicity (Q-TWiST) method \citep{goldhirsch1989costs, lenderking1994eval}, which partitions patients into time (i) having toxic side effects, (ii) without symptoms or toxicities, and (iii) post-progression.
+An example 4-state model is provided on the package website.
+
+As mentioned in the introduction, \code{heemod} is the only other general purpose \proglang{R} package for cost-effectiveness modeling.
+It differs from \pkg{hesim} in that it focuses solely on cDTSTMs (i.e., Markov cohort models) and is not designed to facilitate integration of economic models with the statistical models used for parameter estimation.
+Another important difference is that \code{hesim} is considerably faster since all simulations are vectorized (i.e, performed at the \proglang{C++} level).
+Performance tests are provided on the package website.
+The time-inhomogeneous total hip replacement \citet{briggs2006decision} example was run with both packages using 1,000 PSA iterations.
+A cohort model in \pkg{heemod} completed in approximately 85 seconds while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \code{hesim} ran in 9 seconds.
+
+There are a number of \proglang{R} multi-state modeling packages that can both fit models and make predictions.
+\pkg{hesim} makes use of the \pkg{flexsurv} and \pkg{msm} packages for parameter estimation with parametric models A key difference is that \pkg{hesim} is designed specifically for health-economic modeling whereas the other packages are more general purpose statistical packages.
+
+\pkg{mstate} is another package for multi-state modeling that is designed for estimation of non-parametric and semi-parametric (i.e., Cox proportional hazards) models.
+It supports prediction of state probabilities and simulation of multi-state trajectories (for clock-reset models) from cumulative hazard functions evaluated at specific times.
+\code{williams2017cost} has adapted \pkg{mstate} for cost-effectiveness modeling applications.
+There are however, some disadvantages of using \pkg{mstate} for health-economic modeling since it was not designed for it.
+For instance, the core function for simulating multi-state models, \code{mstate::mssample()}, does not natively support heterogeneous target populations, PSA, simultaneous simulation of multiple treatment strategies, or computation of QALYs and costs from the simulated output.
+It is also computationally inefficient for two main reasons.
+First, simulation of general cumulative hazards functions is considerably slower than simulation of parametric distributions where efficient random number generators are available.
+Second, it is written completely in \proglang{R} and incorporating PSA, multiple treatment strategies, or heterogeneous patients results in code that is not vectorized.
+To compare computational performance, we ran 100 PSA iterations where 1 treatment strategy and 1,000 homogeneous patients were simulated through a 6-state Weibull model during each iteration.
+The model completed in 0.44 second in \pkg{hesim} and took 34 minutes in \pkg{mstate}, showing that computational considerations are not just an academic concern.
+
+A number of extensions of the current version of \pkg{hesim} would be helpful.
+First, we plan to extend the individual-level simulation so that users can update the input data during the simulation as a function of time elapsed and/or the health states transitioned to.
+This would allow potentially time-varying variables such as age to update each time a patients transitions to a new state.
+Second, we aim to continue to add more examples to ensure that \pkg{hesim} is fit for purpose for a wide range of applications in health economics.
+Third, more complex models of utility \citep{kharroubi2007modelling, hernandez2013relationship} and costs \citep{nixon2005methods} could help better capture heterogeneity in preferences and spending.
+Finally, a larger aspiration is to provide support for parameterization via NMA, particularly for estimation of relative treatment effects in survival (e.g., for PSMs) and multi-state models.
+However, since existing \proglang{R} packages are lacking, this could involve creation of a separate NMA package and a new dependency for \pkg{hesim}.
 

### 2020-12-23 manuscript.Rnw
Some updates to the discussion

@@
 \section{Discussion} \label{sec:discussion}
-This article provides a formal description of the economic models and cost-effectiveness framework used by the \pkg{hesim} package.
-An illustrative example is provided for the common three state model used to evaluate treatments in oncology.
+This paper formally describes the economic models and cost-effectiveness framework used by the \pkg{hesim} package.
+An illustrative example is provided for a three state model commonly used to evaluate treatments in oncology.
 However, since \pkg{hesim} aims to be flexible, the example only covers a subset of the relevant modeling approaches and functionality.
@@
 For example, if evenly spaced panel data is available, a multinomial logistic regressions can be used to predict transition probabilities.
-On the other hand, construction of the transition probability matrix and relative treatment effects may not be as straightforward.
-In these cases, it may be easier to use a \code{define_model()} block to construct transition probability matrices using nonstandard evaluation.
-When using this approach, each possible combination of the parameters and input data (treatment strategies and patients) is automatically constructed and the user must only write mathematical expressions that use the parameters and input data to create the transition probability matrices.
+On the other hand, some users may prefer a functional approach that is arguably more human readable.
+To facilitate this type of approach, a \code{define_model()} block can be used to construct transition probability matrices using nonstandard evaluation.
+Each possible combination of the parameters and input data (treatment strategies and patients) is automatically constructed and the user can write custom expressions that create the transition probability matrices as a function of the parameters and input data.
 The same \code{define_model()} block may also be used to construct utility and cost models, if desired.
-Example that leverage both multinomial logistic regession and \code{define_model()} are provided on the course website and the latter is used to reproduce two examples from the reference textbook by \citet{briggs2006decision}.
+Example that leverage multinomial logistic regression and \code{define_model()} are provided on the package website.
+The latter is used to reproduce two examples (a simple Markov model of an HIV treatment and a time-inhomogeneous Markov cohort model of total hip replacement) from the reference textbook by \citet{briggs2006decision}.
 
 In Section~\ref{sec:ictstm-example}, we fit a clock-reset multi-state model and simulated outcomes with an individual-level simulation.
-An iCTSTM can, however, also be developed without fitting a statistical model using \proglang{R} by creating a \code{params_surv} object for each possible transition.
-If IPD is available for the reference treatment, \code{create_params()} can be use to obtain the baseline \code{params_surv} object from the fitted multi-state model, and relative treatment effects (e.g., hazard ratios) could be added as additional columns to the coefficients stored in the \code{params_surv} object.
+An iCTSTM can, however, also be developed without fitting a statistical model using \proglang{R} by manually inputting the arguments (e.g., coefficients, probability distributions) of \code{params_surv} objects for each possible transition.
+If IPD is available for the reference treatment, \code{create_params()} can be use to obtain the baseline \code{params_surv} object from a fitted multi-state model, and relative treatment effects (e.g., hazard ratios) could be added as additional columns to the coefficients stored in the \code{params_surv} object.
 A similar approach was used in the partitioned survival analysis example above (Section~\ref{sec:psm-example}).
-A re-analysis of an time-inhomogeneous Markov cohort model of total hip replacement from the \citet{briggs2006decision} textbook is provided on the package website to illustrate.
-
-A third additional feature worth mentioning is that a PSM can be extended to N-states.
-The 3-state oncology case might, for instance, be extended to account for adverse events or to multiple treatment lines.
-One related application is the quality-adjusted time without symptoms or toxicity (Q-TWiST) method \citep{goldhirsch1989costs, lenderking1994eval}, which partitions patients into time (i) having toxic side effects, (ii) without symptoms or toxicities, and (iii) post-progression.
+A re-analysis of the total hip replacement model from the \citet{briggs2006decision} textbook using an individual-levels simulation is provided on the package website to illustrate.
+
+A third feature worth mentioning is that a PSM can be extended to $N$ health states.
+The 3-state oncology case might, for instance, be extended to account for adverse events or multiple lines of therapy.
+One related application is the quality-adjusted time without symptoms or toxicity (Q-TWiST) method \citep{goldhirsch1989costs, lenderking1994eval}, which partitions patients into time spent (i) having toxic side effects, (ii) without symptoms or toxicities, and (iii) after progression.
 An example 4-state model is provided on the package website.
@@
 Another important difference is that \code{hesim} is considerably faster since all simulations are vectorized (i.e, performed at the \proglang{C++} level).
-Performance tests are provided on the package website.
+Performance comparisons are provided on the package website.
 The time-inhomogeneous total hip replacement \citet{briggs2006decision} example was run with both packages using 1,000 PSA iterations.
@@
 
-There are a number of \proglang{R} multi-state modeling packages that can both fit models and make predictions.
-\pkg{hesim} makes use of the \pkg{flexsurv} and \pkg{msm} packages for parameter estimation with parametric models A key difference is that \pkg{hesim} is designed specifically for health-economic modeling whereas the other packages are more general purpose statistical packages.
+There are a number of multi-state modeling \proglang{R} packages that can both fit models and make predictions.
+\pkg{hesim} leverages some of these for parameters estimation: \pkg{flexsurv} is used to fit parametric and flexible parametric survival and multi-state models when exact event times are available while \pkg{msm} is used to fit multi-state models when only panel data is available.
+A key difference is that \pkg{hesim} is designed specifically for health-economic modeling whereas the other packages are more general purpose statistical packages.
 
 \pkg{mstate} is another package for multi-state modeling that is designed for estimation of non-parametric and semi-parametric (i.e., Cox proportional hazards) models.
-It supports prediction of state probabilities and simulation of multi-state trajectories (for clock-reset models) from cumulative hazard functions evaluated at specific times.
-\code{williams2017cost} has adapted \pkg{mstate} for cost-effectiveness modeling applications.
+It supports prediction of state probabilities and simulation of multi-state trajectories (to facilitate prediction for clock-reset models) from cumulative hazard functions evaluated at specific times.
+\citet{williams2017cost} has adapted \pkg{mstate} for cost-effectiveness modeling applications.
 There are however, some disadvantages of using \pkg{mstate} for health-economic modeling since it was not designed for it.
-For instance, the core function for simulating multi-state models, \code{mstate::mssample()}, does not natively support heterogeneous target populations, PSA, simultaneous simulation of multiple treatment strategies, or computation of QALYs and costs from the simulated output.
+\code{mssample()}---the core function for simulating multi-state models---does not natively support heterogeneous target populations, PSA, simultaneous simulation of multiple treatment strategies, or computation of QALYs and costs from the simulated output.
 It is also computationally inefficient for two main reasons.
@@
 A number of extensions of the current version of \pkg{hesim} would be helpful.
-First, we plan to extend the individual-level simulation so that users can update the input data during the simulation as a function of time elapsed and/or the health states transitioned to.
-This would allow potentially time-varying variables such as age to update each time a patients transitions to a new state.
-Second, we aim to continue to add more examples to ensure that \pkg{hesim} is fit for purpose for a wide range of applications in health economics.
+First, we plan to extend the individual-level simulation so that users can update input data during the simulation as a function of time elapsed and/or the health states transitioned to.
+This would allow potentially time-varying variables such as age to update each time a patient transitions to a new state.
+Second, we would like to add more examples to ensure that \pkg{hesim} is fit for purpose for a wide range of applications in health economics.
 Third, more complex models of utility \citep{kharroubi2007modelling, hernandez2013relationship} and costs \citep{nixon2005methods} could help better capture heterogeneity in preferences and spending.
 Finally, a larger aspiration is to provide support for parameterization via NMA, particularly for estimation of relative treatment effects in survival (e.g., for PSMs) and multi-state models.
-However, since existing \proglang{R} packages are lacking, this could involve creation of a separate NMA package and a new dependency for \pkg{hesim}.
+However, since existing \proglang{R} packages are lacking in this area, this could involve creation of a separate NMA package and a new dependency for \pkg{hesim}.
 

### 2020-12-23 manuscript.Rnw
Tweaks to intro

@@
 \pkg{hesim} is an \proglang{R} package that advances the functionality and performance of the existing software.
-Multiple model types are supported including cohort discrete time state transition models (DTSTMs), N-state PSMs, and individual-level continuous time state transition models (CTSTMs), encompassing both Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
-To maximize flexibility and facilitate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R}, by inputting parameters obtained from external sources.
+Multiple model types are supported including cohort discrete time state transition models (cDTSTMs), N-state PSMs, and individual-level continuous time state transition models (iCTSTMs), encompassing both Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
+To maximize flexibility and facilitate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R} or by inputting parameters obtained from external sources.
 So that individual-level simulation and PSA can be run quickly, \pkg{Rcpp} and \pkg{data.table} are heavily utilized.
@@
 
-The remainder of this article is organized as follows.
-Section~\ref{sec:model-taxonomy} describes the economic models supported by \pkg{hesim}.
+The remainder of this paper is organized as follows.
+Section~\ref{sec:model-taxonomy} provides a mathematical description the economic models supported by \pkg{hesim}.
 An overview of the coding framework is provided in Section~\ref{sec:framework}.
 An illustrative example using a three state model of disease progression in oncology is provided in Section~\ref{sec:illustrative-example}.
-Analyses are conducted for each of the three supported model types, illustrating approaches suitable for both patient level and aggregate level data.
+Analyses are conducted for each of the three supported model types, illustrating approaches suitable for both patient-level and aggregate-level data.
 The cost-effectiveness framework is described in Section~\ref{sec:cea} along with worked examples based on the output from the prior section.
-Section~\ref{sec:discussion} makes comparisons to other software and discusses possible extensions.
+Section~\ref{sec:discussion} discusses additional features not covered in this paper, makes comparisons to other software, and explores possible extensions.
 Finally, Section~\ref{sec:conclusion} concludes.
 
-Analyses were performed using \pkg{hesim} version 0.5.0.
+Analyses were performed using \pkg{hesim} version \code{0.5.0}.
 

### 2020-12-23 manuscript.Rnw
First draft of conculsion

@@
 \section{Conclusion} \label{sec:conclusion}
+\pkg{hesim} is a modular and computationally efficient \proglang{R} package for development and analysis of simulation models for economic evaluation.
+Discrete-time cohort and continuous-time individual-level models are supported, encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes as well as partitioned survival analysis.
+To facilitate integration of statistical methods and economic models and use of both patient- and aggregate-level data, parameters can be estimated by fitting a statistical model with \proglang{R}, inputting parameter estimates from an external source, or a combination of the two.
+Simulated costs and QALYs from a PSA can be used for for decision-analysis within a cost-effectiveness framework, allowing users to represent decision uncertainty (e.g., cost-effectiveness planes, cost-effectiveness acceptability curves, and cost-effectiveness acceptability frontiers), and conduct value of information analysis.
+Algorithms are efficiently coded and written in \proglang{C++} so that approaches that have been considered computationally intensive such as individual-level simulations and PSA can run very quickly.
+The modular design makes modeling more flexible and makes it straightforward to add new features.
 

### 2020-12-23 manuscript.Rnw
Update cDTSTM mathematical description

@@
 STMs simulate transitions between mutually exclusive health states.
-A common assumption is that the STM is a Markov model, meaning that transitions to the next health state can only depend on the present health state.
+A common assumption is that a STM is a Markov model, meaning that transitions to the next health state can only depend on the present health state.
 In a time homogeneous Markov model transition probabilities are constant over time, whereas in a time inhomogeneous model they can depend on time since the start of the model.
@@
 
-Markov and semi-Markov models can be formulated in either continuous or discrete time, at either the cohort or individual-level.
-In health economics, Markov cohort models tend to be formulated in discrete time (i.e., as cDTSTMs), although state probabilities can be computed in continuous time models using the Aalen-Johansen estimator \citep{aalen1978empirical} or the Kolmogorov forward equation \citep{cox1977theory}.
+Markov and semi-Markov models can be formulated in either continuous or discrete time, at either the cohort- or individual-level.
+In health economics, Markov cohort models tend to be formulated in discrete time (i.e., as cDTSTMs), although state probabilities can be computed with continuous time models using the Aalen-Johansen estimator \citep{aalen1978empirical} or the Kolmogorov forward equation \citep{cox1977theory}.
 While tunnel states can be used to approximate a semi-Markov model using a cohort approach, they can only be simulated in a general fashion using individual-level models.
@@
 cDTSTMs simulate the probability that a cohort of patients is in each of $H$ health states over time.
-Time is measured at discrete times with each time point known as a model cycle.
-A $1 \times H$ state vector that stores the probability of being in each health state at time $t$ is written as $x_t = (x_{1t}, x_{2t}, \ldots, x_{Ht})$ where $\sum_i x_{ht} = 1$.
-A transition probability matrix is denoted by $P_t$ where the $(r,s)$th element represents a transition from state $r$ to state $s$ between times $t$ and $t+1$.
-The state vector at time $t+1$ for each of $T$ model cycles is given by,
+Time is partitioned into intervals $\{[0, t_1), [t_1, t_2), \ldots, [t_C, t_{C+1})\}$ with each interval $0,\ldots, C$ known as a model cycle.
+An $H \times 1$ state vector that stores the probability of being in each health state at cycle $c$ is written as $x_c = (x_{1c}, x_{2c}, \ldots, x_{Hc})$ where $\sum_h x_{hc} = 1$.
+A transition probability matrix is an $H \times H$ matrix denoted by $P_c$ where the $(r,s)$th element represents a transition from state $r$ to state $s$ between model cycles $c$ and $c+1$.
+The state vector at cycle $c+1$ for each of $C$ model cycles is given by,
 
 \begin{equation} \label{eqn:markov-sim}
-x_{t+1}^T= x_t^T P_t, \quad t = 0,\ldots, T.
+x_{c+1}^T= x_c^T P_c, \quad c = 0,\ldots, C.
+\end{equation}
+
+The transition probability matrices can also be characterized in terms of transition intensity matrices, $Q_c$, of the same dimensions as $P_c$ and representing an underlying continuous process with a patient in state $X(t)$ at exact time $t$.
+The $(r,s)$th element is the hazard representing the instantaneous risk of moving from state $r$ to state $s$,
+
+\begin{equation} \label{eqn:hazard}
+h_{rs}(t) = \lim_{\Delta t\to 0} \frac{\Prob(X(t + \Delta t) = s | X(t) = r)}{\Delta t}.
+\end{equation}
+
+Each row of $Q_c$ sums to 0 whereas each row of $P_c$ sums to 1.
+If transition intensities are constant over the interval $(t_c, t_{c+1})$ for a model cycle with length $u = t_{c+1}-t_c$, then the transition probability matrix can be solved with the matrix exponential as,
+
+\begin{equation} \label{eqn:expmat}
+P_c = \textrm{Exp}(uQ_c).
 \end{equation}
@@
 
-Expected values are computed by integrating the "weighted" probability of being in each state, where weights are a function of the discount factor and state values.
+Expected values are computed by integrating the ``weighted'' probability of being in each state, where weights are a function of the discount factor and state values.
 That is, for a time horizon $T$, discounted costs and QALYs in health state h are computed as,
@@
 \begin{equation} \label{eqn:cohort-costs-qalys}
-\int_0^{T} z_h(t) e^{-rt} P_h(t) dt,
+\int_0^{T} z_h(t) e^{-rt} p_h(t) dt,
 \end{equation}
 
-where $z_h(t)$ is the predicted cost or utility value at time $t$, $r$ is the discount rate, and $P_h(t)$ is the probability of being in a given health state.
-
-In a discrete time, the integral is approximated with
+where $r$ is the discount rate and at time $t$, $z_h(t)$ is the predicted cost or utility value and $p_{h}(t) = \Prob(X(t) = h)$ is the probability of being in health state $h$.
+
+In discrete time, the integral is approximated with
 
 \begin{equation} \label{eqn:riemann-sum}
-\sum_{j = 1}^{T}f(t_j^{*})\Delta t_j
+\sum_{c = 0}^{C}f(t_c^{*})\Delta t_c,
 \end{equation}
 
-where $\Delta t_j = t_j - t_{j-1}$, $t_j^{*}\in[t_{j-1},t_j]$, and $f(t_j^{*})= z_h(t_j^{*}) e^{-rt_j^{*}} P_h(t_j^{*})$.
-Three methods can be used to estimate $f(t_j^{*})$.
-First, a left Riemann sum uses values at the start of each time interval $f(t_j^{*}) = f(t_{j-1})$.
-Second, a right Riemann sum uses values at the end of each time interval $f(t_j^{*}) = f(t_j)$.
-Finally, the trapezoid rule averages values at the start and end of each interval $f(t_j^{*}) = \frac{1}{2}\Delta t_j[f(t_{j-1}) + f(t_j)]$.
+where $\Delta t_c = t_{c+1} - t_{c}$,  $t_c^{*}\in[t_{c},t_{c+1}]$, and $f(t)= z_h(t) e^{-rt} p_h(t)$.
+Three methods can be used to estimate $f(t_c^{*})$.
+First, a left Riemann sum uses values at the start of each time interval $f(t_c^{*}) = f(t_c)$.
+Second, a right Riemann sum uses values at the end of each time interval $f(t_c^{*}) = f(t_{c+1})$.
+Finally, the trapezoid rule averages values at the start and end of each interval $f(t_c^{*}) = \frac{1}{2}\Delta t_c[f(t_c) + f(t_{c+1})]$.
 
@@
 In continuous time, a patient is in state $X(t)$ at time $t$.
-State transitions are modeled using a multi-state modeling framework \citep{putter2007tutorial} where the probability of a transition from state $r$ to state $s$ is governed by the hazard function,
-
-\begin{equation}
-h_{rs}(t) = \lim_{\Delta t\to 0} \frac{\Prob(X(t + \Delta t) = s | X(t) = r)}{\Delta t}.
-\end{equation}
+State transitions are modeled using a multi-state modeling framework \citep{putter2007tutorial} where the probability of a transition from state $r$ to state $s$ is governed by the hazard function from Equation~\ref{eqn:hazard}.
 
@@
 
-The analysis is performed in the next three steps as shown in Figure~\ref{fig:econ-eval-process-hesim}.
+Development and analysis of the model is performed in the next three steps as shown in Figure~\ref{fig:econ-eval-process-hesim}.
 First, the disease progression, costs, and utility submodels are parameterized using ``estimation'' datasets.

### 2020-12-23 manuscript.Rnw
Updates to model taxonomy section

@@
 cDTSTMs simulate the probability that a cohort of patients is in each of $H$ health states over time.
-Time is partitioned into intervals $\{[0, t_1), [t_1, t_2), \ldots, [t_C, t_{C+1})\}$ with each interval $0,\ldots, C$ known as a model cycle.
+Time is partitioned into intervals $\{[0, t_1), [t_1, t_2), \ldots, [t_C, t_{C+1})\}$ with each interval $0,\ldots, C$ reffered to as a model cycle.
+We denote $t_c$ as the time at the start of model cycle $c$.
 An $H \times 1$ state vector that stores the probability of being in each health state at cycle $c$ is written as $x_c = (x_{1c}, x_{2c}, \ldots, x_{Hc})$ where $\sum_h x_{hc} = 1$.
@@
 
-In $N$-state model, $N-1$ non-mutually exclusive survival curves are required.
+In an $N$-state model, $N-1$ non-mutually exclusive survival curves are required.
 The cumulative survival function, $S_n(t)$, represents the probability that a patient survives to health state $n$ or to a lower indexed state beyond time $t$.
@@
 The $n$th fitted survival model with density $f_n(t)$ has cumulative density function $F_n(t)$, survivor function $1 - F_n(T)$, cumulative hazard $H_n(t) = -\log S_n(t)$, and hazard $h_n(t) = f_n(t)/S_n(t)$.
-State probabilities for each health states can be computed for an arbitrarily fine grid of time points to produce the health state vector $x_t$ for each time $t$ in the grid.
+State probabilities for each health state, $x_{hc}$, can be computed for an arbitrarily fine grid of time periods $c$.
 Costs and QALYs are then computed in the same manner as in the cohort model described in Section~\ref{sec:cDTSTMs}.

### 2020-12-23 manuscript.Rnw
Minor updates to framework section

@@
 Model development proceeds in 4 steps.
-The first step is to setup the model by defining the model structure, target population, and treatment strategies of interest, and storing the relevant information in a \code{hesim_data} object.
+The first step is to setup the model by defining the model structure, target population, and treatment strategies of interest.
+The relevant information is then stored in a \code{hesim_data} object.
 
@@
 The input data always contains variables describing the target population and treatment strategies of interest, but can also contain variables related to time to facilitate use of time-varying covariates or parameters.
-Finally, the simulated outcomes are used to perform decision analysis within a cost-effectiveness framework.
+Finally, the simulated QALYs and costs are used to perform decision analysis within a cost-effectiveness framework.
 While other approaches such as multi-criteria decision analysis (MCDA) could, in principle, be used as well, only CEA is currently supported.
@@
 Each submodel contains fields for the model parameters and the input data.
-Parameters can either be created from a statistical model fit using \proglang{R} or input directly by the user.
-There are two types of parameter objects, standard parameter objects prefixed by \code{``params''} and transformed parameter objects prefixed by \code{``tparams''}.
+Parameters can either be estimated from a statistical model fit using \proglang{R} or input directly by the user.
+There are two types of parameter objects, standard parameter objects prefixed by \code{``params''} and ``transformed'' parameter objects prefixed by \code{``tparams''}.
 The former contain the underlying parameters of a statistical model while the latter contain parameters more immediate to prediction that have been transformed as function of the input data.
-The regression coefficients of a logistic regression are example of a parameter objects while the predicted probabilities are examples of a transformed parameter object.
+The regression coefficients of a logistic regression are an example of a parameter objects while the predicted probabilities are examples of a transformed parameter object.
 
 Transformed parameter objects are often useful when parameterizing a model using external sources.
-Two common examples are predicted probabilities for transition matrices and predicted means for assigning values to health states.
-The advantage of a transformed parameter object is that an explicit statistical model does not need to be specified which affords more flexibility to the user.
+Two common examples are the predicted probabilities comprising transition probability matrices and predicted means for assigning values to health states.
+The advantage of a transformed parameter object is that an explicit statistical model does not need to be specified which gives the user more flexibility.
 
@@
 The latter is useful for storing the parameters of a multi-state model or the independent survival models required for a PSM.
-The parameters of a multinomial logistic regression are stored in a \code{params_mlogit} and can be created by fitting a model with \code{nnet::multinom()}.
-
-\code{tparams_transprobs} objects contain transition probability matrices that have been predicted for each PSA sample, treatment strategy, and patient from the target population.
-Transition probabilities may also vary by time interval to allow for time inhomogeneous Markov models.
+The parameters of a multinomial logistic regression are stored in a \code{params_mlogit} and can be created by fitting a model for each row in a transition probability matrix with \code{nnet::multinom()}.
+
+\code{tparams_transprobs} objects store transition probability matrices that have been predicted for each PSA sample, treatment strategy, patient from the target population, and optionally time interval (to allow for time inhomogeneous Markov models).
 Transition probabilities can be predicted from a fitted multi-state model using the \pkg{msm} package or constructed ``by hand'' in a custom manner.
@@
 \subsubsection{Costs and utility} 
-State values (i.e., costs and utilities) do not depend on the choice of disease model. They can currently either be modeled using a linear model or using predicted means.
+State values (i.e., costs and utilities) do not depend on the choice of disease model.
+They can currently either be modeled using a linear model or using predicted means (Table~\ref{tbl:parameterize-stateval-model}).
 
 \begin{table} [h]
-\caption{Parameterization of state value models}\label{tbl:parameterize-disease-model}
+\caption{Parameterization of state value models}\label{tbl:parameterize-stateval-model}
 \footnotesize
@@
 predicted means) to health states that can vary across PSA samples, treatment strategies, patients, and/or time intervals.
-State values can be specified either as moments (i.e., mean and standard error) or parameters (e.g., shape and scale of gamma distribution) of a probability distribution, or by pre-simulating values from a suitable probability distribution (e.g., from a Bayesian model).
+State values can be specified either as moments (e.g., mean and standard error) or parameters (e.g., shape and scale of gamma distribution) of a probability distribution, or by pre-simulating values from a suitable probability distribution (e.g., from a Bayesian model).
 
@@
 
-The disease and state value models are most easily instantiated using \code{S3} generic functions prefixed by ``\code{create}'' from parameter, transformed parameter, or statistical model objects.
-An \code{IndivCtstmTrans} can, for instance, be created from a \code{flexsurvreg_list} or \code{params_surv_list} object with \code{create_IndivCtstmTrans()}.
+The disease and state value models can either be instantiated from parameter or model objects using \code{S3} generic functions prefixed by ``\code{create}'' or using the \pkg{R6} constructor method \code{$new()}.
+\code{create_IndivCtstmTrans()} is an example of the former that can be used to create an \code{IndivCtstmTrans} object from a \code{flexsurvreg}, \code{flexsurvreg_list}, \code{params_surv}, or \code{params_surv_list} object.
 Similarly, a \code{StateVals} object can, for example, be created from a \code{stateval_tbl} object with \code{create_StateVals()}.
-The complete economic model is instantiated by combining the submodels using the \pkg{R6} constructor method \code{$new()}.
+The complete economic model is instantiated by combining the submodels using \code{$new()}.
 
 Each economic model contains methods for simulating disease progression, QALYs, and costs (Table~\ref{tbl:simulate-outcomes}).
-The cost and utility models always simulate costs and QALYs from the simulated  progression of disease with the methods \code{$sim_qalys()} and \code{$sim_costs()}, respectively.
+The utility and cost models always simulate QALYs and costs from the simulated progression of disease with the methods \code{$sim_qalys()} and \code{$sim_costs()}, respectively.
 Methods for simulating disease progression differ slightly since the simulation approaches differ as described in Section~\ref{sec:model-taxonomy}.
-In an N-state PSM $n-1$ survival curves are generated and in a iCTSTM a trajectory through state transitions is simulated for multiple patients via individual simulation.
+In an N-state PSM $n-1$ survival curves are generated and in a iCTSTM a trajectory through the STM is simulated for multiple patients via individual simulation.
 All models can simulate state probabilities over time.

### 2020-12-24 manuscript.Rnw
Updates to PSM and iCTSTM subsections of illustrative example

@@
 \section{Illustrative example} \label{sec:illustrative-example}
-We consider a three state model commonly used describe disease progression in oncology.
+We consider a three-state model commonly used in oncology.
 As shown in Figure~\ref{fig:three-state-model}, the three health states are stable disease (i.e.
-not progressed), progressed, and death.
-Patients can transition to a more severe health state (stable $\rightarrow$ progressed, stable $\rightarrow$ death, and progression $\rightarrow$ death) but cannot recover to a less severe health state.
+not progressed), progression, and death.
+Patients can transition to a more severe health state (stable $\rightarrow$ progression, stable $\rightarrow$ death, and progression $\rightarrow$ death) but cannot recover to a less severe health state.
 We assume that patients remain on first line (1L) treatment until progression, at which time they switch to a second line (2L) treatment.
@@
 There are 3 treatment strategies: standard of care (SOC) and two new interventions (New 1 and New 2).
-The target population consists of a heterogeneous population of 1000 patients, which is large enough to ensure that averages across patients in the individual simulation are reasonably stable.
+The target population consists of a heterogeneous population of 1,000 patients, which is large enough to ensure that averages across patients in the individual simulation are reasonably stable.
+As we will see in the next section, patients can also be placed into subgroups with a \code{grp_id} column in the \code{patients} table; if left unspecified as done here, then all patients are assumed to belong to a single subgroup.
 
@@
 In a STM, the disease model governs transitions between health states.
-For now, we assume that that patient level data with exact transition times is available so that a multi-state statistical model can be used to parameterize the disease model.
-
-Multi-state data consists of one row for each possible transition from a given health state where only one transition is observed and all others are censored.
+For now, we assume that individual patient data (IPD) with exact transition times is available so that a multi-state statistical model can be used to parameterize the disease model.
+
+Multi-state data consists of one row for each possible transition from a given health state where at most one transition is observed and all others are censored.
 We use the simulated dataset \code{onc3} from the \pkg{hesim} package.
@@
 Multi-state models can be fit in one of two ways.
-First, a joint survival model with interaction terms for different transition can be estimated, which is useful if there are constraints in the parameters across
-transitions, such as coefficients that are assumed equal across transitions.
+First, a joint survival model with interaction terms for different transition can be estimated, which is useful if there are constraints in the parameters, such as coefficients that are assumed equal across transitions.
 Here, we will take a second approach that is computationally more efficient and fit a separate model for each transition.
-We illustrate with parametric Weibull models but multiple distributions should be compared in a real application \citep{williams2017cost}.
+We illustrate with parametric Weibull models but multiple distributions should be compared and evaluated in a real application \citep{williams2017cost}.
 The treatment effect is assumed to only influence transitions from stable disease since patients are assumed to switch to a 2L treatment after progression.
@@
 Both medical and drug costs are considered.
-Medical costs are assumed to follow a gamma distribution which is often appropriate as they tend to be right skewed.
+Medical costs are assumed to follow a gamma distribution which is often appropriate for costs since they tend to be right skewed.
 Like utility, the distribution is characterized by the mean (\$2,000 with stable disease, \$9,500 with progressed disease) and standard error (assumed equal to the mean) and the method of moments is used to derive the underlying shape and scale parameters.
@@
 Drug costs are assumed to be fixed (i.e., measured without uncertainty).
-Costs are \$2,000 with SOC, \$12,000 with New 1, and \$15,000 with New 2 when on 1L treatment.
+When on 1L treatment, costs are \$2,000 with SOC, \$12,000 with New 1, and \$15,000 with New 2.
 All patients are assumed to switch to the same treatment at 2L after progression, which costs \$1,500 for the first 3 months and \$1,200 thereafter.
@@
 \subsubsection{Simulation}
-Before constructing the economic model, we specify the number of parameters samples that will be drawn for each submodel for the PSA.
+Before constructing each submodel, we specify the number of parameters samples that will be drawn for the PSA.
 
@@
 We must also specify the allowed transitions, the desired number of PSA samples (drawn from the multivariate normal distribution of the maximum likelihood estimate of the regression coefficients), the "clock" (recall that we used a clock reset approach when fitting the Weibull models), and the starting age of the patients.
-The starting age of the patients is useful because it allows us to specify a maximum age for patients during the simulation, ensuring that the simulated patients do not live unrealistically long lives.
+Note that the starting age of the patients is useful because it allows us to specify a maximum age for patients during the simulation, ensuring that the simulated patients do not live unrealistically long lives.
 
@@
 The utility and cost models can be easily constructed from the \code{stateval_tbl} objects with the \code{create_StateVals()} function, in which the (transformed) parameters are \code{tparams_mean} objects.
-The \code{hesim_data} object provides the information required to ensures that state values are constant within groups not specified within a particular \code{stateval_tbl}; for instance, since utility only varies by health state, it is assumed constant across patients, treatment strategies, and time.
+The \code{hesim_data} object provides the information required to ensure that state values are constant within groups not specified within a particular \code{stateval_tbl}; for instance, since utility only varies by health state, it is assumed constant across patients, treatment strategies, and time.
 The models for each cost category are combined in a list.
@@
 Once the economic model has been constructed, it is straightforward to simulate outcomes.
-A trajectory through the multi-state model is simulated with the \code{$sim_disease()} method with patients assumed to live to a maximum age of 100.
+Trajectories through the multi-state model are simulated with the \code{$sim_disease()} method with patients assumed to live to a maximum age of 100.
 
@@
 For a given subgroup, QALYs and costs are simulated for each PSA sample, treatment strategy, and health state.
-They are summarized (i.e., averaged) across PSA samples and treatment strategies (and subgroups) with the \code{$summarize()} method.
+They are summarized with the \code{$summarize()} method, which first sums across health states and then computes a (weighted) average across patients, by PSA sample, treatment strategy, and subgroup.
 
@@
 \subsection{Partitioned survival model} \label{sec:psm-example}
-In the previous section multi-state models were fit to individual patient data (IPD).
+In the previous section a multi-state model was fit to IPD.
 Partitioned survival analysis is an alternative approach that is frequently used for economic evaluations of therapies in oncology.
-Although there is increasing evidence that PSM are more biased than multi-state models \citep{woods2020partitioned}, PSMs are easier to parameterize with aggregate level data.
-Techniques for fitting multi-state models with aggregate data require more constraints and are less established \citep{price2011parameterization, pahuta2019technique, jansen2020multi}.
-PSMs, in contrast, consist of separate independent survival models that can be estimated with aggregate level data by reconstructing IPD from digitized survival curves \citep{guyot2012enhanced}.
+Although there is reason to believe that PSMs are more biased than multi-state models \citep{woods2020partitioned}, PSMs are easier to parameterize with aggregate-level data because techniques for fitting multi-state models with aggregate data require more constraints and are less established \citep{price2011parameterization, pahuta2019technique, jansen2020multi}.
+PSMs, in contrast, consist of separate independent survival models that can be estimated with aggregate-level data by reconstructing IPD from digitized survival curves \citep{guyot2012enhanced}.
 
 Since a PSM is a cohort model, the \code{hesim_data} object must be modified accordingly.
-We now specify 4 representative patients that comprise the cohort, each given an equal weight.
+We now assume that the cohort can be characterized by four representative patients, each given an equal weight.
 We consider separate subgroups (identified with \code{grp_id}) for males and females.
@@
 \subsubsection{Parameterization}
-A multi-state dataset can be converted into dataset with one row per patients and outcomes for PFS and OS using the function \code{as_pfs_os()}.
+A multi-state dataset can be converted into dataset with one row per patient and outcomes for PFS and OS using the function \code{as_pfs_os()}.
 
@@
 Survival models must be fit for both PFS and OS.
-To imitate a realistic scenario, we assume that patient level data is only available for SOC and that relative treatment effects are estimated with aggregate data (e.g., using a NMA).
-As such, we start by fitting Weibull models for the SOC.
+To imitate a realistic scenario, we assume that IPD is only available for SOC and that relative treatment effects are estimated with aggregate data (e.g., using a NMA).
+As such, we start by fitting Weibull models for SOC.
 
@@
 
-The relative treatment effects are available in terms of regression coefficients for the scale parameter on the log scale.
+The relative treatment effects are available (from a hypothetical external analysis) in terms of regression coefficients for the scale parameter on the log scale.
 The matrices \code{coef_pfs_wei} and \code{coef_os_wei} contain samples of the treatment effects for the PSA.
@@
 Since we are no longer using an individual-level model, costs cannot vary as a function of time in a health state.
-We now assume conservatively that drug costs remain at initial level when in the progression state.
+We now assume conservatively that drug costs remain at their initial level when in the progression state.
 
@@
 Survival curves for both PFS and OS are predicted using the \code{PsmCurves} class, which is setup using the input data and the parameters of the Weibull model.
-However, since we manually created the parameter object---meaning that the parameters were not generating using a formula interface---the input data must contain each term contained in the parameter object.
+However, since we manually created the parameter object---meaning that the model terms were not generated using a formula interface---the input data must contain each term contained in the parameter object.
 We must consequently create an intercept column of ones.
@@
 
-Finally, the complete partitioned survial model is constructed using the survival, utility, and cost models
+Now that the survival, utility, and cost submodels have been constructed, we can instantiate the complete PSM.
 
@@
 \includegraphics{figure/psm_plot_survival-1.pdf}
-\caption{Mean progression free survival and overall survival, weighted by weight of each patient in the target population.
-Solid and dotted lines represent means and 95\% confidence intervals from the probabilistic sensitivity analysis, respectively.
-weighted} \label{fig:psm-survival}
+\caption{Mean progression free survival and overall survival, weighted by the weight given to each patient in the target population.
+Solid and dotted lines represent means and 95\% confidence intervals from the probabilistic sensitivity analysis, respectively.} \label{fig:psm-survival}
 \end{figure}
@@
 
-QALYs and costs are simulated in using the same \code{sim_qalys()} and \code{sim_costs()} methods as in the previous section, as is the calculation of the ICER.
+QALYs and costs are simulated using the same \code{sim_qalys()} and \code{sim_costs()} methods as in the previous section, and the ICER is computed in the same fashion as well.
 
@@
 One related application is the quality-adjusted time without symptoms or toxicity (Q-TWiST) method \citep{goldhirsch1989costs, lenderking1994eval}, which partitions patients into time spent (i) having toxic side effects, (ii) without symptoms or toxicities, and (iii) after progression.
-An example 4-state model is provided on the package website.
+An example four-state model is provided on the package website.
 

### 2020-12-24 manuscript.Rnw
Edits to cDTSTM section of illustrative example

@@
 We obtain the parameters of the fitted Weibull regression models with the generic function \code{create_params()}, and then manually add the relative treatment effects to the coefficients for the scale parameters.
-(Note that if we had not fit a model for SOC, a survival parameter objects could have still been constructed with \code{params_surv()}.)
+(Note that if we had not fit a model for SOC, the survival parameter objects could have still been constructed with \code{params_surv()}.)
 
@@
 The third model type supported by \pkg{hesim} is the cDTSTM.
-Although it is commonly used in health economics, the approaches taken in Section~\ref{sec:ictstm-example} and \ref{sec:psm-example} make better use of multi-state data if the time to event outcomes are continuously observed.
-However, in some cases panel data is available where observations are only collected at arbitrary times.
+Although it is commonly used in health economics, the approaches taken in Section~\ref{sec:ictstm-example} and \ref{sec:psm-example} make better use of multi-state data if the time-to-event outcomes are continuously observed.
+Unfortunately, observations are sometimes only collected at arbitrary times, which results in what is referred to as panel data.
 This may, for instance, occur in routine practice oncology data when progression of disease is not continuously evaluated.
@@
 \subsubsection{Parameterization}
-As in Section~\ref{sec:psm-example}, we consider a case where patient level data is available for SOC but aggregate level data is used to estimate the relative treatment effects.
+As in Section~\ref{sec:psm-example}, we consider a case where IPD is available for SOC but aggregate-level data is used to estimate the relative treatment effects.
 When exact transition times are not observed, the survival models estimated in Section~\ref{sec:ictstm-example} are inappropriate.
@@
 
-In the prior two sections, we constructed the disease model using parameter (i.e., \code{params_surv}) objects.
+In Sections~\ref{sec:ictstm-example} and \ref{sec:psm-example}, we constructed the disease model using parameter (i.e., \code{params_surv}) objects.
 Here, we will use a transformed parameter (i.e., \code{tparams_transprobs}) object, which is a very flexible way to construct the transition probability matrices for a cDTSTM.
@@
 We start by predicting transition intensity matrices for SOC from the exponential multi-state model using the same input data used for the survival models in the PSM example.
-A prediction is made for each representative patient and samples of the transition intensity matrices are drawn for the PSA using the asymptotic normal approximation.
+A prediction is made for each representative patient and samples of the transition intensity matrices are drawn for the PSA using an asymptotic normal approximation.
 The array contains $B \times N_p$ matrices where $B$ is the number of PSA samples and $N_p$ is the number of patients.
@@
 transmod_data <- survmods_data
-qmat_soc <- qmatrix(msm_fit, newdata = transmod_data[strategy_name == "SOC"],
+qmat_soc <- qmatrix(msm_fit, 
+                    newdata = transmod_data[strategy_name == "SOC"],
                     uncertainty = "normal", n = n_samples)
@@
 In particular, let $x$ be a $N \times k$ matrix where $k$ is the number of predictors and $\beta$ be a $B \times k$ matrix where each row contains a given sample for the PSA.
-The function \code{xbeta} produces a $N \times B$ matrix of predicted relative risks and then ``flattens'' the matrix to create a single vector of length $B \times N$.
+The function \code{xbeta} performs the matrix multiplication operation $x\beta^T$ to produce a $N \times B$ matrix of predicted relative risks and then ``flattens'' the matrix to create a single vector of length $N \times B$.
 
@@
 
-We use this function to predict distributions of relative risks (for the transitions stable $\rightarrow$ progression and stable $\rightarrow$ death) as a function of the three treatment strategies.
+We use this function to predict distributions of relative risks (for the transitions stable $\rightarrow$ progression and stable $\rightarrow$ death) for each row of the input data as a function of the three treatment strategies.
 
@@
 \subsubsection{Simulation}
-One the model has been parameterized, the simulation is straightforward. The transition model depends only on the transition probability matrices and the cycle length,
+Once the model has been parameterized, the simulation is straightforward. The transition model depends only on the transition probability matrices and the cycle length,
 
 <<dtstm_transmod>>=
-transmod <- CohortDtstmTrans$new(params = tprobs, cycle_length = cycle_len)
+transmod <- CohortDtstmTrans$new(params = tprobs, 
+                                 cycle_length = cycle_len)
 @

### 2020-12-24 manuscript.Rnw
Updates to CEA section

@@
 where $e_{j}$ and $c_{j}$ are measures of health outcomes (e.g.
-QALYs) and costs using treatment $j$ respectively, and $k$ is a decision makers willingness to pay (WTP) per unit of health outcomes.
-The optimal treatment is the one that maximizes the expected NMB.
+QALYs) and costs using treatment $j$ respectively, and $k$ is a decision makers willingness to pay (WTP) per unit of a health outcome.
+The optimal treatment is the one that maximizes the expected NMB,
 
@@
 Treatments can be compared in an equivalent manner using the incremental cost-effectiveness ratio (ICER).
-The most common case occurs when a new treatment is more effective and more costly so that treatment $1$ is preferred to treatment $0$ if the ICER is greater than the willingness to pay threshold $k$,
+The most common case occurs when a new treatment is more effective and more costly so that treatment $1$ is preferred to treatment $0$ if the ICER is less than than the willingness to pay threshold $k$,
 
@@
 
-\subsection{Example}
+\subsection{Application}
 Uncertainty over $\theta$ is captured using the PSA.
-The simulated distribution of QALYs and costs for each treatment strategy can be produced with the \code{$summarize()} method.
-Such output was generated above for the iCTSTM.
+The simulated distribution of QALYs and costs for each treatment strategy can be produced with the \code{$summarize()} method as was demonstrated in Section~\ref{sec:illustrative-example}.
+We will use the output from the iCTSTM in this example.
 
@@
 A CEA is performed from this output with the \code{cea()} and \code{cea_pw()} functions.
-\code{cea()} summarizes results by simultaneously accounting for each treatment strategy in the analysis, while \code{cea_pw()} (as used above) summarizes “pairwise” results in which each treatment is compared to a comparator.
+\code{cea()} summarizes results by simultaneously accounting for each treatment strategy in the analysis, while \code{cea_pw()} summarizes ``pairwise'' results in which each treatment is compared to a comparator.
 
@@
 
-A number of common measures can be used to summarize a CEA and represent decision uncertainty.
-\pkg{hesim} provides support for a number of them including the cost-effectiveness plane, cost-effectiveness acceptability curve (CEAC), cost-effectiveness acceptability frontier (CEAF), and expected value of perfect information (EVPI).
-
-\subsubsection{Cost-effectiveness plane}
+\subsubsection{Incremental cost-effectiveness ratio}
+The \code{cea_pw()} function computes the ICER and a summary table can be produced with \code{icer_tbl()}.
+The argument \code{k} is the WTP threshold and in this example the new interventions are labeled as``cost-effective'' relative to SOC if the ICER is below this threshold.
+Incremental QALYs and incremental costs are computed by averaging over PSA samples.
+
+<<cea_icer>>=
+icer_tbl(cea_pw_ictstm, colnames = strategies$strategy_name,
+         k = 100000)
+@
+
+\subsubsection{Representing decision uncertainty}
+A number of common measures can be used to represent decision uncertainty from a CEA, including cost-effectiveness planes, cost-effectiveness acceptability curves (CEAC), cost-effectiveness acceptability frontiers (CEAF), and the expected value of perfect information (EVPI).
+Example plots of each from the code that follows are shown in Figure~\ref{fig:cea-plots}.
+
+\begin{figure}[t!]
+\centering
+% Cost-effectiveness plane
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics{figure/ce_plane-1.pdf}
+\caption{Cost-effectiveness plane} \label{subfig:ce-plane}
+\end{subfigure}
+% CEAC (simultaneous)
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics{figure/ceac_simultaneous-1.pdf}
+\caption{CEAC (simultaneous)} \label{subfig:ceac-simultaneous}
+\end{subfigure}
+\par\bigskip % Add some vertical space for second row
+% CEAC (pairwise)
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics{figure/ceac_pw-1.pdf}
+\caption{CEAC (pairwise)} \label{subfig:ceac-pw}
+\end{subfigure}
+% CEAF
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics{figure/ceaf-1.pdf}
+\caption{CEAF} \label{subfig:ceaf}
+\end{subfigure}
+\par\bigskip % Add some vertical space for third row
+% EVPI
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics{figure/evpi-1.pdf}
+\caption{EVPI} \label{subfig:evpi}
+\end{subfigure}
+\caption{Representations of decision uncertainty}\label{fig:cea-plots}
+\end{figure}
+
 The cost-effectiveness plane plots the incremental effectiveness of a treatment strategy (relative to a comparator) against the incremental cost of the treatment strategy.
@@
 
-<<ce_plane, fig.height = 4, fig.width = 6, fig.show="hide">>=
+<<ce_plane, fig.height = 3, fig.width = 4.5, fig.show="hide">>=
 library("scales")
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics{figure/ce_plane-1.pdf}
-\caption{Cost-effectiveness plane} \label{fig:ce-plane}
-\end{figure}
-
-\subsubsection{Cost-effectiveness acceptability curve}
 A useful summary measure for quantifying uncertainty is the CEAC, which can be generated based on simultaneous comparisons of all treatment strategies or by comparing each treatment strategy to a single comparator.
@@
 
-<<ceac_simultaneous, fig.height = 4, fig.width = 6, fig.show = "hide">>=
+<<ceac_simultaneous, fig.height = 3, fig.width = 4.5, fig.show = "hide">>=
 ggplot(cea_ictstm$mce,
@@
 
-<<ceac_pw, fig.height = 4, fig.width = 6, fig.show = "hide">>=
+<<ceac_pw, fig.height = 3, fig.width = 4.5, fig.show = "hide">>=
 ggplot(cea_pw_ictstm$ceac,
@@
 
-\begin{figure}[t!]
-\centering
-\begin{subfigure}[b]{.8\textwidth}
-\includegraphics{figure/ceac_simultaneous-1.pdf}
-\caption{Simultaneous comparison} \label{fig:ceac-simultaneous}
-\end{subfigure}
-\begin{subfigure}[b]{.8\textwidth}
-\includegraphics{figure/ceac_pw-1.pdf}
-\caption{Pairwise comparison} \label{fig:ceac-pw}
-\end{subfigure}
-\caption{Cost-effectiveness acceptability curves}\label{fig:ceac}
-\end{figure}
-
-\subsubsection{Cost-effectiveness acceptability frontier}
 One drawback of the CEAC is that the probability of being cost-effective cannot be used to determine the optimal treatment option.
@@
 
-<<ceaf, fig.height = 4, fig.width = 6, fig.show = "hide">>=
+<<ceaf, fig.height = 3, fig.width = 4.5, fig.show = "hide">>=
 ggplot(cea_ictstm$mce[best == 1],
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics{figure/ceaf-1.pdf}
-\caption{Cost-effectiveness acceptability frontier} \label{fig:ceaf}
-\end{figure}
-
-\subsubsection{Expected value of perfect information}
 A limitation of the prior measures are that they ignore the magnitude of QALY or cost gains.
@@
 \begin{equation}
-EVPI = E_\theta \left[max_j NMB(j, \theta)\right] - max_j E_\theta \left [ NMB(j, \theta)\right].
+EVPI = E_\theta \left[\max_j NMB(j, \theta)\right] - \max_j E_\theta \left [ NMB(j, \theta)\right].
 \end{equation}
@@
 The \code{cea()} function performs the EVPI calculation across all simulation draws from the PSA and for a number of WTP values.
-A plot of the the EVPI is shown below.
-The kink represents the WTP value where the optimal strategy changes.
-
-<<evpi, fig.height = 4, fig.width = 6, fig.show = "hide">>=
+The kink in the plot represents the WTP value where the optimal strategy changes.
+
+<<evpi, fig.height = 3, fig.width = 4.5, fig.show = "hide">>=
 ggplot(cea_ictstm$evpi, aes(x = k, y = evpi)) +
@@
 @
-
-\begin{figure}[t!]
-\centering
-\includegraphics{figure/evpi-1.pdf}
-\caption{Expected value of perfect information} \label{fig:evpi}
-\end{figure}
 

### 2020-12-24 manuscript.Rnw
Updates to discussion and conclusion

@@
 On the other hand, some users may prefer a functional approach that is arguably more human readable.
-To facilitate this type of approach, a \code{define_model()} block can be used to construct transition probability matrices using nonstandard evaluation.
+To facilitate this type of analysis, a \code{define_model()} block can be used to construct transition probability matrices using nonstandard evaluation.
 Each possible combination of the parameters and input data (treatment strategies and patients) is automatically constructed and the user can write custom expressions that create the transition probability matrices as a function of the parameters and input data.
@@
 Example that leverage multinomial logistic regression and \code{define_model()} are provided on the package website.
-The latter is used to reproduce two examples (a simple Markov model of an HIV treatment and a time-inhomogeneous Markov cohort model of total hip replacement) from the reference textbook by \citet{briggs2006decision}.
+The latter is used to reproduce two examples (a simple Markov cohort model of an HIV treatment and a time-inhomogeneous Markov cohort model of total hip replacement) from the reference textbook by \citet{briggs2006decision}.
 
@@
 A similar approach was used in the partitioned survival analysis example above (Section~\ref{sec:psm-example}).
-A re-analysis of the total hip replacement model from the \citet{briggs2006decision} textbook using an individual-levels simulation is provided on the package website to illustrate.
+A re-analysis of the total hip replacement model from the \citet{briggs2006decision} textbook using an individual-level simulation is provided on the package website to illustrate.
 
@@
 Performance comparisons are provided on the package website.
-The time-inhomogeneous total hip replacement \citet{briggs2006decision} example was run with both packages using 1,000 PSA iterations.
+The time-inhomogeneous total hip replacement example from \citet{briggs2006decision} was run with both packages using 1,000 PSA iterations.
 A cohort model in \pkg{heemod} completed in approximately 85 seconds while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \code{hesim} ran in 9 seconds.
@@
 \pkg{hesim} is a modular and computationally efficient \proglang{R} package for development and analysis of simulation models for economic evaluation.
-Discrete-time cohort and continuous-time individual-level models are supported, encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes as well as partitioned survival analysis.
+Discrete time cohort and continuous time individual-level models are supported, encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes as well as partitioned survival analysis.
 To facilitate integration of statistical methods and economic models and use of both patient- and aggregate-level data, parameters can be estimated by fitting a statistical model with \proglang{R}, inputting parameter estimates from an external source, or a combination of the two.
@@
 Algorithms are efficiently coded and written in \proglang{C++} so that approaches that have been considered computationally intensive such as individual-level simulations and PSA can run very quickly.
-The modular design makes modeling more flexible and makes it straightforward to add new features.
+The modular design makes modeling more flexible and it straightforward to add new features.
 

### 2020-12-24 manuscript.Rnw
Update given that 'intercept' columns returned by create_params.flexsurvreg() in hesim are now named 'intercept'

@@
                                         coef_os_wei)
+survmods_params$pfs$coefs$scale[1:2, ]
 @
@@
 However, since we manually created the parameter object---meaning that the model terms were not generated using a formula interface---the input data must contain each term contained in the parameter object.
-We must consequently create an intercept column of ones.
+We must consequently create a column of ones for the ``intercept``.
 
 <<psm_survmods>>=
-# To remove with update of package
-colnames(survmods_params$pfs$coefs$shape)[1] <- "intercept"
-colnames(survmods_params$pfs$coefs$scale)[1] <- "intercept"
-colnames(survmods_params$os$coefs$shape)[1] <- "intercept"
-colnames(survmods_params$os$coefs$scale)[1] <- "intercept"
-
-# To keep
 survmods_data[, intercept := 1]

### 2020-12-24 manuscript.Rnw
Update given that 'intercept' column is now named '(Intercept)'

@@
 However, since we manually created the parameter object---meaning that the model terms were not generated using a formula interface---the input data must contain each term contained in the parameter object.
-We must consequently create a column of ones for the ``intercept``.
+We must consequently create a column of ones for the intercept.
 
 <<psm_survmods>>=
-survmods_data[, intercept := 1]
+survmods_data[, ("(Intercept)") := 1]
 survmods <- create_PsmCurves(survmods_params, 

### 2020-12-24 manuscript.Rnw
Perform subgroup analysis in PSM example

@@
   patient_id = 1:4,
-  grp_id = c(1, 1, 0, 0),
+  grp_id = c(1, 1, 2, 2),
   patient_wt = rep(1/4, 4),
@@
 <<psm_icer>>=
-ce_sim_psm <- psm$summarize()
+ce_sim_psm <- psm$summarize(by_grp = TRUE)
 cea_pw_psm <- cea_pw(ce_sim_psm, comparator = 1,
                      dr_qalys = .03, dr_costs = .03)
-icer_tbl(cea_pw_psm, colnames = strategies$strategy_name,
+icer_tbl(cea_pw_psm, output = "data.table",
          k = 100000)

### 2020-12-24 manuscript.Rnw
Use partsurvfit object to create PSM survival models to allow for bootstrapping

@@
 
+One difficulty with partitioned survival analysis is that it does not account for the correlation in the survival endpoints and the curves may consequently cross during the PSA.
+One way to better account for correlations between the endpoints is to draw PSA samples via bootstrapping whereby the PFS and OS models are refit repeatedly to resamples of the estimation dataset.
+To allow for bootstrapping we create a \code{partsurvfit} object, which stores both the survival models and the estimation data.
+
 <<psm_fits_soc>>=
@@
 
-psmfit_soc_wei <- flexsurvreg_list(pfs = fit_pfs_soc_wei, os = fit_os_soc_wei)
+psmfit_soc_wei <- partsurvfit(
+  flexsurvreg_list(pfs = fit_pfs_soc_wei, os = fit_os_soc_wei),
+  data = onc_pfs_os
+)
 @
@@
 <<psm_survmods_params>>=
-survmods_params <- create_params(psmfit_soc_wei, n = n_samples)
+survmods_params <- create_params(psmfit_soc_wei, n = n_samples, 
+                                 bootstrap = FALSE)
 survmods_params$pfs$coefs$scale <- cbind(survmods_params$pfs$coefs$scale,

### 2020-12-26 manuscript.Rnw
Pass hesim_data object to create_StateVals() rather than stateval_tbl()

@@
   ),
-  dist = "beta",
-  hesim_data = hesim_dat)
+  dist = "beta")
 print(utility_tbl)
@@
   ),
-  dist = "gamma",
-  hesim_data = hesim_dat)
+  dist = "gamma")
 print(medcost_tbl)
@@
   drugcost_dt,
-  dist = "fixed",
-  hesim_data = hesim_dat
-)
+  dist = "fixed")
 print(drugcost_tbl)
@@
 # Utility
-utilitymod <- create_StateVals(utility_tbl, n = n_samples)
+utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
+                               hesim_data = hesim_dat)
 
@@
 drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples,
-                                time_reset = TRUE)
-medcostmod <- create_StateVals(medcost_tbl, n = n_samples)
+                                time_reset = TRUE, hesim_data = hesim_dat)
+medcostmod <- create_StateVals(medcost_tbl, n = n_samples,
+                               hesim_data = hesim_dat)
 costmods <- list(Drug = drugcostmod,
@@
 <<psm_fits_soc>>=
+onc_pfs_os_soc <- onc_pfs_os[strategy_name == "SOC"]
 fit_pfs_soc_wei <- flexsurv::flexsurvreg(
   Surv(pfs_time, pfs_status) ~ female,
-  data = onc_pfs_os, subset = (strategy_name == "SOC"),
+  data = onc_pfs_os_soc,
   dist = "weibull")
@@
   Surv(os_time, os_status) ~  female,
-  data = onc_pfs_os, subset = (strategy_name == "SOC"),
+  data = onc_pfs_os_soc,
   dist = "weibull")
@@
   flexsurvreg_list(pfs = fit_pfs_soc_wei, os = fit_os_soc_wei),
-  data = onc_pfs_os
+  data = onc_pfs_os_soc
 )
@@
 
-<<psm_stateval_tbl>>=
-attr(utility_tbl, "patients") <- hesim_dat$patients
-attr(medcost_tbl, "patients") <- hesim_dat$patients
-@
-
 <<psm_drugcost>>=
@@
   drugcost_dt[time_start ==  0][, time_start := NULL],
-  dist = "fixed",
-  hesim_data = hesim_dat)
+  dist = "fixed")
 print(drugcost_tbl)
@@
 <<psm_stvalmods>>=
-utilitymod <- create_StateVals(utility_tbl, n = n_samples)
-drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples)
-medcostmod <- create_StateVals(medcost_tbl, n = n_samples)
+utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
+                               hesim_data = hesim_dat)
+drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples, 
+                                hesim_data = hesim_dat)
+medcostmod <- create_StateVals(medcost_tbl, n = n_samples, 
+                               hesim_data = hesim_dat)
 costmods <- list(Drug = drugcostmod, Medical = medcostmod)

### 2020-12-28 manuscript.Rnw
Minor update to discussion adding potential extnesion to automatically create tunnel states
Minor edits to use uncertainty argument when drawing parameters for PSA rather than deprecated point_estimate and/or bootstrap

@@
 The transition model is then constructed as a function of the parameters (from the fitted Weibull models) and the input data.
-We must also specify the allowed transitions, the desired number of PSA samples (drawn from the multivariate normal distribution of the maximum likelihood estimate of the regression coefficients), the "clock" (recall that we used a clock reset approach when fitting the Weibull models), and the starting age of the patients.
-Note that the starting age of the patients is useful because it allows us to specify a maximum age for patients during the simulation, ensuring that the simulated patients do not live unrealistically long lives.
+We must also specify the allowed transitions, the desired number of PSA samples, the "clock" (recall that we used a clock reset approach when fitting the Weibull models), and the starting age of the patients.
+Parameters for the PSA are, by default, drawn from the multivariate normal distribution of the maximum likelihood estimate of the regression coefficients, although we make this explicit with the \code{uncertainty} argument.
+The starting age of the patients is useful because it allows us to specify a maximum age for patients during the simulation, ensuring that the simulated patients do not live unrealistically long lives.
 
@@
                                    trans_mat = tmat, n = n_samples,
+                                   uncertainty = "normal",
                                    clock = "reset",
@@
 survmods_params <- create_params(psmfit_soc_wei, n = n_samples, 
-                                 bootstrap = FALSE)
+                                 uncertainty = "normal")
 survmods_params$pfs$coefs$scale <- cbind(survmods_params$pfs$coefs$scale,
@@
 Third, more complex models of utility \citep{kharroubi2007modelling, hernandez2013relationship} and costs \citep{nixon2005methods} could help better capture heterogeneity in preferences and spending.
+Fourth, it would be useful to add functionality that automatically creates tunnel states for users in cDTSTMs.
 Finally, a larger aspiration is to provide support for parameterization via NMA, particularly for estimation of relative treatment effects in survival (e.g., for PSMs) and multi-state models.

### 2020-12-28 manuscript.Rnw
Updates using new hesim summary.ce() method

@@
 They are summarized with the \code{$summarize()} method, which first sums across health states and then computes a (weighted) average across patients, by PSA sample, treatment strategy, and subgroup.
+The output is a generic cost-effectiveness object (of class \code{ce}) that can be used to perform CEA as demonstrated in Section~\ref{sec:cea}.
+Here, we use the \code{summary()} method to summarize the results.
 
@@
 ce_sim_ictstm <- ictstm$summarize()
-head(ce_sim_ictstm$costs)
-@
-
-The output is a generic cost-effectiveness object (of class \code{ce}) that can be used to perform CEA.
-We demonstrate a formal CEA in Section~\ref{sec:cea}.
-Here, we make a simple pairwise comparisons between each of the new interventions (New 2 and New 1) and the comparator (SOC), and compute the incremental cost-effectiveness ratio (ICER).
-
-<<ictstm_icer>>=
-ce_sim_ictstm <- ictstm$summarize()
-cea_pw_ictstm <- cea_pw(ce_sim_ictstm, comparator = 1,
-                        dr_qalys = .03, dr_costs = .03,
-                        k = seq(0, 25000, 500))
-icer_tbl(cea_pw_ictstm, colnames = strategies$strategy_name,
-         k = 100000)
+summary(ce_sim_ictstm)
 @
@@
 
-QALYs and costs are simulated using the same \code{sim_qalys()} and \code{sim_costs()} methods as in the previous section, and the ICER is computed in the same fashion as well.
+QALYs and costs are simulated using the same \code{sim_qalys()} and \code{sim_costs()} methods as in the previous section, and the results are summarized in the same fashion as well.
 
@@
 ce_sim_psm <- psm$summarize(by_grp = TRUE)
-cea_pw_psm <- cea_pw(ce_sim_psm, comparator = 1,
-                     dr_qalys = .03, dr_costs = .03)
-icer_tbl(cea_pw_psm, output = "data.table",
-         k = 100000)
+summary(ce_sim_psm)
 @
@@
 
-and an ICER is computed.
+and the results are summarized.
 
@@
 ce_sim_cdtstm <- cdtstm$summarize()
-cea_pw_cdtstm <- cea_pw(ce_sim_cdtstm, comparator = 1, 
-                        dr_costs = .03, dr_qalys = .03)
-icer_tbl(cea_pw_cdtstm, output = "data.table", k = 100000)
+summary(ce_sim_cdtstm)
 @

### 2020-12-29 manuscript.Rnw
Minor updates given new icer() function

@@
 \subsubsection{Simulation}
-Before constructing each submodel, we specify the number of parameters samples that will be drawn for the PSA.
+Before constructing each submodel, we specify the number of parameter samples that will be drawn for the PSA.
 
@@
 \subsubsection{Incremental cost-effectiveness ratio}
-The \code{cea_pw()} function computes the ICER and a summary table can be produced with \code{icer_tbl()}.
-The argument \code{k} is the WTP threshold and in this example the new interventions are labeled as``cost-effective'' relative to SOC if the ICER is below this threshold.
-Incremental QALYs and incremental costs are computed by averaging over PSA samples.
+The \code{cea_pw()} function computes the ICER.
+A summary table can be produced with \code{icer()} and \code{format()} can format the table for pretty printing.
+The argument \code{k} is the WTP threshold and is used to compute the INMB.
+Estimates of incremental QALYs and incremental costs are computed by averaging over PSA samples.
+By default, 95\% confidence intervals are presented and are computed using quantiles from the PSA.
 
 <<cea_icer>>=
-icer_tbl(cea_pw_ictstm, colnames = strategies$strategy_name,
-         k = 100000)
+library("magrittr")
+icer(cea_pw_ictstm, k = 100000, 
+     strategy_values = strategies$strategy_name[-1]) %>%
+  format()
 @

### 2020-12-30 manuscript.Rnw
Minor updaters with summary.ce() and icer() functions

@@
 The output is a generic cost-effectiveness object (of class \code{ce}) that can be used to perform CEA as demonstrated in Section~\ref{sec:cea}.
-Here, we use the \code{summary()} method to summarize the results.
+Here, we use the \code{summary()} method to summarize the results and make use the ``pipes'' from the \pkg{magrittr} to nicely format the output.
 
 <<ictstm_summarize>>=
+library("magrittr")
 ce_sim_ictstm <- ictstm$summarize()
-summary(ce_sim_ictstm)
+summary(ce_sim_ictstm, 
+        strategy_values = strategies$strategy_name) %>%
+  format()
 @
@@
 ce_sim_psm <- psm$summarize(by_grp = TRUE)
-summary(ce_sim_psm)
+summary(ce_sim_psm, 
+        strategy_values = strategies$strategy_name) %>%
+  format()
 @
@@
 ce_sim_cdtstm <- cdtstm$summarize()
-summary(ce_sim_cdtstm)
+summary(ce_sim_cdtstm, 
+        strategy_values = strategies$strategy_name) %>%
+  format()
 @
@@
 A CEA is performed from this output with the \code{cea()} and \code{cea_pw()} functions.
-\code{cea()} summarizes results by simultaneously accounting for each treatment strategy in the analysis, while \code{cea_pw()} summarizes ``pairwise'' results in which each treatment is compared to a comparator.
+\code{cea()} summarizes results by simultaneously accounting for each treatment strategy in the analysis, while \code{cea_pw()} summarizes ``pairwise'' results in which each treatment is compared to a comparator (in this case SOC).
 
@@
 <<cea_icer>>=
-library("magrittr")
 icer(cea_pw_ictstm, k = 100000, 

### 2021-01-02 manuscript.Rnw
Use new plotting functions to create CEA plots

@@
 
+<<labels_indiv>>=
+labs_indiv <- get_labels(hesim_dat, grp_label = NULL)
+@
+
 \subsubsection{Parameterization}
@@
 A number of common measures can be used to represent decision uncertainty from a CEA, including cost-effectiveness planes, cost-effectiveness acceptability curves (CEAC), cost-effectiveness acceptability frontiers (CEAF), and the expected value of perfect information (EVPI).
-Example plots of each from the code that follows are shown in Figure~\ref{fig:cea-plots}.
+Plots can be quickly generated with \code{plot_ceplane()}, \code{plot_ceac()}, \code{plot_ceaf()}, and \code{plot_evpi()} functions, respectively.
+Example plots from the code that follows are shown in Figure~\ref{fig:cea-plots}
 
@@
 <<ce_plane, fig.height = 3, fig.width = 4.5, fig.show="hide">>=
-library("scales")
-strategy_factor <- function (x) {
-  factor(x, levels = 1:3, labels = c("SOC", "New 1", "New 2"))
-}
-
-xlim <- ceiling(max(cea_pw_ictstm$delta[, ie]) * 1.2)
-ylim <- max(cea_pw_ictstm$delta[, ic]) * 1.2
-ggplot(cea_pw_ictstm$delta,
-       aes(x = ie, y = ic, col = strategy_factor(strategy_id))) +
-  geom_jitter(size = .5)  +
-  xlab("Incremental QALYs") +
-  ylab("Incremental cost") +
-  scale_y_continuous(limits = c(-ylim, ylim),
-                     labels = scales::dollar) +
-  scale_x_continuous(limits = c(-xlim, xlim), breaks = seq(-2, 2, .5)) +
-  theme(legend.position = "bottom") +
-  scale_colour_discrete(name = "Strategy") +
-  geom_abline(slope = 100000, linetype = "dashed") +
-  geom_hline(yintercept = 0) +
-  geom_vline(xintercept = 0)
-@
-
-A useful summary measure for quantifying uncertainty is the CEAC, which can be generated based on simultaneous comparisons of all treatment strategies or by comparing each treatment strategy to a single comparator.
+plot_ceplane(cea_pw_ictstm, labels = labs_indiv)
+@
+
+The CEAC can be generated based on simultaneous comparisons of all treatment strategies or by comparing each treatment strategy to a single comparator.
 The former is the probability that each strategy is the most cost-effective and is available from the \code{mce} element produced by \code{cea()}.
-
-<<ceac_simultaneous, fig.height = 3, fig.width = 4.5, fig.show = "hide">>=
-ggplot(cea_ictstm$mce,
-       aes(x = k, y = prob, col = strategy_factor(strategy_id))) +
-  geom_line() +
-  xlab("Willingness to pay") +
-  ylab("Probability most cost-effective") +
-  scale_x_continuous(breaks = seq(0, max(wtp), length.out = 6),
-                     label = scales::dollar) +
-  scale_colour_discrete(name = "Strategy") +
-  theme(legend.position = "bottom")
-@
-
 The relevant information for the latter is available from the \code{ceac} element produced by \code{cea_pw()}, and is the probability that each treatment is more cost-effective than the comparator.
 
-<<ceac_pw, fig.height = 3, fig.width = 4.5, fig.show = "hide">>=
-ggplot(cea_pw_ictstm$ceac,
-        aes(x = k, y = prob, col = strategy_factor(strategy_id))) +
-  geom_line() +
-  xlab("Willingness to pay") +
-  ylab("Probability most cost-effective") +
-  scale_x_continuous(breaks = seq(0, max(wtp), length.out = 6),
-                     label = scales::dollar) +
-  theme(legend.position = "bottom") +
-  scale_colour_discrete(name = "Strategy")
+<<ceac, fig.height = 3, fig.width = 4.5, fig.show = "hide">>=
+plot_ceac(cea_ictstm, labels = labs_indiv)
+plot_ceac(cea_pw_ictstm, labels = labs_indiv)
 @
@@
 <<ceaf, fig.height = 3, fig.width = 4.5, fig.show = "hide">>=
-ggplot(cea_ictstm$mce[best == 1],
-       aes(x = k, y = prob, col = strategy_factor(strategy_id))) +
-  geom_line() +
-  xlab("Willingness to pay") +
-  ylab("Probability most cost-effective") +
-  scale_x_continuous(breaks = seq(0, max(wtp), length.out = 6),
-                     label = scales::dollar) +
-  theme(legend.position = "bottom") +
-  scale_colour_discrete(name = "Strategy")
+plot_ceaf(cea_ictstm, labels = labs_indiv)
 @
@@
 <<evpi, fig.height = 3, fig.width = 4.5, fig.show = "hide">>=
-ggplot(cea_ictstm$evpi, aes(x = k, y = evpi)) +
-  geom_line()  +
-  xlab("Willingness to pay") +
-  ylab("Expected value of perfect information") +
-  scale_x_continuous(breaks = seq(0, max(wtp), length.out = 6),
-                     label = scales::dollar) +
-  scale_y_continuous(label = scales::dollar) +
-  theme(legend.position = "bottom")
+plot_evpi(cea_ictstm, labels = labs_indiv)
 @

### 2021-01-04 manuscript.Rnw
Minor plot tweaks
Incorporate new hesim labels functionality

@@
 <<labels_indiv>>=
-labs_indiv <- get_labels(hesim_dat, grp_label = NULL)
+labs_indiv <- get_labels(hesim_dat)
+print(labs_indiv)
 @
@@
 ce_sim_ictstm <- ictstm$summarize()
-summary(ce_sim_ictstm, 
-        strategy_values = strategies$strategy_name) %>%
+summary(ce_sim_ictstm, labels = labs_indiv) %>%
   format()
@@
   age = c(55, 65, 55, 65),
-  female = c(1, 1, 0, 0)
+  female = c(1, 1, 0, 0),
+  grp_name = rep(c("Female", "Male"), 
+                 each = 2)
 )
+@
+
+<<labels_cohort>>=
+labs_cohort <- get_labels(hesim_dat)
 @
@@
 ce_sim_psm <- psm$summarize(by_grp = TRUE)
-summary(ce_sim_psm, 
-        strategy_values = strategies$strategy_name) %>%
+summary(ce_sim_psm, labels = labs_cohort) %>%
   format()
@@
 ce_sim_cdtstm <- cdtstm$summarize()
-summary(ce_sim_cdtstm, 
-        strategy_values = strategies$strategy_name) %>%
+summary(ce_sim_cdtstm, labels = labs_cohort) %>%
   format()
@@
 <<cea_icer>>=
-icer(cea_pw_ictstm, k = 100000, 
-     strategy_values = strategies$strategy_name[-1]) %>%
+icer(cea_pw_ictstm, k = 100000, labels = labs_indiv) %>%
   format()
@@
 % Cost-effectiveness plane
-\begin{subfigure}[b]{.48\textwidth}
+\begin{subfigure}[b]{.49\textwidth}
 \centering
@@
 % CEAC (simultaneous)
-\begin{subfigure}[b]{.48\textwidth}
+\begin{subfigure}[b]{.49\textwidth}
 \centering
-\includegraphics{figure/ceac_simultaneous-1.pdf}
+\includegraphics{figure/ceac-1.pdf}
 \caption{CEAC (simultaneous)} \label{subfig:ceac-simultaneous}
@@
 % CEAC (pairwise)
-\begin{subfigure}[b]{.48\textwidth}
+\begin{subfigure}[b]{.49\textwidth}
 \centering
-\includegraphics{figure/ceac_pw-1.pdf}
+\includegraphics{figure/ceac-2.pdf}
 \caption{CEAC (pairwise)} \label{subfig:ceac-pw}
@@
 % CEAF
-\begin{subfigure}[b]{.48\textwidth}
+\begin{subfigure}[b]{.49\textwidth}
 \centering
@@
 % EVPI
-\begin{subfigure}[b]{.48\textwidth}
+\begin{subfigure}[b]{.49\textwidth}
 \centering

### 2021-01-05 manuscript.Rnw
Manuscript updates

@@
 
-Analyses were performed using \pkg{hesim} version \code{0.5.0}.
+Analyses were performed using \pkg{hesim} version \code{0.5.0}. All code including a replication \proglang{R} script for this paper is available at \url{https://github.com/hesim-dev/hesim-manuscript}.
 
@@
 
-This model can be simulated using either of the model types described in Section~\ref{sec:model-taxonomy}. We demonstrate use of all three below, starting with the iCTSTM.
+This model can be simulated using any of the model types described in Section~\ref{sec:model-taxonomy}. We demonstrate use of all three below, starting with the iCTSTM.
 
@@
 
-The target population, treatment strategies, and (non-death) health states are stored in a \code{hesim_data} object.
+The target population, treatment strategies, and (non-death) health states are stored in a \code{hesim_data} object and identified with integer valued identification variables (\code{patient_id}, \code{strategy_id}, and \code{state_id}).
 There are 3 treatment strategies: standard of care (SOC) and two new interventions (New 1 and New 2).
 The target population consists of a heterogeneous population of 1,000 patients, which is large enough to ensure that averages across patients in the individual simulation are reasonably stable.
-As we will see in the next section, patients can also be placed into subgroups with a \code{grp_id} column in the \code{patients} table; if left unspecified as done here, then all patients are assumed to belong to a single subgroup.
+As we will see in the next section, patients can also be placed into subgroups (with a \code{grp_id} column) or given weights (with a \code{patient_wt} column) in the \code{patients} table; if left unspecified as done here, then all patients are assumed to belong to a single subgroup and are weighted equally.
 
@@
 @
+
+Labels for the identification variables in a \code{hesim_data} object can be generated with \code{get_labels()}.
+These can (as will be shown below), in turn, be passed to plotting and summary functions to create tables and figures with more informative labels.
 
@@
 Here, we use the \code{summary()} method to summarize the results and make use the ``pipes'' from the \pkg{magrittr} to nicely format the output.
+Labels for the treatment strategies are based on output of \code{get_labels()} produced above.
 
@@
 
+Since there are now subgroups, we generate new labels that include the subgroup variables.
+
 <<labels_cohort>>=
@@
 
-We first simulate survival curves by month for 20 years.
+We first simulate survival curves by month for 30 years.
 State probabilities are easily computed from these survival curves with the \code{sim_stateprobs()} method.
@@
 <<psm_sim_survival>>=
-times <- seq(0, 20, by = .1)
+times <- seq(0, 30, by = .1)
 psm$sim_survival(t = times)
@@
 QALYs and costs are simulated using the same \code{sim_qalys()} and \code{sim_costs()} methods as in the previous section, and the results are summarized in the same fashion as well.
+A difference, however, is that we now use the \code{by_grp = TRUE} option to summarize costs and QALYs by subgroup (rather than aggregating across all groups).
 
@@
 Here, we will use a transformed parameter (i.e., \code{tparams_transprobs}) object, which is a very flexible way to construct the transition probability matrices for a cDTSTM.
-A \code{tparams_transprobs} object requires a 3-dimensional array of transition matrices, sorted by PSA sample, treatment strategy, patient, and optionally (in a time-inhomogeneous model) time interval.
+A \code{tparams_transprobs} object requires a 3-dimensional array of transition matrices, where the transition matrices are ordered by PSA sample, treatment strategy, patient, and optionally (in a time-inhomogeneous model) time interval.
 
@@
 
-Identifiers for each matrix in the array can then be generated with \code{tpmatrix_id()}.
+Identifiers for the complete array of transitions matrices (inclusive of SOC and the new interventions) are generated with \code{tpmatrix_id()}.
 
@@
             xbeta(x_rr, params_rr_rng$rr_13))
+head(rr) # Corresponds to 1st 6 rows in tpmat_id
 @
@@
 
-and the results are summarized.
+and the results are summarized. The \code{by_grp} argument of \code{$summarize()} is by default \code{FALSE} so costs and QALYs are aggregated across subgroups here.
 
@@
 The \code{cea_pw()} function computes the ICER.
-A summary table can be produced with \code{icer()} and \code{format()} can format the table for pretty printing.
+A summary table is produced with \code{icer()} and \code{format()} formats the table for pretty printing.
 The argument \code{k} is the WTP threshold and is used to compute the INMB.
@@
 A number of common measures can be used to represent decision uncertainty from a CEA, including cost-effectiveness planes, cost-effectiveness acceptability curves (CEAC), cost-effectiveness acceptability frontiers (CEAF), and the expected value of perfect information (EVPI).
-Plots can be quickly generated with \code{plot_ceplane()}, \code{plot_ceac()}, \code{plot_ceaf()}, and \code{plot_evpi()} functions, respectively.
-Example plots from the code that follows are shown in Figure~\ref{fig:cea-plots}
+Plots using \pkg{ggplot2} graphics can be quickly generated with the \code{plot_ceplane()}, \code{plot_ceac()}, \code{plot_ceaf()}, and \code{plot_evpi()} functions, respectively \citep{wickham2016ggplot2}.
+Output from \code{cea()} and \code{cea_pw()} is readily available, so users may create custom plots as well.
+Example figures from the code that follows are shown in Figure~\ref{fig:cea-plots}.
 
@@
 The CEAF, which plots the probability that the optimal treatment strategy (i.e., the strategy with the highest expected NMB) is cost-effective, is appropriate in this context.
-
 A CEAF curve can be created by using the \code{best} column from the \code{mce} element to subset to the treatment strategy with the highest expected NMB for each WTP value.
@@
 The same \code{define_model()} block may also be used to construct utility and cost models, if desired.
-Example that leverage multinomial logistic regression and \code{define_model()} are provided on the package website.
+Examples that leverage multinomial logistic regression and \code{define_model()} are provided on the package website.
 The latter is used to reproduce two examples (a simple Markov cohort model of an HIV treatment and a time-inhomogeneous Markov cohort model of total hip replacement) from the reference textbook by \citet{briggs2006decision}.
@@
 
-\pkg{mstate} is another package for multi-state modeling that is designed for estimation of non-parametric and semi-parametric (i.e., Cox proportional hazards) models.
+\pkg{mstate} is another package for multi-state modeling that is designed for estimation of non-parametric and semi-parametric (i.e., Cox proportional hazards) models \citep{de2011mstate}.
 It supports prediction of state probabilities and simulation of multi-state trajectories (to facilitate prediction for clock-reset models) from cumulative hazard functions evaluated at specific times.
 \citet{williams2017cost} has adapted \pkg{mstate} for cost-effectiveness modeling applications.
-There are however, some disadvantages of using \pkg{mstate} for health-economic modeling since it was not designed for it.
+There are however, some disadvantages of using \pkg{mstate} for health economic modeling since it was not designed for it.
 \code{mssample()}---the core function for simulating multi-state models---does not natively support heterogeneous target populations, PSA, simultaneous simulation of multiple treatment strategies, or computation of QALYs and costs from the simulated output.
-It is also computationally inefficient for two main reasons.
+It is less computationally efficient than \pkg{hesim} for two main reasons.
 First, simulation of general cumulative hazards functions is considerably slower than simulation of parametric distributions where efficient random number generators are available.
@@
 Algorithms are efficiently coded and written in \proglang{C++} so that approaches that have been considered computationally intensive such as individual-level simulations and PSA can run very quickly.
-The modular design makes modeling more flexible and it straightforward to add new features.
+The modular design makes modeling more flexible and it easier to add new features.
 

### 2021-01-06 manuscript.Rnw
Minor text edits

@@
 Supported model types include cohort discrete time state transition models (cDTSTMs), individual continuous time state transition models (iCTSTMs), and partitioned survival models (PSMs), encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
-A modular design based on \pkg{R6} and \code{S3} classes allows users to combine separate submodels for disease progression, costs, and quality-adjusted life years (QALYs) in a flexible way.
+A modular design based on \pkg{R6} and \code{S3} classes allows users to combine separate submodels for disease progression, costs, and utility in a flexible way.
 Probabilistic sensitivity analysis (PSA) is used to propagate uncertainty in model parameters to model outputs.
 Simulation code is written in \proglang{C++} so complex simulations such as those combining PSA and individual simulation can be run much more quickly than previously possible.
-Decision analysis within a cost-effectiveness framework is performed using simulated costs and QALYs from a PSA.
+Decision analysis within a cost-effectiveness framework is performed using simulated costs and quality-adjusted life years (QALYs) from a PSA.
 }
@@
 It is therefore no surprise that a number of related \proglang{R} packages have recently been developed, such as \pkg{BCEA} \citep{baio2017bayesian}, \pkg{SAVI} \citep{strong2014estimating}, \pkg{survHE} \citep{baio2020survhe}, and \pkg{heemod} \citep{filipovic2017markov}.
-Still, of the available packages, only \pkg{heemod} provides a general purpose framework for developing simulation models and it is limited to cohort Markov models.
+Still, of the available packages, only \pkg{heemod} provides a general purpose framework for developing simulation models and it is limited Markov cohort models.
 

### 2021-01-11 manuscript.Rnw
Typo fix
Minor language tweaks

@@
 Multiple model types are supported including cohort discrete time state transition models (cDTSTMs), N-state PSMs, and individual-level continuous time state transition models (iCTSTMs), encompassing both Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
-To maximize flexibility and facilitate integration of the statistical methods and economic model, parameters can be estimated either by fitting a model in \proglang{R} or by inputting parameters obtained from external sources.
+To maximize flexibility and facilitate integration of the statistical methods and economic model, parameters can be set by either fitting models in \proglang{R} or by inputting values obtained from external sources.
 So that individual-level simulation and PSA can be run quickly, \pkg{Rcpp} and \pkg{data.table} are heavily utilized.
@@
 \hline
-Predicted means & \code{tparams_mean} & \code{stateval_tbl::lm}\\
+Predicted means & \code{tparams_mean} & \code{stateval_tbl}\\
 Linear model & \code{params_lm} & \code{stats::lm}\\
@@
 Discrete time cohort and continuous time individual-level models are supported, encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes as well as partitioned survival analysis.
-To facilitate integration of statistical methods and economic models and use of both patient- and aggregate-level data, parameters can be estimated by fitting a statistical model with \proglang{R}, inputting parameter estimates from an external source, or a combination of the two.
+To facilitate integration of statistical methods and economic models and use of both patient- and aggregate-level data, models can be parameterized by either fitting statistical models using \proglang{R}, inputting values directly, or from a combination of the two.
 Simulated costs and QALYs from a PSA can be used for for decision-analysis within a cost-effectiveness framework, allowing users to represent decision uncertainty (e.g., cost-effectiveness planes, cost-effectiveness acceptability curves, and cost-effectiveness acceptability frontiers), and conduct value of information analysis.

### 2021-01-11 manuscript.Rnw
Typo fix

@@
 The latter is useful for storing the parameters of a multi-state model or the independent survival models required for a PSM.
-The parameters of a multinomial logistic regression are stored in a \code{params_mlogit} and can be created by fitting a model for each row in a transition probability matrix with \code{nnet::multinom()}.
+The parameters of a multinomial logistic regression are stored in a \code{params_mlogit} object and can be created by fitting a model for each row in a transition probability matrix with \code{nnet::multinom()}.
 

### 2021-01-24 manuscript.Rnw
Incorporate hesim autoplot methods

@@
 
-The most straightforward way to construct state values is with a \code{stateval_tbl}, which is a special object used to assign values (i.e.
+The most straightforward way to construct state values is with \code{stateval_tbl}, which is a special object used to assign values (i.e.
 predicted means) to health states that can vary across PSA samples, treatment strategies, patients, and/or time intervals.
@@
 
-<<ictstm_summarize_stateprobs, echo=FALSE, results="hide">>=
-add_state_name <- function(x){
-  x[, state_name := factor(state_id,
-                           levels = 1:nrow(tmat),
-                           labels = colnames(tmat))]
-}
-
-add_strategy_name <- function(x) {
-  x[, strategy_name := factor(strategy_id, 
-                              labels = c("SOC", "New 1", "New 2"))]
-}
-
-summarize_stprobs <- function(stateprobs){
-  x <- stateprobs[, .(prob_mean = mean(prob)),
-                  by = c("strategy_id", "state_id", "t")]
-  add_state_name(x)
-  add_strategy_name(x)
-  return(x[, ])
-}
-@
-
-<<ictstm_plot_stateprobs, echo = FALSE, include = FALSE>>=
+\code{autoplot()} methods are available to quickly visualize simulation output using \pkg{ggplot2} graphics \citep{wickham2016ggplot2}.
+State probabilities from such a plot are shown in Figure~\ref{fig:ictstm-stprobs}, with labels for the treatment strategies and health states based on the output of \code{get_labels()} produced above.
+
+<<ictstm_plot_stateprobs, fig.show = "hide">>=
 library("ggplot2")
-stprobs <- summarize_stprobs(ictstm$stateprobs_)
-ggplot(stprobs, 
-       aes(x = t, y = prob_mean, col = strategy_name)) +
-  geom_line() + facet_wrap(~state_name) +
-  xlab("Years") + ylab("Probability in health state") +
-  scale_color_discrete(name = "Strategy") +
-  theme(legend.position = "bottom") 
-@
-
-\begin{figure}[t!]
+autoplot(ictstm$stateprobs_, labels = labs_indiv,
+         ci = FALSE) 
+@
+
+\begin{figure}[h]
 \centering
@@
 Here, we use the \code{summary()} method to summarize the results and make use the ``pipes'' from the \pkg{magrittr} to nicely format the output.
-Labels for the treatment strategies are based on output of \code{get_labels()} produced above.
 
@@
 
-We first simulate survival curves by month for 30 years.
-State probabilities are easily computed from these survival curves with the \code{sim_stateprobs()} method.
-The area under the PFS curve is the proportion of patients in stable disease while the area between the PFS and OS curves is the proportion of patient whose disease has progressed.
+We first simulate survival curves by month for 30 years. Survival curves produced using an \code{autoplot()} method are displayed in Figure~\ref{fig:psm-survival}.
 
@@
 
-<<psm_plot_survival, echo = FALSE, include = FALSE>>=
-# Survival curve plotting data
-surv_means <- psm$survival_[, lapply(.SD, weighted.mean, w = patient_wt),
-                                by = c("strategy_id", "curve", "t"),
-                               .SDcols = "survival"]
-surv_mean_min <- surv_means[curve != 2][, curve := curve + 1]
-setnames(surv_mean_min, "survival", "survival_min")
-surv_means <- merge(surv_means, surv_mean_min,
-                    by = c("strategy_id", "curve", "t"),
-                    all.x = TRUE)
-surv_means[, survival_min := ifelse(curve == 1, 0, survival_min)]
-surv_means[, curve := factor(curve, labels = c("PFS", "OS"))]
-surv_means[, strategy_name := factor(strategy_id, labels = c("SOC", "New 1", "New 2"))]
-
-# Plot
-ggplot(surv_means, aes(x = t, y = survival, fill = curve)) +
-  geom_line(aes(col = curve)) +
-  facet_wrap(~strategy_name) +
-  geom_ribbon(aes(ymin = survival_min, ymax = survival), alpha = .5) +
-  xlab("Years") + ylab("Proportion surviving") +
-  scale_color_discrete(name = "Survival curve") +
-  guides(fill = FALSE) +
-  theme_bw() +
-  theme(legend.position = "bottom") 
-@
-
-\begin{figure}[t!]
+<<psm_plot_survival, fig.show = "hide">>=
+autoplot(psm$survival_, 
+         labels = c(labs_cohort,
+                    list(curve = c("PFS" = 1, "OS" = 2))
+                    ),
+         ci = TRUE, ci_style = "ribbon")
+@
+
+\begin{figure}[h]
 \centering
@@
 \caption{Mean progression free survival and overall survival, weighted by the weight given to each patient in the target population.
-Solid and dotted lines represent means and 95\% confidence intervals from the probabilistic sensitivity analysis, respectively.} \label{fig:psm-survival}
+The solid line and shaded regions represent means and 95\% confidence intervals from the probabilistic sensitivity analysis, respectively.} \label{fig:psm-survival}
 \end{figure}
+
+State probabilities are easily computed from these survival curves with the \code{$sim_stateprobs()} method.
+The area under the PFS curve is the proportion of patients in stable disease while the area between the PFS and OS curves is the proportion of patient whose disease has progressed.
 
@@
 
-<<cdtstm_plot_stateprobs, echo = FALSE, include = FALSE>>=
-stprobs_tot <- cdtstm$stateprobs_[, lapply(.SD, weighted.mean, w = patient_wt),
-                                   by = c("sample", "strategy_id", "state_id", "t"),
-                                   .SDcols = "prob"]
-stprobs_summary <- summarize_stprobs(stprobs_tot)
-ggplot(stprobs_summary, aes(x = t, y = prob_mean)) +
-  geom_line(aes(col = strategy_name)) +
-  # geom_ribbon(aes(x = t, ymin = prob_lower, ymax = prob_upper,
-  #                 fill = strategy_name), alpha = .3) +
-  facet_wrap(~state_name) +
-  xlab("Year") + ylab("Probability") +
-  scale_fill_discrete("Strategy") + scale_color_discrete("Strategy")
-@
-
 QALYs and costs are again simulated using a 3 percent discount rate
@@
 A number of common measures can be used to represent decision uncertainty from a CEA, including cost-effectiveness planes, cost-effectiveness acceptability curves (CEAC), cost-effectiveness acceptability frontiers (CEAF), and the expected value of perfect information (EVPI).
-Plots using \pkg{ggplot2} graphics can be quickly generated with the \code{plot_ceplane()}, \code{plot_ceac()}, \code{plot_ceaf()}, and \code{plot_evpi()} functions, respectively \citep{wickham2016ggplot2}.
+Plots using \pkg{ggplot2} can be quickly generated with the \code{plot_ceplane()}, \code{plot_ceac()}, \code{plot_ceaf()}, and \code{plot_evpi()} functions, respectively.
 Output from \code{cea()} and \code{cea_pw()} is readily available, so users may create custom plots as well.

### 2021-02-12 manuscript.Rnw
Typo fixes

@@
 It is therefore no surprise that a number of related \proglang{R} packages have recently been developed, such as \pkg{BCEA} \citep{baio2017bayesian}, \pkg{SAVI} \citep{strong2014estimating}, \pkg{survHE} \citep{baio2020survhe}, and \pkg{heemod} \citep{filipovic2017markov}.
-Still, of the available packages, only \pkg{heemod} provides a general purpose framework for developing simulation models and it is limited Markov cohort models.
+Still, of the available packages, only \pkg{heemod} provides a general purpose framework for developing simulation models and it is limited to Markov cohort models.
 
@@
 The remainder of this paper is organized as follows.
-Section~\ref{sec:model-taxonomy} provides a mathematical description the economic models supported by \pkg{hesim}.
+Section~\ref{sec:model-taxonomy} provides a mathematical description of the economic models supported by \pkg{hesim}.
 An overview of the coding framework is provided in Section~\ref{sec:framework}.
@@
 While the algorithm repeats until a patient dies, it can also be stopped at a specified time $t$ or when a patient reaches a maximum age.
-In the latter scenario, death is assume to occur at the maximum age.
+In the latter scenario, death is assumed to occur at the maximum age.
 
@@
 Model development proceeds in 4 steps.
-The first step is to setup the model by defining the model structure, target population, and treatment strategies of interest.
+The first step is to set up the model by defining the model structure, target population, and treatment strategies of interest.
 The relevant information is then stored in a \code{hesim_data} object.
@@
 Next, the submodels are combined to construct an economic model.
-The economic model is then used to simulate, disease progression, QALYs, and costs as a function of ``input data''.
+The economic model is then used to simulate disease progression, QALYs, and costs as a function of ``input data''.
 The input data always contains variables describing the target population and treatment strategies of interest, but can also contain variables related to time to facilitate use of time-varying covariates or parameters.
@@
 There are two types of parameter objects, standard parameter objects prefixed by \code{``params''} and ``transformed'' parameter objects prefixed by \code{``tparams''}.
-The former contain the underlying parameters of a statistical model while the latter contain parameters more immediate to prediction that have been transformed as function of the input data.
-The regression coefficients of a logistic regression are an example of a parameter objects while the predicted probabilities are examples of a transformed parameter object.
+The former contain the underlying parameters of a statistical model while the latter contain parameters more immediate to prediction that have been transformed as a function of the input data.
+The regression coefficients of a logistic regression are an example of parameter objects while the predicted probabilities are examples of a transformed parameter object.
 
@@
 Methods for simulating disease progression differ slightly since the simulation approaches differ as described in Section~\ref{sec:model-taxonomy}.
-In an N-state PSM $n-1$ survival curves are generated and in a iCTSTM a trajectory through the STM is simulated for multiple patients via individual simulation.
+In an N-state PSM $n-1$ survival curves are generated and in an iCTSTM a trajectory through the STM is simulated for multiple patients via individual simulation.
 All models can simulate state probabilities over time.
@@
 \subsection{Individual continuous time state transition model} \label{sec:ictstm-example}
-We setup the model by defining the model structure, target population, and treatment strategies of interest.
+We set up the model by defining the model structure, target population, and treatment strategies of interest.
 The transitions are characterized by a matrix where each element denotes a transition from a row to a column.
@@
 Multi-state models can be fit in one of two ways.
-First, a joint survival model with interaction terms for different transition can be estimated, which is useful if there are constraints in the parameters, such as coefficients that are assumed equal across transitions.
+First, a joint survival model with interaction terms for different transitions can be estimated, which is useful if there are constraints in the parameters, such as coefficients that are assumed equal across transitions.
 Here, we will take a second approach that is computationally more efficient and fit a separate model for each transition.
@@
 
-The complete economic model is an constructed by combining the transition, utility, and cost submodels.
+The complete economic model is constructed by combining the transition, utility, and cost submodels.
 
@@
 The output is a generic cost-effectiveness object (of class \code{ce}) that can be used to perform CEA as demonstrated in Section~\ref{sec:cea}.
-Here, we use the \code{summary()} method to summarize the results and make use the ``pipes'' from the \pkg{magrittr} to nicely format the output.
+Here, we use the \code{summary()} method to summarize the results and make use of the ``pipes'' from the \pkg{magrittr} to nicely format the output.
 
@@
 \subsubsection{Parameterization}
-A multi-state dataset can be converted into dataset with one row per patient and outcomes for PFS and OS using the function \code{as_pfs_os()}.
+A multi-state dataset can be converted into a dataset with one row per patient and outcomes for PFS and OS using the function \code{as_pfs_os()}.
 
@@
 \subsubsection{Simulation}
-As with with the multi-state model, input data for the survival models is constructed with the \code{expand()} function.
+As with  the multi-state model, input data for the survival models is constructed with the \code{expand()} function.
 
@@
 State probabilities are easily computed from these survival curves with the \code{$sim_stateprobs()} method.
-The area under the PFS curve is the proportion of patients in stable disease while the area between the PFS and OS curves is the proportion of patient whose disease has progressed.
+The area under the PFS curve is the proportion of patients in stable disease while the area between the PFS and OS curves is the proportion of patients whose disease has progressed.
 
@@
 When exact transition times are not observed, the survival models estimated in Section~\ref{sec:ictstm-example} are inappropriate.
-The \pkg{msm} package, can, however, be used to estimate models from such data.
+The \pkg{msm} package can, however, be used to estimate models from such data.
 The loss of information does come at a price as only exponential and piecewise exponential models are possible.
@@
 
-A limitation of the prior measures are that they ignore the magnitude of QALY or cost gains.
+A limitation of the prior measures is that they ignore the magnitude of QALY or cost gains.
 A measure which combines the probability of being most effective with the magnitude of the expected NMB is the EVPI.
@@
 On the other hand, some users may prefer a functional approach that is arguably more human readable.
-To facilitate this type of analysis, a \code{define_model()} block can be used to construct transition probability matrices using nonstandard evaluation.
-Each possible combination of the parameters and input data (treatment strategies and patients) is automatically constructed and the user can write custom expressions that create the transition probability matrices as a function of the parameters and input data.
+To facilitate this type of analysis, a \code{define_model()} block can be used to construct transition probability matrices using non-standard evaluation.
+Each possible combination of the parameters and input data (treatment strategies and patients) is automatically constructed and the user can define expressions that create the transition probability matrices as a function of the parameters and input data.
 The same \code{define_model()} block may also be used to construct utility and cost models, if desired.

### 2021-02-13 manuscript.Rnw
Some minor tweaks

--- 2021-01-11 manuscript.Rnw
@@
 
-  Jeroen Jansen\\
+  Jeroen P Jansen\\
   University of California, San Francisco\\
@@
 Publicly funded health systems such as those in Australia, Canada, and the United Kingdom, among others, use HTA to help maximize health gains for the population given a fixed budget.
-HTA is also commonly used by private payers to guide coverage decisions and the adoption of new technologies \citep{trosman2011health}.
+HTA is also used by private payers to guide coverage decisions and the adoption of new technologies \citep{trosman2011health}.
 Such assessments usually rely heavily on cost-effectiveness analysis (CEA) so that decisions can be made on the basis of a formal evaluation of costs and effects \citep{dakin2015influence}.
-CEA is made feasible by development of health economic models that simulate costs and effects over relevant time horizons using the totality of available evidence.
+CEA is made feasible by development of health economic models that simulate costs and effects over relevant time horizons for the compared health technologies using the totality of relevant evidence for the decision problem at hand.
 
@@
 
-Expected values are computed by integrating the ``weighted'' probability of being in each state, where weights are a function of the discount factor and state values.
+Expected values are computed by integrating the probability of being in each state and ``weighted'' state values, where weights are a function of the discount factor.
 That is, for a time horizon $T$, discounted costs and QALYs in health state h are computed as,
@@
 iCTSTMs simulate individual trajectories between health states using random number generation.
-Trajectories are simulated for multiple patients and costs and QALYs are computed by averaging across the simulated patients.
+Trajectories are simulated for multiple patients and expected costs and QALYs are computed by averaging across the simulated patients.
 A reasonably large number of patients must be simulated to ensure that expected values are stable \citep{ohagan2007monte}.
@@
 
-Survival functions are estimated by fitting parametric or flexible parametric survival models as described in Section~\ref{sec:iCTSTMs}.
+Survival functions are estimated by fitting (flexible) parametric survival models as described in Section~\ref{sec:iCTSTMs}.
 The $n$th fitted survival model with density $f_n(t)$ has cumulative density function $F_n(t)$, survivor function $1 - F_n(T)$, cumulative hazard $H_n(t) = -\log S_n(t)$, and hazard $h_n(t) = f_n(t)/S_n(t)$.
@@
 \section{Framework} \label{sec:framework}
-Economic models consist of a disease model, a utility model, and a set of cost models for each cost category.
+Economic models with \pkg{hesim} consist of a disease model, a utility model, and a set of cost models for each cost category.
 Model development proceeds in 4 steps.
@@
 While other approaches such as multi-criteria decision analysis (MCDA) could, in principle, be used as well, only CEA is currently supported.
-All models are simulated using PSA so that decision uncertainty can be represented.
+All models are simulated using PSA so that parameter uncertainty can be represented.
 
@@
 
-The three economic models described in Section~\ref{sec:model-taxonomy} are implemented as the \pkg{R6} classes \code{CohortDtstm}, \code{IndivCtstm}, and \code{Psm}.
+The three types of economic models described in Section~\ref{sec:model-taxonomy} are implemented as the \pkg{R6} classes \code{CohortDtstm}, \code{IndivCtstm}, and \code{Psm}.
 Each class contains methods for simulating disease progression, QALYs, and costs.
@@
 The utility and cost models are always \pkg{R6} objects of class \code{StateVals}, whereas the disease models vary by economic model (Table~\ref{tbl:submodels}).
-The disease model is used to simulate survival curves in a PSM and health state transitions in a cDTSTM and iCTSTM.
+The disease model is used to simulate health state transitions in a cDTSTM and iCTSTM, and survival curves in a PSM.
 
@@
 \code{CohortDtstm} & \code{CohortDtstmTrans} & \code{StateVals} & \code{StateVals}\\
+\code{IndivCtstm} & \code{IndivCtstmTrans} & \code{StateVals} & \code{StateVals}\\
 \code{Psm} & \code{PsmCurves} & \code{StateVals} & \code{StateVals}\\
-\code{IndivCtstm} & \code{IndivCtstmTrans} & \code{StateVals} & \code{StateVals}\\
 \hline
@@
 Methods for simulating disease progression differ slightly since the simulation approaches differ as described in Section~\ref{sec:model-taxonomy}.
-In an N-state PSM $n-1$ survival curves are generated and in a iCTSTM a trajectory through the STM is simulated for multiple patients via individual simulation.
+In a iCTSTM a trajectory through the STM is simulated for multiple patients via individual simulation and in an N-state PSM $n-1$ survival curves are generated and .
 All models can simulate state probabilities over time.
@@
 \code{CohortDtstm} & \code{sim_stateprobs()} & \code{sim_qalys()} & \code{sim_costs()}\\
+\code{IndivCtstm} & \code{sim_disease()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
 \code{Psm} & \code{sim_survival()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
-\code{IndivCtstm} & \code{sim_disease()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
 \hline
@@
 
-The complete economic model is an constructed by combining the transition, utility, and cost submodels.
+The complete economic model is a constructed by combining the transition, utility, and cost submodels.
 

### 2021-02-14 manuscript.Rnw
Update manuscript.Rnw
Increase number of PSA iterations and use bootstrap for partitioned survival model

--- 2021-02-12 manuscript.Rnw
@@
 One source of decision uncertainty stems from uncertainty in the underlying model parameters.
-Parameter uncertainty is typically quantified using probabilistic sensitivity analysis (PSA), which involves randomly sampling the model parameters from suitable probability distribution and simulating the model for each sampled parameter set \citep{claxton2005probabilistic}.
+Parameter uncertainty is typically quantified using probabilistic sensitivity analysis (PSA), which involves randomly sampling the model parameters from suitable probability distributions and simulating the model for each sampled parameter set \citep{claxton2005probabilistic}.
 When combined with individual-level simulation, PSA can take an appreciable amount of time to run \citep{ohagan2007monte}.
@@
 cDTSTMs simulate the probability that a cohort of patients is in each of $H$ health states over time.
-Time is partitioned into intervals $\{[0, t_1), [t_1, t_2), \ldots, [t_C, t_{C+1})\}$ with each interval $0,\ldots, C$ reffered to as a model cycle.
+Time is partitioned into intervals $\{[0, t_1), [t_1, t_2), \ldots, [t_C, t_{C+1})\}$ with each interval $0,\ldots, C$ referred to as a model cycle.
 We denote $t_c$ as the time at the start of model cycle $c$.
@@
 <<ictstm_n_samples>>=
-n_samples <- 100
+n_samples <- 1000
 @
@@
 
-<<psm_survmods_params>>=
+<<psm_survmods_params, warning = FALSE>>=
 survmods_params <- create_params(psmfit_soc_wei, n = n_samples, 
-                                 uncertainty = "normal")
+                                 uncertainty = "bootstrap", 
+                                 max_errors = 5, silent = TRUE)
 survmods_params$pfs$coefs$scale <- cbind(survmods_params$pfs$coefs$scale,

### 2021-02-15 manuscript.Rnw
Some text updates

@@
 
-This model can be simulated using any of the model types described in Section~\ref{sec:model-taxonomy}. We demonstrate use of all three below, starting with the iCTSTM.
+This model can be simulated using any of the model types described in Section~\ref{sec:model-taxonomy}.
+We demonstrate use of all three below, noting how the type of model employed should depend on the available data.
+We begin with the richest dataset in Section~\ref{sec:ictstm-example} and show how a flexible iCTSTM can be parameterized when continuously observed patient-level multi-state data is available.
+The next two sections then show how models can be developed when the data is more limited.
+Section~\ref{sec:psm-example} uses a PSM in a case where PFS and OS are available but multi-state data is not, and Section~\ref{sec:cdtstm-example} develops a cDTSTM in a setting where multi-state data is available but it is not continuously observed.
+For the sake of illustration, in both Sections~\ref{sec:psm-example} and ~\ref{sec:cdtstm-example}, patient-level data is available for the SOC, but not for estimation of relative treatment effects.
 
@@
 
-Since we did not fit a survival regression model containing both the absolute and relative treatment effects, we must construct the survival parameter object manually.
+Since we did not fit a survival regression model containing both the baseline risk and relative treatment effects, the survival parameter object must be constructed manually.
 We obtain the parameters of the fitted Weibull regression models with the generic function \code{create_params()}, and then manually add the relative treatment effects to the coefficients for the scale parameters.
 (Note that if we had not fit a model for SOC, the survival parameter objects could have still been constructed with \code{params_surv()}.)
+
+The parameters for SOC are drawn by bootstrapping.
+Although this better accounts for the correlation between PFS and OS, it is slower than drawing the parameters from a multivariate normal distribution (as employed in Section~\ref{sec:ictstm-example}) because the survival models are refit for each PSA sample.
+We use the \code{max_errors} argument to allow for errors during some of the bootstrap replications (e.g., if a model fails to converge).
 
@@
 Fourth, it would be useful to add functionality that automatically creates tunnel states for users in cDTSTMs.
+Fifth, equity considerations could be incorporated by adding support for distributional CEA \citep{asaria2015distributional, asaria2016distributional, cookson2017using}.
 Finally, a larger aspiration is to provide support for parameterization via NMA, particularly for estimation of relative treatment effects in survival (e.g., for PSMs) and multi-state models.

### 2021-02-16 manuscript.Rnw
Add Jeroen's middle initial
Add health economic R packages to references

@@
 \author{Devin Incerti\\Genentech
-   \And Jeroen Jansen\\University of California, San Francisco}
+   \And Jeroen P Jansen\\University of California, San Francisco}
 \Plainauthor{Devin Incerti, Jeroen Jansen}
@@
 The limitations of such software relative to programming languages like \proglang{R} have been increasingly emphasized in the literature \citep{baio2017simple, incerti2019r, jalal2017overview}.
-It is therefore no surprise that a number of related \proglang{R} packages have recently been developed, such as \pkg{BCEA} \citep{baio2017bayesian}, \pkg{SAVI} \citep{strong2014estimating}, \pkg{survHE} \citep{baio2020survhe}, and \pkg{heemod} \citep{filipovic2017markov}.
+It is therefore no surprise that a number of related \proglang{R} packages have recently been developed, such as \pkg{BCEA} \citep{baio2017bayesian}, \pkg{SAVI} \citep{strong2014estimating}, \pkg{dampack} \citep{alarid2018dampack}, \pkg{EVSI} \citep{heath2018efficient}, \pkg{survHE} \citep{baio2020survhe}, and \pkg{heemod} \citep{filipovic2017markov}.
 Still, of the available packages, only \pkg{heemod} provides a general purpose framework for developing simulation models and it is limited to Markov cohort models.

### 2021-02-16 manuscript.Rnw
Some text edits

@@
 cDTSTMs simulate the probability that a cohort of patients is in each of $H$ health states over time.
-Time is partitioned into intervals $\{[0, t_1), [t_1, t_2), \ldots, [t_C, t_{C+1})\}$ with each interval $0,\ldots, C$ referred to as a model cycle.
+Time is partitioned into intervals $\{[0, t_1], [t_1, t_2], \ldots, [t_C, t_{C+1}]\}$ with each interval $0,\ldots, C$ referred to as a model cycle.
 We denote $t_c$ as the time at the start of model cycle $c$.
@@
 Three methods can be used to estimate $f(t_c^{*})$.
-First, a left Riemann sum uses values at the start of each time interval $f(t_c^{*}) = f(t_c)$.
-Second, a right Riemann sum uses values at the end of each time interval $f(t_c^{*}) = f(t_{c+1})$.
-Finally, the trapezoid rule averages values at the start and end of each interval $f(t_c^{*}) = \frac{1}{2}\Delta t_c[f(t_c) + f(t_{c+1})]$.
+First, a left Riemann sum uses values at the start of each time interval, $f(t_c^{*}) = f(t_c)$.
+Second, a right Riemann sum uses values at the end of each time interval, $f(t_c^{*}) = f(t_{c+1})$.
+Finally, the trapezoid rule averages values at the start and end of each interval, $f(t_c^{*}) = \frac{1}{2}\Delta t_c[f(t_c) + f(t_{c+1})]$.
 
@@
 Costs and QALYs are computed using the continuous time present value given a flow of state values, which change as patients transition between health states or as costs vary as a function of time.
-The state values can be partitioned into $M$ time intervals indexed by $m = 1,\ldots, M$ where interval $m$ contains times $t$ such that $t_m\leq t \leq t_{m+1}$ and values for state $h$ are equal to $z_{hm}$ during interval $m$.
+The state values can be partitioned into $M$ time intervals indexed by $m = 1,\ldots, M$ where interval $m$ contains times $t$ such that $t_m < t \leq t_{m+1}$ and values for state $h$ are equal to $z_{hm}$ during interval $m$.
 $z_{hm}$ will equal zero during time intervals in which a patient is not in state $h$.
@@
 They differ, however, in that state probabilities are not computed via matrix multiplication or individual simulation, but from a set of non-mutually exclusive survival curves \citep{glasziou1990quality, woods2018nice}.
-Each survival curve represents time to transitioning to that state or to a more severe health state.
+Each survival curve represents time to transition to that state or to a more severe health state.
 
@@
 State membership in health states $2,\ldots, N-1$ is computed as $S_{n}(t) - S_{n-1}(t)$.
-Finally, the probability of being in the final health state $n$ (i.e., the death state) is $1-S_{N-1}(t)$, or one minus overall survival function.
+Finally, the probability of being in the final health state (i.e., the death state) is $1-S_{N-1}(t)$, or one minus overall survival function.
 
@@
 
-The most straightforward way to construct state values is with \code{stateval_tbl}, which is a special object used to assign values (i.e.
-predicted means) to health states that can vary across PSA samples, treatment strategies, patients, and/or time intervals.
+The most straightforward way to construct state values is with \code{stateval_tbl}, which is a special object used to assign values (i.e., predicted means) to health states that can vary across PSA samples, treatment strategies, patients, and/or time intervals.
 State values can be specified either as moments (e.g., mean and standard error) or parameters (e.g., shape and scale of gamma distribution) of a probability distribution, or by pre-simulating values from a suitable probability distribution (e.g., from a Bayesian model).
@@
 Methods for simulating disease progression differ slightly since the simulation approaches differ as described in Section~\ref{sec:model-taxonomy}.
-In an N-state PSM $n-1$ survival curves are generated and in an iCTSTM a trajectory through the STM is simulated for multiple patients via individual simulation.
+In an N-state PSM $N-1$ survival curves are generated and in an iCTSTM a trajectory through the STM is simulated for multiple patients via individual simulation.
 All models can simulate state probabilities over time.
@@
 We consider a three-state model commonly used in oncology.
-As shown in Figure~\ref{fig:three-state-model}, the three health states are stable disease (i.e.
-not progressed), progression, and death.
+As shown in Figure~\ref{fig:three-state-model}, the three health states are stable disease (i.e., not progressed), progression, and death.
 Patients can transition to a more severe health state (stable $\rightarrow$ progression, stable $\rightarrow$ death, and progression $\rightarrow$ death) but cannot recover to a less severe health state.
@@
 This model can be simulated using any of the model types described in Section~\ref{sec:model-taxonomy}.
-We demonstrate use of all three below, noting how the type of model employed should depend on the available data.
+We demonstrate use of all three below, noting that the type of model employed should depend on the available data.
 We begin with the richest dataset in Section~\ref{sec:ictstm-example} and show how a flexible iCTSTM can be parameterized when continuously observed patient-level multi-state data is available.
@@
 Section~\ref{sec:psm-example} uses a PSM in a case where PFS and OS are available but multi-state data is not, and Section~\ref{sec:cdtstm-example} develops a cDTSTM in a setting where multi-state data is available but it is not continuously observed.
-For the sake of illustration, in both Sections~\ref{sec:psm-example} and ~\ref{sec:cdtstm-example}, patient-level data is available for the SOC, but not for estimation of relative treatment effects.
+For the sake of illustration, in both Sections~\ref{sec:psm-example} and~\ref{sec:cdtstm-example}, patient-level data is available for the SOC, but not for estimation of relative treatment effects.
 
@@
 
-strategies <- data.frame(
+strategies <- data.table(
   strategy_id = 1:3,
@@
 for (i in 1:length(wei_fits)){
-  if (i == 3) f <- update.formula(f, .~.-factor(strategy_name)) 
+  if (i == 3) f <- update(f, .~.-factor(strategy_name)) 
   wei_fits[[i]] <- flexsurvreg(f, data = onc3, 
@@
 Medical costs are assumed to follow a gamma distribution which is often appropriate for costs since they tend to be right skewed.
-Like utility, the distribution is characterized by the mean (\$2,000 with stable disease, \$9,500 with progressed disease) and standard error (assumed equal to the mean) and the method of moments is used to derive the underlying shape and scale parameters.
+Like utility, the distribution is characterized by the mean (\$2,000 with stable disease, \$9,500 with progressed disease) and SE (assumed equal to the mean) and the method of moments is used to derive the underlying shape and scale parameters.
 
@@
 
-The input data for the transition model consists of one row for each treatment strategy and patient and can be easily generated from a \code{hesim_data} object with the \code{expand} function.
+The input data for the transition model consists of one row for each treatment strategy and patient and can be easily generated from a \code{hesim_data} object with the \code{expand()} function.
 
@@
 
-QALYs and costs are simulated from the disease progression simulation output using the utility and cost submodels with the \code{$sim_qalys()} and \code{$sim_costs()} methods.
+QALYs and costs are simulated from the disease progression simulation output using the utility and cost models with the \code{$sim_qalys()} and \code{$sim_costs()} methods.
 Outcomes are simulated for each discount rate specified with the \code{dr} argument.
@@
 
-We first simulate survival curves by month for 30 years. Survival curves produced using an \code{autoplot()} method are displayed in Figure~\ref{fig:psm-survival}.
+We first simulate survival curves by month for 30 years. Survival curves produced using \code{autoplot()} are displayed in Figure~\ref{fig:psm-survival}.
 
@@
 The third model type supported by \pkg{hesim} is the cDTSTM.
-Although it is commonly used in health economics, the approaches taken in Section~\ref{sec:ictstm-example} and \ref{sec:psm-example} make better use of multi-state data if the time-to-event outcomes are continuously observed.
+Although it is commonly used in health economics, the approaches taken in Section~\ref{sec:ictstm-example} and \ref{sec:psm-example} make better use of the data if the time-to-event outcomes are continuously observed.
 Unfortunately, observations are sometimes only collected at arbitrary times, which results in what is referred to as panel data.
@@
 Relative risks can be predicted for each row of the input data and each sample of the coefficients for the PSA using matrix multiplication.
-In particular, let $x$ be a $N \times k$ matrix where $k$ is the number of predictors and $\beta$ be a $B \times k$ matrix where each row contains a given sample for the PSA.
+In particular, let $x$ be a $N \times k$ matrix where $N$ is the number of rows of the input data and $k$ is the number of predictors, and $\beta$ be a $B \times k$ matrix where each row contains a given sample for the PSA.
 The function \code{xbeta} performs the matrix multiplication operation $x\beta^T$ to produce a $N \times B$ matrix of predicted relative risks and then ``flattens'' the matrix to create a single vector of length $N \times B$.
@@
 
-where $e_{j}$ and $c_{j}$ are measures of health outcomes (e.g.
-QALYs) and costs using treatment $j$ respectively, and $k$ is a decision makers willingness to pay (WTP) per unit of a health outcome.
+where $e_{j}$ and $c_{j}$ are measures of health outcomes (e.g., QALYs) and costs using treatment $j$ respectively, and $k$ is a decision makers willingness to pay (WTP) per unit of a health outcome.
 The optimal treatment is the one that maximizes the expected NMB,
@@
 \subsubsection{Representing decision uncertainty}
-A number of common measures can be used to represent decision uncertainty from a CEA, including cost-effectiveness planes, cost-effectiveness acceptability curves (CEAC), cost-effectiveness acceptability frontiers (CEAF), and the expected value of perfect information (EVPI).
+A number of common measures can be used to represent decision uncertainty from a CEA, including cost-effectiveness planes, cost-effectiveness acceptability curves (CEACs), cost-effectiveness acceptability frontiers (CEAFs), and the expected value of perfect information (EVPI).
 Plots using \pkg{ggplot2} can be quickly generated with the \code{plot_ceplane()}, \code{plot_ceac()}, \code{plot_ceaf()}, and \code{plot_evpi()} functions, respectively.
@@
 
-A limitation of the prior measures is that they ignore the magnitude of QALY or cost gains.
+A limitation of CEACs and CEAFs is that they ignore the magnitude of QALY or cost gains.
 A measure which combines the probability of being most effective with the magnitude of the expected NMB is the EVPI.
@@
 However, since \pkg{hesim} aims to be flexible, the example only covers a subset of the relevant modeling approaches and functionality.
-Additional examples as well as documentation for all functions and classes are consequently provided on the package website (\url{https://hesim-dev.github.io/hesim/}).
+Additional examples as well as documentation for all functions and classes are provided on the package website (\url{https://hesim-dev.github.io/hesim/}).
 
@@
 Alternative approaches might be appropriate in other contexts.
-For example, if evenly spaced panel data is available, a multinomial logistic regressions can be used to predict transition probabilities.
+For example, if evenly spaced panel data is available, multinomial logistic regressions can be used to predict transition probabilities.
 On the other hand, some users may prefer a functional approach that is arguably more human readable.
-To facilitate this type of analysis, a \code{define_model()} block can be used to construct transition probability matrices using non-standard evaluation.
-Each possible combination of the parameters and input data (treatment strategies and patients) is automatically constructed and the user can define expressions that create the transition probability matrices as a function of the parameters and input data.
+To facilitate this type of analysis, a \code{define_model()} block can be used to define transition probability matrices in terms of expressions that are evaluated later.
+Each possible combination of the parameters and input data (treatment strategies and patients) is automatically constructed and the expressions are used to create the transition probability matrices as a function of the parameters and input data.
 The same \code{define_model()} block may also be used to construct utility and cost models, if desired.
@@
 The time-inhomogeneous total hip replacement example from \citet{briggs2006decision} was run with both packages using 1,000 PSA iterations.
-A cohort model in \pkg{heemod} completed in approximately 85 seconds while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \code{hesim} ran in 9 seconds.
+At the time of writing, a cohort model in \pkg{heemod} completed in approximately 1.9 minutes while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \code{hesim} ran in 7.4 seconds.
 
@@
 \pkg{hesim} leverages some of these for parameters estimation: \pkg{flexsurv} is used to fit parametric and flexible parametric survival and multi-state models when exact event times are available while \pkg{msm} is used to fit multi-state models when only panel data is available.
-A key difference is that \pkg{hesim} is designed specifically for health-economic modeling whereas the other packages are more general purpose statistical packages.
+A key difference is that \pkg{hesim} is designed specifically for health economic modeling whereas the other packages are more general purpose statistical packages.
 
@@
 \code{mssample()}---the core function for simulating multi-state models---does not natively support heterogeneous target populations, PSA, simultaneous simulation of multiple treatment strategies, or computation of QALYs and costs from the simulated output.
-It is less computationally efficient than \pkg{hesim} for two main reasons.
-First, simulation of general cumulative hazards functions is considerably slower than simulation of parametric distributions where efficient random number generators are available.
+It is also less computationally efficient than \pkg{hesim} for two main reasons.
+First, simulation of general cumulative hazards functions is considerably slower than simulation of parametric distributions when efficient random number generators are available.
 Second, it is written completely in \proglang{R} and incorporating PSA, multiple treatment strategies, or heterogeneous patients results in code that is not vectorized.
-To compare computational performance, we ran 100 PSA iterations where 1 treatment strategy and 1,000 homogeneous patients were simulated through a 6-state Weibull model during each iteration.
-The model completed in 0.44 second in \pkg{hesim} and took 34 minutes in \pkg{mstate}, showing that computational considerations are not just an academic concern.
+To compare computational performance, we ran 1,000 PSA iterations where 1 treatment strategy and 1,000 homogeneous patients were simulated through a 6-state Weibull model during each iteration.
+The model completed in 2.8 seconds in \pkg{hesim} and 3.1 hours in \pkg{mstate}, showing that computational considerations are not just an academic concern.
 
@@
 To facilitate integration of statistical methods and economic models and use of both patient- and aggregate-level data, models can be parameterized by either fitting statistical models using \proglang{R}, inputting values directly, or from a combination of the two.
-Simulated costs and QALYs from a PSA can be used for for decision-analysis within a cost-effectiveness framework, allowing users to represent decision uncertainty (e.g., cost-effectiveness planes, cost-effectiveness acceptability curves, and cost-effectiveness acceptability frontiers), and conduct value of information analysis.
+Simulated costs and QALYs from a PSA can be used for decision-analysis within a cost-effectiveness framework, allowing users to represent decision uncertainty (e.g., cost-effectiveness planes, cost-effectiveness acceptability curves, and cost-effectiveness acceptability frontiers), and conduct value of information analysis.
 Algorithms are efficiently coded and written in \proglang{C++} so that approaches that have been considered computationally intensive such as individual-level simulations and PSA can run very quickly.

### 2021-02-17 manuscript.Rnw
Minor fixes to use \pkg when was incorrectly using \code

@@
 
-As mentioned in the introduction, \code{heemod} is the only other general purpose \proglang{R} package for cost-effectiveness modeling.
+As mentioned in the introduction, \pkg{heemod} is the only other general purpose \proglang{R} package for cost-effectiveness modeling.
 It differs from \pkg{hesim} in that it focuses solely on cDTSTMs (i.e., Markov cohort models) and is not designed to facilitate integration of economic models with the statistical models used for parameter estimation.
-Another important difference is that \code{hesim} is considerably faster since all simulations are vectorized (i.e, performed at the \proglang{C++} level).
+Another important difference is that \pkg{hesim} is considerably faster since all simulations are vectorized (i.e, performed at the \proglang{C++} level).
 Performance comparisons are provided on the package website.
 The time-inhomogeneous total hip replacement example from \citet{briggs2006decision} was run with both packages using 1,000 PSA iterations.
-At the time of writing, a cohort model in \pkg{heemod} completed in approximately 1.9 minutes while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \code{hesim} ran in 7.4 seconds.
+At the time of writing, a cohort model in \pkg{heemod} completed in approximately 1.9 minutes while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \pkg{hesim} ran in 7.4 seconds.
 

### 2021-02-22 manuscript.Rnw
Remove extra comma in code
Add sessionInfo() to facilitate replication
Update R packages to newest versions

@@
   Surv(pfs_time, pfs_status) ~ female + new1 + new2,
-  data = onc_pfs_os,,
+  data = onc_pfs_os,
   dist = "weibull")
@@
 
-<<psm_icer>>=
+<<psm_summary>>=
 ce_sim_psm <- psm$summarize(by_grp = TRUE)
@@
 
-<<cdtstm_icer>>=
+<<cdtstm_summary>>=
 ce_sim_cdtstm <- cdtstm$summarize()
@@
 
+%% -- Session info -------------------------------------------------------------
+<<sessionInfo, include=FALSE>>=
+sessionInfo()
+@
+
 %% -- Bibliography -------------------------------------------------------------

### 2021-02-24 manuscript.Rnw
Updates to references

@@
 PSMs are conceptually similar to STMs in that they are characterized by mutually exclusive health states.
-They differ, however, in that state probabilities are not computed via matrix multiplication or individual simulation, but from a set of non-mutually exclusive survival curves \citep{glasziou1990quality, woods2018nice}.
+They differ, however, in that state probabilities are not computed via matrix multiplication or individual simulation, but from a set of non-mutually exclusive survival curves \citep{glasziou1990quality, woods2017nice}.
 Each survival curve represents time to transition to that state or to a more severe health state.

### 2021-02-25 manuscript.Rnw
Minor code styling edit

@@
 onc_pfs_os_soc <- onc_pfs_os[strategy_name == "SOC"]
-fit_pfs_soc_wei <- flexsurv::flexsurvreg(
+fit_pfs_soc_wei <- flexsurvreg(
   Surv(pfs_time, pfs_status) ~ female,

### 2021-02-27 manuscript.Rnw
Minor edit to abstract

@@
 Models have historically been developed using spreadsheet software (e.g., Microsoft \proglang{Excel}) and while use of \proglang{R} is growing, general purpose modeling software is still limited.
-\pkg{hesim} helps fill this gap by facilitating parameterization, simulation, and analysis of economic models in an integrated manner.
+\pkg{hesim} is an \proglang{R} package that helps fill this gap by facilitating parameterization, simulation, and analysis of economic models in an integrated manner.
 Supported model types include cohort discrete time state transition models (cDTSTMs), individual continuous time state transition models (iCTSTMs), and partitioned survival models (PSMs), encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.

### 2021-02-28 manuscript.Rnw
Some style updates to captions

@@
 \usepackage{algorithm}
+\captionsetup[algorithm]{labelfont=rm,labelsep=colon}
 \usepackage{multirow}
@@
 \begin{algorithm}
-\caption{Simulation of individual continuous time state transition model}
+\caption{Simulation of individual continuous time state transition model.}
 \label{alg:sim-ictstm}
@@
 \end{tikzpicture}
-\caption{Economic modeling process}\label{fig:econ-eval-process-hesim}
+\caption{Economic modeling process.}\label{fig:econ-eval-process-hesim}
 \end{figure}
@@
 \begin{table} [h]
-\caption{Parameterization of disease models}\label{tbl:parameterize-disease-model}
 \footnotesize
@@
 \end{tabular*}
+\caption{Parameterization of disease models.}\label{tbl:parameterize-disease-model}
 \end{table}
@@
 \begin{table} [h]
-\caption{Parameterization of state value models}\label{tbl:parameterize-stateval-model}
 \footnotesize
@@
 \end{tabular*}
+\caption{Parameterization of state value models.}\label{tbl:parameterize-stateval-model}
 \end{table}
@@
 \begin{table} [h]
-\caption{Submodels comprising an economic model}\label{tbl:submodels}
 \footnotesize
@@
 \end{tabular*}
+\caption{Submodels comprising an economic model.}\label{tbl:submodels}
 \end{table}
@@
 \begin{table} [h]
-\caption{Methods for simulating outcomes from economic models}\label{tbl:simulate-outcomes}
 \footnotesize
@@
 \end{tabular*}
+\caption{Methods for simulating outcomes from economic models.}\label{tbl:simulate-outcomes}
 \end{table}
@@
 \end{tikzpicture}
-\caption{Three state model of disease progression in oncology}\label{fig:three-state-model}
+\caption{Three state model of disease progression in oncology.}\label{fig:three-state-model}
 \end{figure}
@@
 \includegraphics{figure/ictstm_plot_stateprobs-1.pdf}
-\caption{Mean simulated state probabilities from the probabilistic sensitivity analysis} \label{fig:ictstm-stprobs}
+\caption{Mean simulated state probabilities from the probabilistic sensitivity analysis.} \label{fig:ictstm-stprobs}
 \end{figure}
@@
 \end{subfigure}
-\caption{Representations of decision uncertainty}\label{fig:cea-plots}
+\caption{Representations of decision uncertainty.}\label{fig:cea-plots}
 \end{figure}

### 2021-02-28 manuscript.Rnw
Remove comments from code chunks

@@
   state_id = c(1, 2),
-  state_name = c("Stable", "Progression") # Non-death health states
+  state_name = c("Stable", "Progression") 
 )
@@
 <<ictstm_stvalmods>>=
-# Utility
 utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
                                hesim_data = hesim_dat)
-
-# Costs
 drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples,
@@
             xbeta(x_rr, params_rr_rng$rr_13))
-head(rr) # Corresponds to 1st 6 rows in tpmat_id
+head(rr) 
 @
@@
 <<cea>>=
-wtp <- seq(0, 250000, 500) # Willingness to pay per QALY
+wtp <- seq(0, 250000, 500) 
 cea_ictstm <- cea(ce_sim_ictstm, dr_costs = .03, dr_qalys = .03, k = wtp)

### 2021-02-28 manuscript.Rnw
Minor edits to cDTSTM math notation

@@
 cDTSTMs simulate the probability that a cohort of patients is in each of $H$ health states over time.
-Time is partitioned into intervals $\{[0, t_1], [t_1, t_2], \ldots, [t_C, t_{C+1}]\}$ with each interval $0,\ldots, C$ referred to as a model cycle.
-We denote $t_c$ as the time at the start of model cycle $c$.
+Time is partitioned into intervals $\{[0, t_1], [t_1, t_2], \ldots, [t_{C-1}, t_C]\}$ with each interval $1,\ldots, C$ referred to as a model cycle.
+We denote $t_c$ as the time at the end of model cycle $c$.
 An $H \times 1$ state vector that stores the probability of being in each health state at cycle $c$ is written as $x_c = (x_{1c}, x_{2c}, \ldots, x_{Hc})$ where $\sum_h x_{hc} = 1$.
@@
 \begin{equation} \label{eqn:markov-sim}
-x_{c+1}^T= x_c^T P_c, \quad c = 0,\ldots, C.
+x_{c+1}^T= x_c^T P_c, \quad c = 0,\ldots, C-1.
 \end{equation}
@@
 \begin{equation} \label{eqn:riemann-sum}
-\sum_{c = 0}^{C}f(t_c^{*})\Delta t_c,
+\sum_{c = 1}^{C}f(t_c^{*})\Delta t_c,
 \end{equation}
 
-where $\Delta t_c = t_{c+1} - t_{c}$,  $t_c^{*}\in[t_{c},t_{c+1}]$, and $f(t)= z_h(t) e^{-rt} p_h(t)$.
+where $\Delta t_c = t_{c} - t_{c-1}$,  $t_c^{*}\in[t_{c-1},t_{c}]$, and $f(t)= z_h(t) e^{-rt} p_h(t)$.
 Three methods can be used to estimate $f(t_c^{*})$.
-First, a left Riemann sum uses values at the start of each time interval, $f(t_c^{*}) = f(t_c)$.
-Second, a right Riemann sum uses values at the end of each time interval, $f(t_c^{*}) = f(t_{c+1})$.
-Finally, the trapezoid rule averages values at the start and end of each interval, $f(t_c^{*}) = \frac{1}{2}\Delta t_c[f(t_c) + f(t_{c+1})]$.
+First, a left Riemann sum uses values at the start of each time interval, $f(t_c^{*}) = f(t_{c-1})$.
+Second, a right Riemann sum uses values at the end of each time interval, $f(t_c^{*}) = f(t_{c})$.
+Finally, the trapezoid rule averages values at the start and end of each interval, $f(t_c^{*}) = \frac{1}{2}\Delta t_c[f(t_{c-1}) + f(t_{c})]$.
 

### 2021-03-07 replication.R
Commit replication R file

@@ -1,0 +1,496 @@
+#' ---
+#' title: "hesim: Health Economic Simulation Modeling and Decision Analysis"
+#' author: "Devin Incerti, Jeroen P Jansen"
+#' date: "`r Sys.Date()`"
+#' output:
+#'   html_document:
+#'     toc: true
+#'     toc_float: true
+#' ---
+
+#' This is a replication file for the manuscript. It can be used to produce
+#' an HTML file that displays the results without the text with:
+# rmarkdown::render("replication.R", output_format = "html_document")
+#' Replication was performed by the authors on macOS Mojave 10.14.6. See the
+#' [Session Information](#sessionInfo) below for further details.
+
+#+ message = FALSE, warning = FALSE
+#' # Settings and dependencies
+#' Package versioning is managed using the `renv` package. Packages can be 
+#' installed by cloning the GitHub repository available at 
+#' https://github.com/hesim-dev/hesim-manuscript and running `renv::restore()`. 
+library("data.table")
+library("flexsurv")
+library("ggplot2")
+library("hesim")
+library("knitr")
+library("magrittr")
+library("msm")
+opts_chunk$set(
+  tidy = FALSE, eval = TRUE, cache = TRUE,
+  fig.height = 4, fig.width = 6
+)
+options(prompt = "> ", continue = "+  ", width = 120, useFancyQuotes = TRUE)
+ggplot2::theme_set(ggplot2::theme_bw())
+set.seed(102)
+
+#' # Illustrative example
+#' ## Individual continuous time state transition model
+#' ### Setup model
+#' #### Target population, treatment strategies, and model structure
+# Define transitions
+tmat <- rbind(
+  c(NA, 1, 2),
+  c(NA, NA, 3),
+  c(NA, NA, NA)
+)
+colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
+print(tmat)
+
+# hesim data
+n_patients <- 1000
+patients <- data.table(
+  patient_id = 1:n_patients,
+  age = rnorm(n_patients, mean = 45, sd = 7),
+  female = rbinom(n_patients, size = 1, prob = .51)
+)
+
+states <- data.table(
+  state_id = c(1, 2),
+  state_name = c("Stable", "Progression") # Non-death health states
+)
+
+strategies <- data.table(
+  strategy_id = 1:3,
[…]

### 2021-03-07 manuscript.tex
Add files needed for arXiv submission

--- 2021-02-28 manuscript.Rnw
@@
-\documentclass[article, nojss]{jss}
+\documentclass[article, nojss]{jss}\usepackage[]{graphicx}\usepackage[]{color}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\usepackage{Sweave}
 
@@
 
-<<include=FALSE>>=
-library(knitr)
-knitr::render_sweave()
-opts_chunk$set(
-  engine="R", tidy = FALSE, eval = TRUE, cache = TRUE,
-  fig.height = 4, fig.width = 6
-)
-@
-
-<<preliminaries, echo=FALSE, results="hide">>=
-options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
-ggplot2::theme_set(ggplot2::theme_bw())
-set.seed(102)
-@
-
 %% -- Article metainformation (author, title, ...) -----------------------------
@@
 }
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 \begin{table} [h]
+\caption{Parameterization of disease models.}\label{tbl:parameterize-disease-model}
 \footnotesize
@@
 \end{tabular*}
-\caption{Parameterization of disease models.}\label{tbl:parameterize-disease-model}
 \end{table}
@@
 \begin{table} [h]
+\caption{Parameterization of state value models.}\label{tbl:parameterize-stateval-model}
 \footnotesize
@@
 \end{tabular*}
-\caption{Parameterization of state value models.}\label{tbl:parameterize-stateval-model}
 \end{table}
@@
 \begin{table} [h]
+\caption{Submodels comprising an economic model.}\label{tbl:submodels}
 \footnotesize
@@
 \end{tabular*}
-\caption{Submodels comprising an economic model.}\label{tbl:submodels}
 \end{table}
@@
 \begin{table} [h]
+\caption{Methods for simulating outcomes from economic models.}\label{tbl:simulate-outcomes}
 \footnotesize
@@
 \end{tabular*}
-\caption{Methods for simulating outcomes from economic models.}\label{tbl:simulate-outcomes}
 \end{table}
@@
 
-<<tmat>>=
-tmat <- rbind(
-  c(NA, 1, 2),
-  c(NA, NA, 3),
-  c(NA, NA, NA)
-)
-colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
-print(tmat)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> tmat <- rbind(
++    c(NA, 1, 2),
++    c(NA, NA, 3),
++    c(NA, NA, NA)
++  )
+R> colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
+R> print(tmat)
+\end{Sinput}
+\begin{Soutput}
+            Stable Progression Dead
+Stable          NA           1    2
+Progression     NA          NA    3
+Dead            NA          NA   NA
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<hesim_data>>=
-library("hesim")
-library("data.table")
-n_patients <- 1000
-patients <- data.table(
-  patient_id = 1:n_patients,
-  age = rnorm(n_patients, mean = 45, sd = 7),
-  female = rbinom(n_patients, size = 1, prob = .51)
-)
-
-states <- data.table(
-  state_id = c(1, 2),
-  state_name = c("Stable", "Progression") 
-)
-
-strategies <- data.table(
-  strategy_id = 1:3,
-  strategy_name = c("SOC", "New 1", "New 2"),
-  soc = c(1, 0, 0),
-  new1 = c(0, 1, 0),
-  new2 = c(0, 0, 1)
-)
-
-hesim_dat <- hesim_data(
-  strategies = strategies,
-  patients = patients,
-  states = states
-)
-print(hesim_dat)
-@
+\begin{Schunk}
+\begin{Sinput}
+> library("hesim")
+> library("data.table")
+> n_patients <- 1000
+> patients <- data.table(
++   patient_id = 1:n_patients,
++   age = rnorm(n_patients, mean = 45, sd = 7),
++   female = rbinom(n_patients, size = 1, prob = .51)
++ )
+> 
+> states <- data.table(
++   state_id = c(1, 2),
++   state_name = c("Stable", "Progression") 
++ )
+> 
+> strategies <- data.table(
++   strategy_id = 1:3,
++   strategy_name = c("SOC", "New 1", "New 2"),
++   soc = c(1, 0, 0),
++   new1 = c(0, 1, 0),
++   new2 = c(0, 0, 1)
++ )
+> 
+> hesim_dat <- hesim_data(
++   strategies = strategies,
++   patients = patients,
++   states = states
++ )
+> print(hesim_dat)
+\end{Sinput}
+\begin{Soutput}
+$strategies
+   strategy_id strategy_name soc new1 new2
+1:           1           SOC   1    0    0
+2:           2         New 1   0    1    0
+3:           3         New 2   0    0    1
+
+$patients
+      patient_id      age female
+   1:          1 46.26366      1
+   2:          2 50.49314      1
+   3:          3 35.52785      1
+   4:          4 58.88309      0
+   5:          5 53.66930      0
+  ---                           
+ 996:        996 54.60555      0
+ 997:        997 50.10654      0
+ 998:        998 45.02035      0
+ 999:        999 45.60573      1
+1000:       1000 39.45426      0
+
+$states
+   state_id  state_name
+1:        1      Stable
+2:        2 Progression
+
+attr(,"class")
+[1] "hesim_data"
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<labels_indiv>>=
-labs_indiv <- get_labels(hesim_dat)
-print(labs_indiv)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> labs_indiv <- get_labels(hesim_dat)
+R> print(labs_indiv)
+\end{Sinput}
+\begin{Soutput}
+$strategy_id
+  SOC New 1 New 2 
+    1     2     3
+
+$state_id
+     Stable Progression       Death 
+          1           2           3 
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<ictstm_data>>=
-onc3[patient_id %in% c(1, 2)]
-@
+\begin{Schunk}
+\begin{Sinput}
+R> onc3[patient_id %in% c(1, 2)]
+\end{Sinput}
+\begin{Soutput}
+          from          to strategy_name female      age patient_id
+1:      Stable Progression         New 2      0 59.85813          1
+2:      Stable       Death         New 2      0 59.85813          1
+3: Progression       Death         New 2      0 59.85813          1
+4:      Stable Progression         New 2      0 62.57282          2
+5:      Stable       Death         New 2      0 62.57282          2
+   time_start time_stop status transition_id strategy_id      time
+1:   0.000000  2.420226      1             1           3  2.420226
+2:   0.000000  2.420226      0             2           3  2.420226
+3:   2.420226 14.620258      1             3           3 12.200032
+4:   0.000000  7.497464      0             1           3  7.497464
+5:   0.000000  7.497464      0             2           3  7.497464
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<ictstm_fits, warning = FALSE, message = FALSE>>=
-library("flexsurv")
-n_trans <- max(tmat, na.rm = TRUE) 
-wei_fits <- vector(length = n_trans, mode = "list")
-f <- as.formula(Surv(time, status) ~ factor(strategy_name) + female + age)
-for (i in 1:length(wei_fits)){
-  if (i == 3) f <- update(f, .~.-factor(strategy_name)) 
-  wei_fits[[i]] <- flexsurvreg(f, data = onc3, 
-                               subset = (transition_id == i),
-                               dist = "weibull")
-}
-wei_fits <- flexsurvreg_list(wei_fits)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> library("flexsurv")
+R> n_trans <- max(tmat, na.rm = TRUE) 
+R> wei_fits <- vector(length = n_trans, mode = "list")
+R> f <- as.formula(Surv(time, status) ~ factor(strategy_name) + female + age)
+R> for (i in 1:length(wei_fits)){
++    if (i == 3) f <- update(f, .~.-factor(strategy_name)) 
++    wei_fits[[i]] <- flexsurvreg(f, data = onc3, 
++                                 subset = (transition_id == i),
++                                 dist = "weibull")
++  }
+R> wei_fits <- flexsurvreg_list(wei_fits)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<ictstm_utility>>=
-utility_tbl <- stateval_tbl(
-  data.table(state_id = states$state_id,
-             mean = c(.8, .6),
-             se = c(0.02, .05)
-  ),
-  dist = "beta")
-print(utility_tbl)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> utility_tbl <- stateval_tbl(
++    data.table(state_id = states$state_id,
++               mean = c(.8, .6),
++               se = c(0.02, .05)
++    ),
++    dist = "beta")
+R> print(utility_tbl)
+\end{Sinput}
+\begin{Soutput}
+   state_id mean   se
+1:        1  0.8 0.02
+2:        2  0.6 0.05
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<ictstm_medcosts>>=
-medcost_tbl <- stateval_tbl(
-  data.table(state_id = states$state_id,
-             mean = c(2000, 9500),
-             se = c(2000, 9500)
-  ),
-  dist = "gamma")
-print(medcost_tbl)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> medcost_tbl <- stateval_tbl(
++    data.table(state_id = states$state_id,
++               mean = c(2000, 9500),
++               se = c(2000, 9500)
++    ),
++    dist = "gamma")
+R> print(medcost_tbl)
+\end{Sinput}
+\begin{Soutput}
+   state_id mean   se
+1:        1 2000 2000
+2:        2 9500 9500
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<ictstm_drugcost_setup, echo=FALSE, results="hide">>=
-n_times <- 2
-n_states <- nrow(states)
-n_strategies <- nrow(strategies)
-n_i <- n_states * n_times
-drugcost_dt <-  data.table(
-    strategy_id = rep(strategies$strategy_id, each = n_i),
-    state_id = rep(rep(states$state_id, each = n_times), n_strategies),
-    time_start = rep(c(0, 3/12), n_states * n_strategies),
-    est = c(2000, 2000, 1500, 1200,
-           12000, 12000, 1500, 1200,
-           15000, 15000, 1500, 1200)
-) 
-@
-
-<<ictstm_drugcost>>=
-drugcost_tbl <- stateval_tbl(
-  drugcost_dt,
-  dist = "fixed")
-print(drugcost_tbl)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> drugcost_tbl <- stateval_tbl(
++    drugcost_dt,
++    dist = "fixed")
+R> print(drugcost_tbl)
+\end{Sinput}
+\begin{Soutput}
+    strategy_id state_id time_id time_start time_stop   est
+ 1:           1        1       1       0.00      0.25  2000
+ 2:           1        1       2       0.25       Inf  2000
+ 3:           1        2       1       0.00      0.25  1500
+ 4:           1        2       2       0.25       Inf  1200
+ 5:           2        1       1       0.00      0.25 12000
+ 6:           2        1       2       0.25       Inf 12000
+ 7:           2        2       1       0.00      0.25  1500
+ 8:           2        2       2       0.25       Inf  1200
+ 9:           3        1       1       0.00      0.25 15000
+10:           3        1       2       0.25       Inf 15000
+11:           3        2       1       0.00      0.25  1500
+12:           3        2       2       0.25       Inf  1200
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<ictstm_n_samples>>=
-n_samples <- 1000
-@
+\begin{Schunk}
+\begin{Sinput}
+R> n_samples <- 1000
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<ictstm_transmod_data>>=
-transmod_data <- expand(hesim_dat,
-                        by = c("strategies", "patients"))
-head(transmod_data)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> transmod_data <- expand(hesim_dat,
++                          by = c("strategies", "patients"))
+R> head(transmod_data)
+\end{Sinput}
+\begin{Soutput}
+   strategy_id patient_id strategy_name soc new1 new2      age female
+1:           1          1           SOC   1    0    0 46.26366      1
+2:           1          2           SOC   1    0    0 50.49314      1
+3:           1          3           SOC   1    0    0 35.52785      1
+4:           1          4           SOC   1    0    0 58.88309      0
+5:           1          5           SOC   1    0    0 53.66930      0
+6:           1          6           SOC   1    0    0 53.40432      1
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<ictstm_transmod>>=
-transmod <- create_IndivCtstmTrans(wei_fits, transmod_data,
-                                   trans_mat = tmat, n = n_samples,
-                                   uncertainty = "normal",
-                                   clock = "reset",
-                                   start_age = patients$age)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> transmod <- create_IndivCtstmTrans(wei_fits, transmod_data,
++                                     trans_mat = tmat, n = n_samples,
++                                     uncertainty = "normal",
++                                     clock = "reset",
++                                     start_age = patients$age)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<ictstm_stvalmods>>=
-utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
-                               hesim_data = hesim_dat)
-drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples,
-                                time_reset = TRUE, hesim_data = hesim_dat)
-medcostmod <- create_StateVals(medcost_tbl, n = n_samples,
-                               hesim_data = hesim_dat)
-costmods <- list(Drug = drugcostmod,
-                 Medical = medcostmod)
-@
+\begin{Schunk}
+\begin{Sinput}
+> utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
++                                hesim_data = hesim_dat)
+> drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples,
++                                 time_reset = TRUE, hesim_data = hesim_dat)
+> medcostmod <- create_StateVals(medcost_tbl, n = n_samples,
++                                hesim_data = hesim_dat)
+> costmods <- list(Drug = drugcostmod,
++                  Medical = medcostmod)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<ictstm_econmod>>=
-ictstm <- IndivCtstm$new(trans_model = transmod,
-                         utility_model = utilitymod,
-                         cost_models = costmods)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> ictstm <- IndivCtstm$new(trans_model = transmod,
++                           utility_model = utilitymod,
++                           cost_models = costmods)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<ictstm_sim_disease>>=
-ictstm$sim_disease(max_age = 100)
-head(ictstm$disprog_)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> ictstm$sim_disease(max_age = 100)
+R> head(ictstm$disprog_)
+\end{Sinput}
+\begin{Soutput}
+   sample strategy_id patient_id grp_id from to final time_start
+1:      1           1          1      1    1  3     1   0.000000
+2:      1           1          2      1    1  2     0   0.000000
+3:      1           1          2      1    2  3     1   8.058198
+4:      1           1          3      1    1  2     0   0.000000
+5:      1           1          3      1    2  3     1  15.250551
+6:      1           1          4      1    1  3     1   0.000000
+   time_stop
+1: 11.183651
+2:  8.058198
+3: 12.794488
+4: 15.250551
+5: 23.791357
+6:  6.047286
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<ictstm_sim_stateprobs>>=
-ictstm$sim_stateprobs(t = seq(0, 30 , 1/12))
-head(ictstm$stateprobs_)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> ictstm$sim_stateprobs(t = seq(0, 30 , 1/12))
+R> head(ictstm$stateprobs_)
+\end{Sinput}
+\begin{Soutput}
+   sample strategy_id grp_id state_id          t  prob
+1:      1           1      1        1 0.00000000 1.000
+2:      1           1      1        1 0.08333333 1.000
+3:      1           1      1        1 0.16666667 1.000
+4:      1           1      1        1 0.25000000 0.999
+5:      1           1      1        1 0.33333333 0.998
+6:      1           1      1        1 0.41666667 0.996
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<ictstm_plot_stateprobs, fig.show = "hide">>=
-library("ggplot2")
-autoplot(ictstm$stateprobs_, labels = labs_indiv,
-         ci = FALSE) 
-@
+\begin{Schunk}
+\begin{Sinput}
+R> library("ggplot2")
+R> autoplot(ictstm$stateprobs_, labels = labs_indiv,
++           ci = FALSE) 
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<ictstm_sim_qalys>>=
-ictstm$sim_qalys(dr = c(0,.03))
-head(ictstm$qalys_)
-@
-
-<<ictstm_sim_costs>>=
-ictstm$sim_costs(dr = .03)
-head(ictstm$costs_)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> ictstm$sim_qalys(dr = c(0,.03))
+R> head(ictstm$qalys_)
+\end{Sinput}
+\begin{Soutput}
+   sample strategy_id grp_id state_id dr    qalys      lys
+1:      1           1      1        1  0 4.933188 6.374648
+2:      1           1      1        2  0 3.233055 5.255885
+3:      1           2      1        1  0 5.604984 7.242740
+4:      1           2      1        2  0 2.683823 4.363013
+5:      1           3      1        1  0 6.289275 8.126979
+6:      1           3      1        2  0 2.915707 4.739981
+\end{Soutput}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+R> ictstm$sim_costs(dr = .03)
+R> head(ictstm$costs_)
+\end{Sinput}
+\begin{Soutput}
+   sample strategy_id grp_id state_id   dr category      costs
+1:      1           1      1        1 0.03     Drug  11352.373
+2:      1           1      1        2 0.03     Drug   4691.177
+3:      1           2      1        1 0.03     Drug  76292.008
+4:      1           2      1        2 0.03     Drug   3819.849
+5:      1           3      1        1 0.03     Drug 105209.810
+6:      1           3      1        2 0.03     Drug   4042.658
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<ictstm_summarize>>=
-library("magrittr")
-ce_sim_ictstm <- ictstm$summarize()
-summary(ce_sim_ictstm, labels = labs_indiv) %>%
-  format()
-@
+\begin{Schunk}
+\begin{Sinput}
+R> library("magrittr")
+R> ce_sim_ictstm <- ictstm$summarize()
+R> summary(ce_sim_ictstm, labels = labs_indiv) %>%
++    format()
+\end{Sinput}
+\begin{Soutput}
+   Discount rate        Outcome                      SOC
+1:          0.00          QALYs        7.87 (7.12, 8.60)
+2:          0.03          QALYs        6.58 (6.03, 7.14)
+3:          0.03    Costs: Drug  15,352 (14,431, 16,234)
+4:          0.03 Costs: Medical  46,784 (4,518, 144,256)
+5:          0.03   Costs: total 62,136 (19,888, 159,037)
+                       New 1                      New 2
+1:         8.59 (7.81, 9.40)         9.20 (8.42, 10.06)
+2:         7.10 (6.52, 7.69)          7.52 (6.96, 8.11)
+3:   78,465 (73,854, 83,200)  105,306 (99,228, 111,686)
+4:   46,222 (4,543, 137,303)    46,840 (4,746, 143,093)
+5: 124,687 (82,038, 218,484) 152,146 (109,625, 245,279)
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<psm_hesim_data>>=
-hesim_dat$patients <- data.table(
-  patient_id = 1:4,
-  grp_id = c(1, 1, 2, 2),
-  patient_wt = rep(1/4, 4),
-  age = c(55, 65, 55, 65),
-  female = c(1, 1, 0, 0),
-  grp_name = rep(c("Female", "Male"), 
-                 each = 2)
-)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> hesim_dat$patients <- data.table(
++    patient_id = 1:4,
++    grp_id = c(1, 1, 2, 2),
++    patient_wt = rep(1/4, 4),
++    age = c(55, 65, 55, 65),
++    female = c(1, 1, 0, 0),
++    grp_name = rep(c("Female", "Male"), 
++                   each = 2)
++  )
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<labels_cohort>>=
-labs_cohort <- get_labels(hesim_dat)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> labs_cohort <- get_labels(hesim_dat)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<psm_data>>=
-onc_pfs_os <- as_pfs_os(onc3, patient_vars = c("patient_id", "female", "age", 
-                                               "strategy_name"))
-onc_pfs_os[patient_id %in% c(1, 2)]
-@
+\begin{Schunk}
+\begin{Sinput}
+R> onc_pfs_os <- as_pfs_os(onc3, patient_vars = c("patient_id", "female", "age", 
++                                                 "strategy_name"))
+R> onc_pfs_os[patient_id %in% c(1, 2)]
+\end{Sinput}
+\begin{Soutput}
+   patient_id female      age strategy_name pfs_time pfs_status
+1:          1      0 59.85813         New 2 2.420226          1
+2:          2      0 62.57282         New 2 7.497464          0
+   os_status   os_time
+1:         1 14.620258
+2:         0  7.497464
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<psm_fits_soc>>=
-onc_pfs_os_soc <- onc_pfs_os[strategy_name == "SOC"]
-fit_pfs_soc_wei <- flexsurvreg(
-  Surv(pfs_time, pfs_status) ~ female,
-  data = onc_pfs_os_soc,
-  dist = "weibull")
-
-fit_os_soc_wei <- flexsurvreg(
-  Surv(os_time, os_status) ~  female,
-  data = onc_pfs_os_soc,
-  dist = "weibull")
-
-psmfit_soc_wei <- partsurvfit(
-  flexsurvreg_list(pfs = fit_pfs_soc_wei, os = fit_os_soc_wei),
-  data = onc_pfs_os_soc
-)
-@
+\begin{Schunk}
+\begin{Sinput}
+> onc_pfs_os_soc <- onc_pfs_os[strategy_name == "SOC"]
+> fit_pfs_soc_wei <- flexsurvreg(
++   Surv(pfs_time, pfs_status) ~ female,
++   data = onc_pfs_os_soc,
++   dist = "weibull")
+> 
+> fit_os_soc_wei <- flexsurvreg(
++   Surv(os_time, os_status) ~  female,
++   data = onc_pfs_os_soc,
++   dist = "weibull")
+> 
+> psmfit_soc_wei <- partsurvfit(
++   flexsurvreg_list(pfs = fit_pfs_soc_wei, os = fit_os_soc_wei),
++   data = onc_pfs_os_soc
++ )
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<psm_fits, include = FALSE>>=
-# Create dummies
-onc_pfs_os[, new1 := ifelse(strategy_name == "New 1", 1, 0)]
-onc_pfs_os[, new2 := ifelse(strategy_name == "New 2", 1, 0)]
-
-# Fit models
-fit_pfs_wei <- flexsurv::flexsurvreg(
-  Surv(pfs_time, pfs_status) ~ female + new1 + new2,
-  data = onc_pfs_os,
-  dist = "weibull")
-
-fit_os_wei <- flexsurvreg(
-  Surv(os_time, os_status) ~ female + new1 + new2,
-  data = onc_pfs_os,
-  dist = "weibull")
-
-# Extract relative treatment effects
-params_pfs_wei <- create_params(fit_pfs_wei, n = n_samples)
-params_os_wei <- create_params(fit_os_wei, n = n_samples)
-coef_pfs_wei <- params_pfs_wei$coefs$scale[, c("new1", "new2")]
-coef_os_wei <- params_os_wei$coefs$scale[, c("new1", "new2")]
-@
-
-<<psm_coefs>>=
-head(coef_pfs_wei)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> head(coef_pfs_wei)
+\end{Sinput}
+\begin{Soutput}
+          new1      new2
+[1,] 0.1814189 0.2654431
+[2,] 0.1390351 0.2907570
+[3,] 0.1298460 0.2307868
+[4,] 0.1729702 0.3143110
+[5,] 0.1988663 0.3179721
+[6,] 0.1305311 0.2261630
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<psm_drugcost>>=
-drugcost_tbl <- stateval_tbl(
-  drugcost_dt[time_start ==  0][, time_start := NULL],
-  dist = "fixed")
-print(drugcost_tbl)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> drugcost_tbl <- stateval_tbl(
++    drugcost_dt[time_start ==  0][, time_start := NULL],
++    dist = "fixed")
+R> print(drugcost_tbl)
+\end{Sinput}
+\begin{Soutput}
+   strategy_id state_id   est
+1:           1        1  2000
+2:           1        2  1500
+3:           2        1 12000
+4:           2        2  1500
+5:           3        1 15000
+6:           3        2  1500
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<psm_survmods_data>>=
-survmods_data <- expand(hesim_dat, by = c("strategies", "patients"))
-head(survmods_data)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> survmods_data <- expand(hesim_dat, by = c("strategies", "patients"))
+R> head(survmods_data)
+\end{Sinput}
+\begin{Soutput}
+   strategy_id patient_id strategy_name soc new1 new2 grp_id
+1:           1          1           SOC   1    0    0      1
+2:           1          2           SOC   1    0    0      1
+3:           1          3           SOC   1    0    0      2
+4:           1          4           SOC   1    0    0      2
+5:           2          1         New 1   0    1    0      1
+6:           2          2         New 1   0    1    0      1
+   patient_wt age female grp_name
+1:       0.25  55      1   Female
+2:       0.25  65      1   Female
+3:       0.25  55      0     Male
+4:       0.25  65      0     Male
+5:       0.25  55      1   Female
+6:       0.25  65      1   Female
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<psm_survmods_params, warning = FALSE>>=
-survmods_params <- create_params(psmfit_soc_wei, n = n_samples, 
-                                 uncertainty = "bootstrap", 
-                                 max_errors = 5, silent = TRUE)
-survmods_params$pfs$coefs$scale <- cbind(survmods_params$pfs$coefs$scale,
-                                         coef_pfs_wei)
-survmods_params$os$coefs$scale <- cbind(survmods_params$os$coefs$scale,
-                                        coef_os_wei)
-survmods_params$pfs$coefs$scale[1:2, ]
-@
+\begin{Schunk}
+\begin{Sinput}
+R> survmods_params <- create_params(psmfit_soc_wei, n = n_samples, 
++                                   uncertainty = "bootstrap", 
++                                   max_errors = 5, silent = TRUE)
+R> survmods_params$pfs$coefs$scale <- cbind(survmods_params$pfs$coefs$scale,
++                                           coef_pfs_wei)
+R> survmods_params$os$coefs$scale <- cbind(survmods_params$os$coefs$scale,
++                                          coef_os_wei)
+R> survmods_params$pfs$coefs$scale[1:2, ]
+\end{Sinput}
+\begin{Soutput}
+     (Intercept)     female      new1      new2
+[1,]    1.784666 -0.1364781 0.1814189 0.2654431
+[2,]    1.776743 -0.1173777 0.1390351 0.2907570
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<psm_survmods>>=
-survmods_data[, ("(Intercept)") := 1]
-survmods <- create_PsmCurves(survmods_params, 
-                             input_data = survmods_data)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> survmods_data[, ("(Intercept)") := 1]
+R> survmods <- create_PsmCurves(survmods_params, 
++                               input_data = survmods_data)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<psm_stvalmods>>=
-utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
-                               hesim_data = hesim_dat)
-drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples, 
-                                hesim_data = hesim_dat)
-medcostmod <- create_StateVals(medcost_tbl, n = n_samples, 
-                               hesim_data = hesim_dat)
-costmods <- list(Drug = drugcostmod, Medical = medcostmod)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
++                                 hesim_data = hesim_dat)
+R> drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples, 
++                                  hesim_data = hesim_dat)
+R> medcostmod <- create_StateVals(medcost_tbl, n = n_samples, 
++                                 hesim_data = hesim_dat)
+R> costmods <- list(Drug = drugcostmod, Medical = medcostmod)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<psm_econmod>>=
-psm <- Psm$new(survival_models = survmods,
-               utility_model = utilitymod,
-               cost_models = costmods)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> psm <- Psm$new(survival_models = survmods,
++                 utility_model = utilitymod,
++                 cost_models = costmods)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<psm_sim_survival>>=
-times <- seq(0, 30, by = .1)
-psm$sim_survival(t = times)
-head(psm$survival_)
-@
-
-<<psm_plot_survival, fig.show = "hide">>=
-autoplot(psm$survival_, 
-         labels = c(labs_cohort,
-                    list(curve = c("PFS" = 1, "OS" = 2))
-                    ),
-         ci = TRUE, ci_style = "ribbon")
-@
+\begin{Schunk}
+\begin{Sinput}
+R> times <- seq(0, 30, by = .1)
+R> psm$sim_survival(t = times)
+R> head(psm$survival_)
+\end{Sinput}
+\begin{Soutput}
+   sample strategy_id patient_id grp_id patient_wt curve   t
+1:      1           1          1      1       0.25     1 0.0
+2:      1           1          1      1       0.25     1 0.1
+3:      1           1          1      1       0.25     1 0.2
+4:      1           1          1      1       0.25     1 0.3
+5:      1           1          1      1       0.25     1 0.4
+6:      1           1          1      1       0.25     1 0.5
+    survival
+1: 1.0000000
+2: 0.9998540
+3: 0.9993126
+4: 0.9982992
+5: 0.9967670
+6: 0.9946815
+\end{Soutput}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+R> autoplot(psm$survival_, 
++           labels = c(labs_cohort,
++                      list(curve = c("PFS" = 1, "OS" = 2))
++                      ),
++           ci = TRUE, ci_style = "ribbon")
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<psm_sim_stateprobs>>=
-psm$sim_stateprobs()
-psm$stateprobs_[sample == 1 & patient_id == 1 & state_id == 2 & t == 12]
-@
+\begin{Schunk}
+\begin{Sinput}
+R> psm$sim_stateprobs()
+R> psm$stateprobs_[sample == 1 & patient_id == 1 & state_id == 2 & t == 12]
+\end{Sinput}
+\begin{Soutput}
+   sample strategy_id patient_id grp_id patient_wt state_id  t
+1:      1           1          1      1       0.25        2 12
+2:      1           2          1      1       0.25        2 12
+3:      1           3          1      1       0.25        2 12
+        prob
+1: 0.3215069
+2: 0.3934037
+3: 0.4148810
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<psm_costs_qalys>>=
-psm$sim_qalys(dr = .03) 
-psm$sim_costs(dr = .03)
-@
-
-<<psm_summary>>=
-ce_sim_psm <- psm$summarize(by_grp = TRUE)
-summary(ce_sim_psm, labels = labs_cohort) %>%
-  format()
-@
+\begin{Schunk}
+\begin{Sinput}
+R> psm$sim_qalys(dr = .03) 
+R> psm$sim_costs(dr = .03)
+\end{Sinput}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+R> ce_sim_psm <- psm$summarize(by_grp = TRUE)
+R> summary(ce_sim_psm, labels = labs_cohort) %>%
++    format()
+\end{Sinput}
+\begin{Soutput}
+    Group Discount rate        Outcome                      SOC
+1: Female          0.03          QALYs        6.10 (5.55, 6.63)
+2: Female          0.03    Costs: Drug  15,225 (14,629, 15,900)
+3: Female          0.03 Costs: Medical  52,296 (4,435, 180,774)
+4: Female          0.03   Costs: total 67,521 (19,657, 195,939)
+5:   Male          0.03          QALYs        6.51 (5.99, 7.08)
+6:   Male          0.03    Costs: Drug  16,247 (15,553, 17,004)
+7:   Male          0.03 Costs: Medical  52,677 (4,767, 171,325)
+8:   Male          0.03   Costs: total 68,924 (20,902, 187,307)
+                       New 1                      New 2
+1:         6.61 (6.00, 7.22)          6.96 (6.36, 7.60)
+2:   66,473 (62,592, 70,188)    88,466 (83,561, 93,548)
+3:   52,717 (4,871, 171,715)    53,138 (5,156, 169,727)
+4: 119,189 (70,613, 239,546)  141,604 (92,445, 260,157)
+5:         7.05 (6.46, 7.70)          7.42 (6.81, 8.07)
+6:   74,027 (69,772, 78,081)   98,691 (93,156, 104,387)
+7:   52,709 (5,188, 165,850)    52,874 (5,417, 162,113)
+8: 126,736 (78,810, 238,080) 151,565 (103,556, 262,961)
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<cdtstm_data, warning = FALSE>>=
-head(onc3p)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> head(onc3p)
+\end{Sinput}
+\begin{Soutput}
+         state strategy_name female      age patient_id      time
+1:      Stable         New 2      0 59.85813          1  0.000000
+2: Progression         New 2      0 59.85813          1  2.420226
+3:       Death         New 2      0 59.85813          1 14.620258
+4:      Stable         New 2      0 62.57282          2  0.000000
+5:      Stable         New 2      0 62.57282          2  7.497464
+6:      Stable           SOC      1 61.44379          3  0.000000
+   strategy_id state_id
+1:           3        1
+2:           3        2
+3:           3        3
+4:           3        1
+5:           3        1
+6:           1        1
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<cdtstm_fit_soc, warning = FALSE, message = FALSE>>=
-library("msm")
-qinit <- matrix(0, nrow = 3, ncol = 3)
-qinit[1, 2] <-  0.28; qinit[1, 3] <-  0.013; qinit[2, 3] <-  0.10
-msm_fit <- msm(state_id ~ time, subject = patient_id, 
-               data = onc3p[strategy_name == "SOC"],
-               exacttimes = FALSE,
-               covariates = ~ age + female,
-               qmatrix = qinit, gen.inits = FALSE)
-@
-
-<<msm_exact, include = FALSE, message = FALSE>>=
-# msm with exact transition times
-tmat2 <- tmat; tmat2[is.na(tmat2)] <- 0
-msm_efit <- msm(state_id ~ time, subject = patient_id, 
-                data = onc3p[strategy_name == "SOC"],
-                qmatrix = tmat2, exacttimes = TRUE)
-
-# flexsurv with exponential model
-exp_fits <- vector(length = n_trans, mode = "list")
-for (i in 1:length(wei_fits)){
-  exp_fits[[i]] <- flexsurvreg(Surv(time, status) ~ 1, data = onc3, 
-                               subset = (transition_id == i & 
-                                         strategy_name == "SOC"),
-                               dist = "exponential")
-}
-
-# Compare
-msm_efit
-exp_fits[[1]]$res
-exp_fits[[2]]$res
-exp_fits[[3]]$res
-@
+\begin{Schunk}
+\begin{Sinput}
+R> library("msm")
+R> qinit <- matrix(0, nrow = 3, ncol = 3)
+R> qinit[1, 2] <-  0.28; qinit[1, 3] <-  0.013; qinit[2, 3] <-  0.10
+R> msm_fit <- msm(state_id ~ time, subject = patient_id, 
++                 data = onc3p[strategy_name == "SOC"],
++                 exacttimes = FALSE,
++                 covariates = ~ age + female,
++                 qmatrix = qinit, gen.inits = FALSE)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<cdtstm_params>>=
-params_rr <- list(
-  lrr_12_est = c(soc = log(1), new1 = log(.80), new2 = log(.71)),
-  lrr_12_se = c(soc = 0, new1 = .03, new2 = .04),
-  lrr_13_est = c(soc = log(1), new1 = log(.90), new2 = log(.85)),
-  lrr_13_se = c(soc = 0, new1 = .02, new2 = .03)
-)
-
-params_rr_def <- define_rng({
-  list(
-    rr_12 = lognormal_rng(lrr_12_est, lrr_12_se),
-    rr_13 = lognormal_rng(lrr_13_est, lrr_13_se)
-  )}, n = n_samples)
-params_rr_rng <- eval_rng(params_rr_def, params_rr)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> params_rr <- list(
++    lrr_12_est = c(soc = log(1), new1 = log(.80), new2 = log(.71)),
++    lrr_12_se = c(soc = 0, new1 = .03, new2 = .04),
++    lrr_13_est = c(soc = log(1), new1 = log(.90), new2 = log(.85)),
++    lrr_13_se = c(soc = 0, new1 = .02, new2 = .03)
++  )  
+R> 
+R> params_rr_def <- define_rng({
++    list(
++      rr_12 = lognormal_rng(lrr_12_est, lrr_12_se),
++      rr_13 = lognormal_rng(lrr_13_est, lrr_13_se)
++    )}, n = n_samples)
+R> params_rr_rng <- eval_rng(params_rr_def, params_rr)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<qmat_soc>>=
-transmod_data <- survmods_data
-qmat_soc <- qmatrix(msm_fit, 
-                    newdata = transmod_data[strategy_name == "SOC"],
-                    uncertainty = "normal", n = n_samples)
-dim(qmat_soc)
-qmat_soc[,, 1]
-@
+\begin{Schunk}
+\begin{Sinput}
+R> transmod_data <- survmods_data
+R> qmat_soc <- qmatrix(msm_fit, 
++                      newdata = transmod_data[strategy_name == "SOC"],
++                      uncertainty = "normal", n = n_samples)
+R> dim(qmat_soc)
+\end{Sinput}
+\begin{Soutput}
+[1]    3    3 4000
+\end{Soutput}
+\begin{Sinput}
+R> qmat_soc[,, 1]
+\end{Sinput}
+\begin{Soutput}
+           [,1]       [,2]       [,3]
+[1,] -0.3533989  0.3371267 0.01627226
+[2,]  0.0000000 -0.1190753 0.11907530
+[3,]  0.0000000  0.0000000 0.00000000
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<pmat_soc>>=
-cycle_len <- 1/4
-pmat_soc <- expmat(qmat_soc, t = cycle_len)
-pmat_soc[,, 1]
-@
+\begin{Schunk}
+\begin{Sinput}
+R> cycle_len <- 1/4
+R> pmat_soc <- expmat(qmat_soc, t = cycle_len)
+R> pmat_soc[,, 1]
+\end{Sinput}
+\begin{Soutput}
+          [,1]       [,2]       [,3]
+[1,] 0.9154407 0.07945955 0.00509979
+[2,] 0.0000000 0.97066990 0.02933010
+[3,] 0.0000000 0.00000000 1.00000000
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<transprobs_id>>=
-tpmat_id <- tpmatrix_id(transmod_data, n_samples)
-head(tpmat_id)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> tpmat_id <- tpmatrix_id(transmod_data, n_samples)
+R> head(tpmat_id)
+\end{Sinput}
+\begin{Soutput}
+   sample strategy_id patient_id grp_id patient_wt
+1:      1           1          1      1       0.25
+2:      1           1          2      1       0.25
+3:      1           1          3      2       0.25
+4:      1           1          4      2       0.25
+5:      1           2          1      1       0.25
+6:      1           2          2      1       0.25
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<xbeta>>=
-xbeta <- function(x, beta) c(x %*% t(beta))
-@
+\begin{Schunk}
+\begin{Sinput}
+R> xbeta <- function(x, beta) c(x %*% t(beta))
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<rr>>=
-x_rr <- as.matrix(transmod_data[, .(soc, new1, new2)])
-rr <- cbind(xbeta(x_rr, params_rr_rng$rr_12),
-            xbeta(x_rr, params_rr_rng$rr_13))
-head(rr) 
-@
+\begin{Schunk}
+\begin{Sinput}
+> x_rr <- as.matrix(transmod_data[, .(soc, new1, new2)])
+> rr <- cbind(xbeta(x_rr, params_rr_rng$rr_12),
++             xbeta(x_rr, params_rr_rng$rr_13))
+> head(rr) 
+\end{Sinput}
+\begin{Soutput}
+         [,1]      [,2]
+[1,] 1.000000 1.0000000
+[2,] 1.000000 1.0000000
+[3,] 1.000000 1.0000000
+[4,] 1.000000 1.0000000
+[5,] 0.833283 0.9039595
+[6,] 0.833283 0.9039595
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<apply_rr>>=
-pmat <- apply_rr(pmat_soc, rr = rr,
-                 index = list(c(1, 2), c(1, 3)))
-tprobs <- tparams_transprobs(pmat, tpmat_id)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> pmat <- apply_rr(pmat_soc, rr = rr,
++                   index = list(c(1, 2), c(1, 3)))
+R> tprobs <- tparams_transprobs(pmat, tpmat_id)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<dtstm_transmod>>=
-transmod <- CohortDtstmTrans$new(params = tprobs, 
-                                 cycle_length = cycle_len)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> transmod <- CohortDtstmTrans$new(params = tprobs, 
++                                   cycle_length = cycle_len)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<dtstm_econmod>>=
-cdtstm <- CohortDtstm$new(trans_model = transmod,
-                          utility_model = psm$utility_model,
-                          cost_models = psm$cost_models)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> cdtstm <- CohortDtstm$new(trans_model = transmod,
++                            utility_model = psm$utility_model,
++                            cost_models = psm$cost_models)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<cdtstm_sim_stateprobs>>=
-cdtstm$sim_stateprobs(n_cycles = 30/cycle_len)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> cdtstm$sim_stateprobs(n_cycles = 30/cycle_len)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<cdtstm_sim_qalys_costs>>=
-cdtstm$sim_qalys(dr = .03)
-cdtstm$sim_costs(dr = .03)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> cdtstm$sim_qalys(dr = .03)
+R> cdtstm$sim_costs(dr = .03)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<cdtstm_summary>>=
-ce_sim_cdtstm <- cdtstm$summarize()
-summary(ce_sim_cdtstm, labels = labs_cohort) %>%
-  format()
-@
+\begin{Schunk}
+\begin{Sinput}
+R> ce_sim_cdtstm <- cdtstm$summarize()
+R> summary(ce_sim_cdtstm, labels = labs_cohort) %>%
++    format()
+\end{Sinput}
+\begin{Soutput}
+   Discount rate        Outcome                      SOC
+1:          0.03          QALYs        5.56 (4.78, 6.26)
+2:          0.03    Costs: Drug  13,862 (12,563, 14,711)
+3:          0.03 Costs: Medical  62,515 (4,186, 217,838)
+4:          0.03   Costs: total 76,378 (18,215, 231,236)
+                       New 1                     New 2
+1:         5.92 (5.10, 6.67)         6.14 (5.27, 6.84)
+2:   46,196 (41,201, 50,104)   60,714 (53,418, 66,584)
+3:   62,011 (4,351, 223,642)   61,636 (4,671, 213,904)
+4: 108,207 (50,113, 267,630) 122,350 (64,868, 274,308)
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<ce>>=
-ce_sim_ictstm
-@
+\begin{Schunk}
+\begin{Sinput}
+R> ce_sim_ictstm
+\end{Sinput}
+\begin{Soutput}
+$costs
+      category   dr sample strategy_id     costs grp_id
+   1:     Drug 0.03      1           1  16043.55      1
+   2:     Drug 0.03      1           2  80111.86      1
+   3:     Drug 0.03      1           3 109252.47      1
+   4:     Drug 0.03      2           1  14645.16      1
+   5:     Drug 0.03      2           2  77962.49      1
+  ---                                                  
+8996:    total 0.03    999           2  95024.16      1
+8997:    total 0.03    999           3 122357.69      1
+8998:    total 0.03   1000           1  52638.13      1
+8999:    total 0.03   1000           2 119410.47      1
+9000:    total 0.03   1000           3 151164.45      1
+
+$qalys
+        dr sample strategy_id    qalys grp_id
+   1: 0.00      1           1 8.166243      1
+   2: 0.00      1           2 8.288806      1
+   3: 0.00      1           3 9.204982      1
+   4: 0.00      2           1 7.440970      1
+   5: 0.00      2           2 8.414148      1
+  ---                                        
+5996: 0.03    999           2 7.134710      1
+5997: 0.03    999           3 7.399217      1
+5998: 0.03   1000           1 6.435002      1
+5999: 0.03   1000           2 7.103850      1
+6000: 0.03   1000           3 7.584523      1
+
+attr(,"class")
+[1] "ce"
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<cea>>=
-wtp <- seq(0, 250000, 500) 
-cea_ictstm <- cea(ce_sim_ictstm, dr_costs = .03, dr_qalys = .03, k = wtp)
-cea_pw_ictstm <- cea_pw(ce_sim_ictstm, comparator = 1,
-                        dr_qalys = .03, dr_costs = .03,
-                        k = wtp)
-@
+\begin{Schunk}
+\begin{Sinput}
+> wtp <- seq(0, 250000, 500) 
+> cea_ictstm <- cea(ce_sim_ictstm, dr_costs = .03, dr_qalys = .03, k = wtp)
+> cea_pw_ictstm <- cea_pw(ce_sim_ictstm, comparator = 1,
++                         dr_qalys = .03, dr_costs = .03,
++                         k = wtp)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<cea_icer>>=
-icer(cea_pw_ictstm, k = 100000, labels = labs_indiv) %>%
-  format()
-@
+\begin{Schunk}
+\begin{Sinput}
+R> icer(cea_pw_ictstm, k = 100000, labels = labs_indiv) %>%
++    format()
+\end{Sinput}
+\begin{Soutput}
+             Outcome                     New 1
+1: Incremental QALYs         0.52 (0.16, 0.90)
+2: Incremental costs   62,551 (51,685, 70,389)
+3:   Incremental NMB -11,003 (-42,132, 21,247)
+4:              ICER                   121,345
+                      New 2
+1:        0.93 (0.58, 1.27)
+2: 90,010 (77,342, 100,581)
+3:  3,223 (-28,582, 33,456)
+4:                   96,543
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<ce_plane, fig.height = 3, fig.width = 4.5, fig.show="hide">>=
-plot_ceplane(cea_pw_ictstm, labels = labs_indiv)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> plot_ceplane(cea_pw_ictstm, labels = labs_indiv)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<ceac, fig.height = 3, fig.width = 4.5, fig.show = "hide">>=
-plot_ceac(cea_ictstm, labels = labs_indiv)
-plot_ceac(cea_pw_ictstm, labels = labs_indiv)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> plot_ceac(cea_ictstm, labels = labs_indiv)
+\end{Sinput}
+\begin{Sinput}
+R> plot_ceac(cea_pw_ictstm, labels = labs_indiv)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<ceaf, fig.height = 3, fig.width = 4.5, fig.show = "hide">>=
-plot_ceaf(cea_ictstm, labels = labs_indiv)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> plot_ceaf(cea_ictstm, labels = labs_indiv)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<evpi, fig.height = 3, fig.width = 4.5, fig.show = "hide">>=
-plot_evpi(cea_ictstm, labels = labs_indiv)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> plot_evpi(cea_ictstm, labels = labs_indiv)
+\end{Sinput}
+\end{Schunk}
 
@@
 %% -- Session info -------------------------------------------------------------
-<<sessionInfo, include=FALSE>>=
-sessionInfo()
-@
 

### 2021-07-11 manuscript.tex
Updates to R packages including renv; rebuild manuscript

@@
 \begin{Sinput}
-> library("hesim")
-> library("data.table")
-> n_patients <- 1000
-> patients <- data.table(
+R> library("hesim")
+R> library("data.table")
+R> n_patients <- 1000
+R> patients <- data.table(
 +    patient_id = 1:n_patients,
@@
 +  )
-> 
-> states <- data.table(
+R> 
+R> states <- data.table(
 +    state_id = c(1, 2),
@@
 +  )
-> 
-> strategies <- data.table(
+R> 
+R> strategies <- data.table(
 +    strategy_id = 1:3,
@@
 +  )
-> 
-> hesim_dat <- hesim_data(
+R> 
+R> hesim_dat <- hesim_data(
 +    strategies = strategies,
@@
 +  )
-> print(hesim_dat)
+R> print(hesim_dat)
 \end{Sinput}
@@
 \begin{Sinput}
-> utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
+R> utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
 +                                 hesim_data = hesim_dat)
-> drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples,
+R> drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples,
 +                                  time_reset = TRUE, hesim_data = hesim_dat)
-> medcostmod <- create_StateVals(medcost_tbl, n = n_samples,
+R> medcostmod <- create_StateVals(medcost_tbl, n = n_samples,
 +                                 hesim_data = hesim_dat)
-> costmods <- list(Drug = drugcostmod,
+R> costmods <- list(Drug = drugcostmod,
 +                   Medical = medcostmod)
@@
 \begin{Sinput}
-> onc_pfs_os_soc <- onc_pfs_os[strategy_name == "SOC"]
-> fit_pfs_soc_wei <- flexsurvreg(
+R> onc_pfs_os_soc <- onc_pfs_os[strategy_name == "SOC"]
+R> fit_pfs_soc_wei <- flexsurvreg(
 +    Surv(pfs_time, pfs_status) ~ female,
@@
 +    dist = "weibull")
-> 
-> fit_os_soc_wei <- flexsurvreg(
+R> 
+R> fit_os_soc_wei <- flexsurvreg(
 +    Surv(os_time, os_status) ~  female,
@@
 +    dist = "weibull")
-> 
-> psmfit_soc_wei <- partsurvfit(
+R> 
+R> psmfit_soc_wei <- partsurvfit(
 +    flexsurvreg_list(pfs = fit_pfs_soc_wei, os = fit_os_soc_wei),
@@
 \begin{Sinput}
-> x_rr <- as.matrix(transmod_data[, .(soc, new1, new2)])
-> rr <- cbind(xbeta(x_rr, params_rr_rng$rr_12),
+R> x_rr <- as.matrix(transmod_data[, .(soc, new1, new2)])
+R> rr <- cbind(xbeta(x_rr, params_rr_rng$rr_12),
 +              xbeta(x_rr, params_rr_rng$rr_13))
-> head(rr) 
+R> head(rr) 
 \end{Sinput}
@@
 \begin{Sinput}
-> wtp <- seq(0, 250000, 500) 
-> cea_ictstm <- cea(ce_sim_ictstm, dr_costs = .03, dr_qalys = .03, k = wtp)
-> cea_pw_ictstm <- cea_pw(ce_sim_ictstm, comparator = 1,
+R> wtp <- seq(0, 250000, 500) 
+R> cea_ictstm <- cea(ce_sim_ictstm, dr_costs = .03, dr_qalys = .03, k = wtp)
+R> cea_pw_ictstm <- cea_pw(ce_sim_ictstm, comparator = 1,
 +                          dr_qalys = .03, dr_costs = .03,

### 2021-07-11 manuscript.Rnw
Rename Section 4 to 'Simulation of model outcomes'

@@
 An overview of the coding framework is provided in Section~\ref{sec:framework}.
-An illustrative example using a three state model of disease progression in oncology is provided in Section~\ref{sec:illustrative-example}.
+Section~\ref{sec:model-simulation} uses a three state model of disease progression in oncology to show how \pkg{hesim} can be used to simulate model outcomes.
 Analyses are conducted for each of the three supported model types, illustrating approaches suitable for both patient-level and aggregate-level data.
@@
 Second, a right Riemann sum uses values at the end of each time interval, $f(t_c^{*}) = f(t_{c})$.
-Finally, the trapezoid rule averages values at the start and end of each interval, $f(t_c^{*}) = \frac{1}{2}\Delta t_c[f(t_{c-1}) + f(t_{c})]$.
+Finally, the trapezoid rule averages values at the start and end of each interval, $f(t_c^{*}) = \frac{1}{2}[f(t_{c-1}) + f(t_{c})]$.
 
@@
 %% -- Illustrative example -----------------------------------------------------
-\section{Illustrative example} \label{sec:illustrative-example}
-We consider a three-state model commonly used in oncology.
+\section{Simulation of model outcomes} \label{sec:model-simulation}
+To demonstrate how \pkg{hesim} can be used to simulate outcomes from an economic model, we consider a three state model commonly used in oncology.
 As shown in Figure~\ref{fig:three-state-model}, the three health states are stable disease (i.e., not progressed), progression, and death.
@@
 Uncertainty over $\theta$ is captured using the PSA.
-The simulated distribution of QALYs and costs for each treatment strategy can be produced with the \code{$summarize()} method as was demonstrated in Section~\ref{sec:illustrative-example}.
+The simulated distribution of QALYs and costs for each treatment strategy can be produced with the \code{$summarize()} method as was demonstrated in Section~\ref{sec:model-simulation}.
 We will use the output from the iCTSTM in this example.
@@
 This paper formally describes the economic models and cost-effectiveness framework used by the \pkg{hesim} package.
-An illustrative example is provided for a three state model commonly used to evaluate treatments in oncology.
+An illustrative example was provided for a three state model commonly used to evaluate treatments in oncology.
 However, since \pkg{hesim} aims to be flexible, the example only covers a subset of the relevant modeling approaches and functionality.

### 2021-07-12 manuscript.Rnw
Remove explicit print statements

@@
 colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
-print(tmat)
+tmat
 @
@@
 )
-print(hesim_dat)
+hesim_dat
 @
@@
 labs_indiv <- get_labels(hesim_dat)
-print(labs_indiv)
+labs_indiv
 @
@@
   dist = "beta")
-print(utility_tbl)
+utility_tbl
 @
@@
   dist = "gamma")
-print(medcost_tbl)
+medcost_tbl
 @
@@
   dist = "fixed")
-print(drugcost_tbl)
+drugcost_tbl
 @

### 2021-07-12 manuscript.tex
Remove explicit print statements

@@
 \begin{Sinput}
-R> tmat <- rbind(
+> tmat <- rbind(
 +   c(NA, 1, 2),
@@
 + )
-R> colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
-R> print(tmat)
+> colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
+> tmat
 \end{Sinput}
@@
 \begin{Sinput}
-R> library("hesim")
-R> library("data.table")
-R> n_patients <- 1000
-R> patients <- data.table(
+> library("hesim")
+> library("data.table")
+> n_patients <- 1000
+> patients <- data.table(
 +   patient_id = 1:n_patients,
@@
 + )
-R> 
-R> states <- data.table(
+> 
+> states <- data.table(
 +   state_id = c(1, 2),
@@
 + )
-R> 
-R> strategies <- data.table(
+> 
+> strategies <- data.table(
 +   strategy_id = 1:3,
@@
 + )
-R> 
-R> hesim_dat <- hesim_data(
+> 
+> hesim_dat <- hesim_data(
 +   strategies = strategies,
@@
 + )
-R> print(hesim_dat)
+> hesim_dat
 \end{Sinput}
@@
 \begin{Sinput}
-R> labs_indiv <- get_labels(hesim_dat)
-R> print(labs_indiv)
+> labs_indiv <- get_labels(hesim_dat)
+> labs_indiv
 \end{Sinput}
@@
 \begin{Sinput}
-R> utility_tbl <- stateval_tbl(
+> utility_tbl <- stateval_tbl(
 +   data.table(state_id = states$state_id,
@@
 +   dist = "beta")
-R> print(utility_tbl)
+> utility_tbl
 \end{Sinput}
@@
 \begin{Sinput}
-R> medcost_tbl <- stateval_tbl(
+> medcost_tbl <- stateval_tbl(
 +   data.table(state_id = states$state_id,
@@
 +   dist = "gamma")
-R> print(medcost_tbl)
+> medcost_tbl
 \end{Sinput}
@@
 \begin{Sinput}
-R> drugcost_tbl <- stateval_tbl(
+> drugcost_tbl <- stateval_tbl(
 +   drugcost_dt,
 +   dist = "fixed")
-R> print(drugcost_tbl)
+> drugcost_tbl
 \end{Sinput}
@@
 \begin{Sinput}
-R> drugcost_tbl <- stateval_tbl(
+> drugcost_tbl <- stateval_tbl(
 +   drugcost_dt[time_start ==  0][, time_start := NULL],
 +   dist = "fixed")
-R> print(drugcost_tbl)
+> drugcost_tbl
 \end{Sinput}

### 2021-07-12 manuscript.Rnw
Remove unnecessary abbreviations within abstract

@@
 \Abstract{
-Health economic models simulate the costs and effects of health technologies for use in health technology assessment (HTA) to inform efficient use of scarce resources.
+Health economic models simulate the costs and effects of health technologies for use in health technology assessment to inform efficient use of scarce resources.
 Models have historically been developed using spreadsheet software (e.g., Microsoft \proglang{Excel}) and while use of \proglang{R} is growing, general purpose modeling software is still limited.
 \pkg{hesim} is an \proglang{R} package that helps fill this gap by facilitating parameterization, simulation, and analysis of economic models in an integrated manner.
-Supported model types include cohort discrete time state transition models (cDTSTMs), individual continuous time state transition models (iCTSTMs), and partitioned survival models (PSMs), encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
+Supported model types include cohort discrete time state transition models, individual continuous time state transition models, and partitioned survival models, encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
 A modular design based on \pkg{R6} and \code{S3} classes allows users to combine separate submodels for disease progression, costs, and utility in a flexible way.
@@
 Simulation code is written in \proglang{C++} so complex simulations such as those combining PSA and individual simulation can be run much more quickly than previously possible.
-Decision analysis within a cost-effectiveness framework is performed using simulated costs and quality-adjusted life years (QALYs) from a PSA.
+Decision analysis within a cost-effectiveness framework is performed using simulated costs and quality-adjusted life years from a PSA.
 }

### 2021-07-23 manuscript.Rnw
Add references for data.table and rcpp

@@
 To maximize flexibility and facilitate integration of the statistical methods and economic model, parameters can be set by either fitting models in \proglang{R} or by inputting values obtained from external sources.
-So that individual-level simulation and PSA can be run quickly, \pkg{Rcpp} and \pkg{data.table} are heavily utilized.
+So that individual-level simulation and PSA can be run quickly, \pkg{Rcpp} \citep{eddelbuettel2011rcpp, eddelbuettel2013rcpp, eddelbuettel2018rcpp} and \pkg{data.table} \citep{dowle2021datatable} are heavily utilized.
 After simulating costs and quality-adjusted life-years (QALYs) from a PSA, decision analysis can be performed within a cost-effectiveness framework.

### 2021-07-23 manuscript.tex
Update package versions

@@
 \begin{Sinput}
-> tmat <- rbind(
+R> tmat <- rbind(
 +    c(NA, 1, 2),
@@
 +  )
-> colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
-> tmat
+R> colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
+R> tmat
 \end{Sinput}
@@
 +  )
-> hesim_dat
+R> hesim_dat
 \end{Sinput}
@@
 \begin{Sinput}
-> labs_indiv <- get_labels(hesim_dat)
-> labs_indiv
+R> labs_indiv <- get_labels(hesim_dat)
+R> labs_indiv
 \end{Sinput}
@@
 \begin{Sinput}
-> utility_tbl <- stateval_tbl(
+R> utility_tbl <- stateval_tbl(
 +    data.table(state_id = states$state_id,
@@
 +    dist = "beta")
-> utility_tbl
+R> utility_tbl
 \end{Sinput}
@@
 \begin{Sinput}
-> medcost_tbl <- stateval_tbl(
+R> medcost_tbl <- stateval_tbl(
 +    data.table(state_id = states$state_id,
@@
 +    dist = "gamma")
-> medcost_tbl
+R> medcost_tbl
 \end{Sinput}
@@
 \begin{Sinput}
-> drugcost_tbl <- stateval_tbl(
+R> drugcost_tbl <- stateval_tbl(
 +    drugcost_dt,
 +    dist = "fixed")
-> drugcost_tbl
+R> drugcost_tbl
 \end{Sinput}
@@
 \begin{Sinput}
-> drugcost_tbl <- stateval_tbl(
+R> drugcost_tbl <- stateval_tbl(
 +    drugcost_dt[time_start ==  0][, time_start := NULL],
 +    dist = "fixed")
-> drugcost_tbl
+R> drugcost_tbl
 \end{Sinput}

### 2021-07-24 manuscript.Rnw
Modify R style

@@
 <<preliminaries, echo=FALSE, results="hide">>=
-options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+options(prompt = "R> ", continue = "+  ", width = 76, useFancyQuotes = FALSE)
 ggplot2::theme_set(ggplot2::theme_bw())
@@
 <<tmat>>=
-tmat <- rbind(
-  c(NA, 1, 2),
-  c(NA, NA, 3),
-  c(NA, NA, NA)
-)
+tmat <- rbind(c(NA, 1, 2), c(NA, NA, 3), c(NA, NA, NA))
 colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
@@
 )
-
 states <- data.table(
@@
 )
-
 strategies <- data.table(
@@
 )
-
 hesim_dat <- hesim_data(
@@
   wei_fits[[i]] <- flexsurvreg(f, data = onc3, 
-                               subset = (transition_id == i),
-                               dist = "weibull")
+    subset = (transition_id == i), dist = "weibull")
 }
@@
 utility_tbl <- stateval_tbl(
-  data.table(state_id = states$state_id,
+  data.table(
+    state_id = states$state_id,
     mean = c(.8, .6),
@@
   ),
-  dist = "beta")
+  dist = "beta"
+)
 utility_tbl
@@
 medcost_tbl <- stateval_tbl(
-  data.table(state_id = states$state_id,
+  data.table(
+    state_id = states$state_id,
     mean = c(2000, 9500),
@@
   ),
-  dist = "gamma")
+  dist = "gamma"
+)
 medcost_tbl
@@
   time_start = rep(c(0, 3/12), n_states * n_strategies),
-    est = c(2000, 2000, 1500, 1200,
+  est = c(
+    2000, 2000, 1500, 1200,
     12000, 12000, 1500, 1200,
-           15000, 15000, 1500, 1200)
+    15000, 15000, 1500, 1200
   )
+) 
 @
@@
 <<ictstm_drugcost>>=
-drugcost_tbl <- stateval_tbl(
-  drugcost_dt,
-  dist = "fixed")
+drugcost_tbl <- stateval_tbl(drugcost_dt, dist = "fixed")
 drugcost_tbl
@@
 <<ictstm_transmod_data>>=
-transmod_data <- expand(hesim_dat,
-                        by = c("strategies", "patients"))
+transmod_data <- expand(hesim_dat, by = c("strategies", "patients"))
 head(transmod_data)
@@
 transmod <- create_IndivCtstmTrans(wei_fits, transmod_data,
-                                   trans_mat = tmat, n = n_samples,
-                                   uncertainty = "normal",
-                                   clock = "reset",
-                                   start_age = patients$age)
+  trans_mat = tmat, n = n_samples, uncertainty = "normal",
+  clock = "reset", start_age = patients$age)
 @
@@
   hesim_data = hesim_dat)
-costmods <- list(Drug = drugcostmod,
-                 Medical = medcostmod)
+costmods <- list(Drug = drugcostmod, Medical = medcostmod)
 @
@@
 ictstm <- IndivCtstm$new(trans_model = transmod, 
-                         utility_model = utilitymod,
-                         cost_models = costmods)
+  utility_model = utilitymod, cost_models = costmods)
 @
@@
 library("ggplot2")
-autoplot(ictstm$stateprobs_, labels = labs_indiv,
-         ci = FALSE) 
+autoplot(ictstm$stateprobs_, labels = labs_indiv, ci = FALSE) 
 @
@@
   female = c(1, 1, 0, 0),
-  grp_name = rep(c("Female", "Male"), 
-                 each = 2)
+  grp_name = rep(c("Female", "Male"), each = 2)
 )
@@
 <<psm_data>>=
-onc_pfs_os <- as_pfs_os(onc3, patient_vars = c("patient_id", "female", "age", 
-                                               "strategy_name"))
+onc_pfs_os <- as_pfs_os(onc3, 
+  patient_vars = c("patient_id", "female", "age", "strategy_name"))
 onc_pfs_os[patient_id %in% c(1, 2)]
@@
 onc_pfs_os_soc <- onc_pfs_os[strategy_name == "SOC"]
-fit_pfs_soc_wei <- flexsurvreg(
-  Surv(pfs_time, pfs_status) ~ female,
-  data = onc_pfs_os_soc,
-  dist = "weibull")
-
-fit_os_soc_wei <- flexsurvreg(
-  Surv(os_time, os_status) ~  female,
-  data = onc_pfs_os_soc,
-  dist = "weibull")
-
+fit_pfs_soc_wei <- flexsurvreg(Surv(pfs_time, pfs_status) ~ female,
+  data = onc_pfs_os_soc, dist = "weibull")
+fit_os_soc_wei <- flexsurvreg(Surv(os_time, os_status) ~  female,
+  data = onc_pfs_os_soc, dist = "weibull")
 psmfit_soc_wei <- partsurvfit(
   flexsurvreg_list(pfs = fit_pfs_soc_wei, os = fit_os_soc_wei), 
-  data = onc_pfs_os_soc
-)
+  data = onc_pfs_os_soc)
 @
@@
 # Fit models
-fit_pfs_wei <- flexsurv::flexsurvreg(
-  Surv(pfs_time, pfs_status) ~ female + new1 + new2,
-  data = onc_pfs_os,
-  dist = "weibull")
-
-fit_os_wei <- flexsurvreg(
-  Surv(os_time, os_status) ~ female + new1 + new2,
-  data = onc_pfs_os,
-  dist = "weibull")
+fit_pfs_wei <- flexsurvreg(Surv(pfs_time, pfs_status) ~ female + new1 + new2,
+  data = onc_pfs_os, dist = "weibull")
+
+fit_os_wei <- flexsurvreg(Surv(os_time, os_status) ~ female + new1 + new2,
+  data = onc_pfs_os, dist = "weibull")
 
@@
   drugcost_dt[time_start ==  0][, time_start := NULL],
-  dist = "fixed")
+  dist = "fixed"
+)
 drugcost_tbl
@@
 survmods_params <- create_params(psmfit_soc_wei, n = n_samples, 
-                                 uncertainty = "bootstrap", 
-                                 max_errors = 5, silent = TRUE)
+  uncertainty = "bootstrap", max_errors = 5, silent = TRUE)
 survmods_params$pfs$coefs$scale <- cbind(survmods_params$pfs$coefs$scale,
@@
 survmods_data[, ("(Intercept)") := 1]
-survmods <- create_PsmCurves(survmods_params, 
-                             input_data = survmods_data)
+survmods <- create_PsmCurves(survmods_params, input_data = survmods_data)
 @
@@
 <<psm_econmod>>=
-psm <- Psm$new(survival_models = survmods,
-               utility_model = utilitymod,
+psm <- Psm$new(survival_models = survmods, utility_model = utilitymod,
   cost_models = costmods)
@@
 autoplot(psm$survival_, 
-         labels = c(labs_cohort,
-                    list(curve = c("PFS" = 1, "OS" = 2))
-                    ),
+  labels = c(labs_cohort, list(curve = c("PFS" = 1, "OS" = 2))),
   ci = TRUE, ci_style = "ribbon")
@@
 msm_fit <- msm(state_id ~ time, subject = patient_id, 
-               data = onc3p[strategy_name == "SOC"],
-               exacttimes = FALSE,
-               covariates = ~ age + female,
-               qmatrix = qinit, gen.inits = FALSE)
+  data = onc3p[strategy_name == "SOC"], exacttimes = FALSE,
+  covariates = ~ age + female, qmatrix = qinit, gen.inits = FALSE)
 @
@@
   exp_fits[[i]] <- flexsurvreg(Surv(time, status) ~ 1, data = onc3, 
-                               subset = (transition_id == i & 
-                                         strategy_name == "SOC"),
+    subset = (transition_id == i & strategy_name == "SOC"),
     dist = "exponential")
@@
 )  
-
 params_rr_def <- define_rng({
@@
     rr_13 = lognormal_rng(lrr_13_est, lrr_13_se)
-  )}, n = n_samples)
+  )}, 
+  n = n_samples
+)
 params_rr_rng <- eval_rng(params_rr_def, params_rr)
@@
 x_rr <- as.matrix(transmod_data[, .(soc, new1, new2)])
-rr <- cbind(xbeta(x_rr, params_rr_rng$rr_12),
-            xbeta(x_rr, params_rr_rng$rr_13))
+rr <- cbind(
+  xbeta(x_rr, params_rr_rng$rr_12),
+  xbeta(x_rr, params_rr_rng$rr_13)
+)
 head(rr) 
@@
 <<apply_rr>>=
-pmat <- apply_rr(pmat_soc, rr = rr,
-                 index = list(c(1, 2), c(1, 3)))
+pmat <- apply_rr(pmat_soc, rr = rr, index = list(c(1, 2), c(1, 3)))
 tprobs <- tparams_transprobs(pmat, tpmat_id)
@@
 cdtstm <- CohortDtstm$new(trans_model = transmod, 
-                          utility_model = psm$utility_model,
-                          cost_models = psm$cost_models)
+  utility_model = psm$utility_model, cost_models = psm$cost_models)
 @
@@
 cea_ictstm <- cea(ce_sim_ictstm, dr_costs = .03, dr_qalys = .03, k = wtp)
-cea_pw_ictstm <- cea_pw(ce_sim_ictstm, comparator = 1,
-                        dr_qalys = .03, dr_costs = .03,
-                        k = wtp)
+cea_pw_ictstm <- cea_pw(ce_sim_ictstm, comparator = 1, dr_qalys = .03, 
+  dr_costs = .03, k = wtp)
 @

### 2021-07-24 manuscript.tex
Modify R style

@@
 \begin{Sinput}
-R> tmat <- rbind(
-+    c(NA, 1, 2),
-+    c(NA, NA, 3),
-+    c(NA, NA, NA)
-+  )
+R> tmat <- rbind(c(NA, 1, 2), c(NA, NA, 3), c(NA, NA, NA))
 R> colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
@@
 +  )
-R> 
 R> states <- data.table(
@@
 +  )
-R> 
 R> strategies <- data.table(
@@
 +  )
-R> 
 R> hesim_dat <- hesim_data(
@@
 +    wei_fits[[i]] <- flexsurvreg(f, data = onc3, 
-+                                 subset = (transition_id == i),
-+                                 dist = "weibull")
++      subset = (transition_id == i), dist = "weibull")
 +  }
@@
 R> utility_tbl <- stateval_tbl(
-+    data.table(state_id = states$state_id,
++    data.table(
++      state_id = states$state_id,
 +      mean = c(.8, .6),
@@
 +    ),
-+    dist = "beta")
++    dist = "beta"
++  )
 R> utility_tbl
@@
 R> medcost_tbl <- stateval_tbl(
-+    data.table(state_id = states$state_id,
++    data.table(
++      state_id = states$state_id,
 +      mean = c(2000, 9500),
@@
 +    ),
-+    dist = "gamma")
++    dist = "gamma"
++  )
 R> medcost_tbl
@@
 \begin{Sinput}
-R> drugcost_tbl <- stateval_tbl(
-+    drugcost_dt,
-+    dist = "fixed")
+R> drugcost_tbl <- stateval_tbl(drugcost_dt, dist = "fixed")
 R> drugcost_tbl
@@
 \begin{Sinput}
-R> transmod_data <- expand(hesim_dat,
-+                          by = c("strategies", "patients"))
+R> transmod_data <- expand(hesim_dat, by = c("strategies", "patients"))
 R> head(transmod_data)
@@
 R> transmod <- create_IndivCtstmTrans(wei_fits, transmod_data,
-+                                     trans_mat = tmat, n = n_samples,
-+                                     uncertainty = "normal",
-+                                     clock = "reset",
-+                                     start_age = patients$age)
++    trans_mat = tmat, n = n_samples, uncertainty = "normal",
++    clock = "reset", start_age = patients$age)
 \end{Sinput}
@@
 +    hesim_data = hesim_dat)
-R> costmods <- list(Drug = drugcostmod,
-+                   Medical = medcostmod)
+R> costmods <- list(Drug = drugcostmod, Medical = medcostmod)
 \end{Sinput}
@@
 R> ictstm <- IndivCtstm$new(trans_model = transmod, 
-+                           utility_model = utilitymod,
-+                           cost_models = costmods)
++    utility_model = utilitymod, cost_models = costmods)
 \end{Sinput}
@@
 \begin{Soutput}
-   sample strategy_id patient_id grp_id from to final time_start
-1:      1           1          1      1    1  3     1   0.000000
-2:      1           1          2      1    1  2     0   0.000000
-3:      1           1          2      1    2  3     1   8.058198
-4:      1           1          3      1    1  2     0   0.000000
-5:      1           1          3      1    2  3     1  15.250551
-6:      1           1          4      1    1  3     1   0.000000
-   time_stop
-1: 11.183651
-2:  8.058198
-3: 12.794488
-4: 15.250551
-5: 23.791357
-6:  6.047286
+   sample strategy_id patient_id grp_id from to final time_start time_stop
+1:      1           1          1      1    1  3     1   0.000000 11.183651
+2:      1           1          2      1    1  2     0   0.000000  8.058198
+3:      1           1          2      1    2  3     1   8.058198 12.794488
+4:      1           1          3      1    1  2     0   0.000000 15.250551
+5:      1           1          3      1    2  3     1  15.250551 23.791357
+6:      1           1          4      1    1  3     1   0.000000  6.047286
 \end{Soutput}
@@
 R> library("ggplot2")
-R> autoplot(ictstm$stateprobs_, labels = labs_indiv,
-+           ci = FALSE) 
+R> autoplot(ictstm$stateprobs_, labels = labs_indiv, ci = FALSE) 
 \end{Sinput}
@@
 +    female = c(1, 1, 0, 0),
-+    grp_name = rep(c("Female", "Male"), 
-+                   each = 2)
++    grp_name = rep(c("Female", "Male"), each = 2)
 +  )
@@
 \begin{Sinput}
-R> onc_pfs_os <- as_pfs_os(onc3, patient_vars = c("patient_id", "female", "age", 
-+                                                 "strategy_name"))
+R> onc_pfs_os <- as_pfs_os(onc3, 
++    patient_vars = c("patient_id", "female", "age", "strategy_name"))
 R> onc_pfs_os[patient_id %in% c(1, 2)]
@@
 \begin{Soutput}
-   patient_id female      age strategy_name pfs_time pfs_status
-1:          1      0 59.85813         New 2 2.420226          1
-2:          2      0 62.57282         New 2 7.497464          0
-   os_status   os_time
-1:         1 14.620258
-2:         0  7.497464
+   patient_id female      age strategy_name pfs_time pfs_status os_status
+1:          1      0 59.85813         New 2 2.420226          1         1
+2:          2      0 62.57282         New 2 7.497464          0         0
+     os_time
+1: 14.620258
+2:  7.497464
 \end{Soutput}
@@
 R> onc_pfs_os_soc <- onc_pfs_os[strategy_name == "SOC"]
-R> fit_pfs_soc_wei <- flexsurvreg(
-+    Surv(pfs_time, pfs_status) ~ female,
-+    data = onc_pfs_os_soc,
-+    dist = "weibull")
-R> 
-R> fit_os_soc_wei <- flexsurvreg(
-+    Surv(os_time, os_status) ~  female,
-+    data = onc_pfs_os_soc,
-+    dist = "weibull")
-R> 
+R> fit_pfs_soc_wei <- flexsurvreg(Surv(pfs_time, pfs_status) ~ female,
++    data = onc_pfs_os_soc, dist = "weibull")
+R> fit_os_soc_wei <- flexsurvreg(Surv(os_time, os_status) ~  female,
++    data = onc_pfs_os_soc, dist = "weibull")
 R> psmfit_soc_wei <- partsurvfit(
 +    flexsurvreg_list(pfs = fit_pfs_soc_wei, os = fit_os_soc_wei), 
-+    data = onc_pfs_os_soc
-+  )
++    data = onc_pfs_os_soc)
 \end{Sinput}
@@
 +    drugcost_dt[time_start ==  0][, time_start := NULL],
-+    dist = "fixed")
++    dist = "fixed"
++  )
 R> drugcost_tbl
@@
 \begin{Soutput}
-   strategy_id patient_id strategy_name soc new1 new2 grp_id
-1:           1          1           SOC   1    0    0      1
-2:           1          2           SOC   1    0    0      1
-3:           1          3           SOC   1    0    0      2
-4:           1          4           SOC   1    0    0      2
-5:           2          1         New 1   0    1    0      1
-6:           2          2         New 1   0    1    0      1
-   patient_wt age female grp_name
-1:       0.25  55      1   Female
-2:       0.25  65      1   Female
-3:       0.25  55      0     Male
-4:       0.25  65      0     Male
-5:       0.25  55      1   Female
-6:       0.25  65      1   Female
+   strategy_id patient_id strategy_name soc new1 new2 grp_id patient_wt age
+1:           1          1           SOC   1    0    0      1       0.25  55
+2:           1          2           SOC   1    0    0      1       0.25  65
+3:           1          3           SOC   1    0    0      2       0.25  55
+4:           1          4           SOC   1    0    0      2       0.25  65
+5:           2          1         New 1   0    1    0      1       0.25  55
+6:           2          2         New 1   0    1    0      1       0.25  65
+   female grp_name
+1:      1   Female
+2:      1   Female
+3:      0     Male
+4:      0     Male
+5:      1   Female
+6:      1   Female
 \end{Soutput}
@@
 R> survmods_params <- create_params(psmfit_soc_wei, n = n_samples, 
-+                                   uncertainty = "bootstrap", 
-+                                   max_errors = 5, silent = TRUE)
++    uncertainty = "bootstrap", max_errors = 5, silent = TRUE)
 R> survmods_params$pfs$coefs$scale <- cbind(survmods_params$pfs$coefs$scale,
@@
 R> survmods_data[, ("(Intercept)") := 1]
-R> survmods <- create_PsmCurves(survmods_params, 
-+                               input_data = survmods_data)
+R> survmods <- create_PsmCurves(survmods_params, input_data = survmods_data)
 \end{Sinput}
@@
 \begin{Sinput}
-R> psm <- Psm$new(survival_models = survmods,
-+                 utility_model = utilitymod,
+R> psm <- Psm$new(survival_models = survmods, utility_model = utilitymod,
 +    cost_models = costmods)
@@
 \begin{Soutput}
-   sample strategy_id patient_id grp_id patient_wt curve   t
-1:      1           1          1      1       0.25     1 0.0
-2:      1           1          1      1       0.25     1 0.1
-3:      1           1          1      1       0.25     1 0.2
-4:      1           1          1      1       0.25     1 0.3
-5:      1           1          1      1       0.25     1 0.4
-6:      1           1          1      1       0.25     1 0.5
-    survival
-1: 1.0000000
-2: 0.9998540
-3: 0.9993126
-4: 0.9982992
-5: 0.9967670
-6: 0.9946815
+   sample strategy_id patient_id grp_id patient_wt curve   t  survival
+1:      1           1          1      1       0.25     1 0.0 1.0000000
+2:      1           1          1      1       0.25     1 0.1 0.9998540
+3:      1           1          1      1       0.25     1 0.2 0.9993126
+4:      1           1          1      1       0.25     1 0.3 0.9982992
+5:      1           1          1      1       0.25     1 0.4 0.9967670
+6:      1           1          1      1       0.25     1 0.5 0.9946815
 \end{Soutput}
@@
 R> autoplot(psm$survival_, 
-+           labels = c(labs_cohort,
-+                      list(curve = c("PFS" = 1, "OS" = 2))
-+                      ),
++    labels = c(labs_cohort, list(curve = c("PFS" = 1, "OS" = 2))),
 +    ci = TRUE, ci_style = "ribbon")
@@
 \begin{Soutput}
-   sample strategy_id patient_id grp_id patient_wt state_id  t
-1:      1           1          1      1       0.25        2 12
-2:      1           2          1      1       0.25        2 12
-3:      1           3          1      1       0.25        2 12
-        prob
-1: 0.3215069
-2: 0.3934037
-3: 0.4148810
+   sample strategy_id patient_id grp_id patient_wt state_id  t      prob
+1:      1           1          1      1       0.25        2 12 0.3215069
+2:      1           2          1      1       0.25        2 12 0.3934037
+3:      1           3          1      1       0.25        2 12 0.4148810
 \end{Soutput}
@@
 R> msm_fit <- msm(state_id ~ time, subject = patient_id, 
-+                 data = onc3p[strategy_name == "SOC"],
-+                 exacttimes = FALSE,
-+                 covariates = ~ age + female,
-+                 qmatrix = qinit, gen.inits = FALSE)
++    data = onc3p[strategy_name == "SOC"], exacttimes = FALSE,
++    covariates = ~ age + female, qmatrix = qinit, gen.inits = FALSE)
 \end{Sinput}
@@
 +  )  
-R> 
 R> params_rr_def <- define_rng({
@@
 +      rr_13 = lognormal_rng(lrr_13_est, lrr_13_se)
-+    )}, n = n_samples)
++    )}, 
++    n = n_samples
++  )
 R> params_rr_rng <- eval_rng(params_rr_def, params_rr)
@@
 R> x_rr <- as.matrix(transmod_data[, .(soc, new1, new2)])
-R> rr <- cbind(xbeta(x_rr, params_rr_rng$rr_12),
-+              xbeta(x_rr, params_rr_rng$rr_13))
+R> rr <- cbind(
++    xbeta(x_rr, params_rr_rng$rr_12),
++    xbeta(x_rr, params_rr_rng$rr_13)
++  )
 R> head(rr) 
@@
 \begin{Sinput}
-R> pmat <- apply_rr(pmat_soc, rr = rr,
-+                   index = list(c(1, 2), c(1, 3)))
+R> pmat <- apply_rr(pmat_soc, rr = rr, index = list(c(1, 2), c(1, 3)))
 R> tprobs <- tparams_transprobs(pmat, tpmat_id)
@@
 R> cdtstm <- CohortDtstm$new(trans_model = transmod, 
-+                            utility_model = psm$utility_model,
-+                            cost_models = psm$cost_models)
++    utility_model = psm$utility_model, cost_models = psm$cost_models)
 \end{Sinput}
@@
 R> cea_ictstm <- cea(ce_sim_ictstm, dr_costs = .03, dr_qalys = .03, k = wtp)
-R> cea_pw_ictstm <- cea_pw(ce_sim_ictstm, comparator = 1,
-+                          dr_qalys = .03, dr_costs = .03,
-+                          k = wtp)
+R> cea_pw_ictstm <- cea_pw(ce_sim_ictstm, comparator = 1, dr_qalys = .03, 
++    dr_costs = .03, k = wtp)
 \end{Sinput}
@@
 \begin{Soutput}
-             Outcome                     New 1
-1: Incremental QALYs         0.52 (0.16, 0.90)
-2: Incremental costs   62,551 (51,685, 70,389)
-3:   Incremental NMB -11,003 (-42,132, 21,247)
-4:              ICER                   121,345
-                      New 2
-1:        0.93 (0.58, 1.27)
-2: 90,010 (77,342, 100,581)
-3:  3,223 (-28,582, 33,456)
-4:                   96,543
+             Outcome                     New 1                    New 2
+1: Incremental QALYs         0.52 (0.16, 0.90)        0.93 (0.58, 1.27)
+2: Incremental costs   62,551 (51,685, 70,389) 90,010 (77,342, 100,581)
+3:   Incremental NMB -11,003 (-42,132, 21,247)  3,223 (-28,582, 33,456)
+4:              ICER                   121,345                   96,543
 \end{Soutput}

### 2021-08-01 manuscript.Rnw
Some edits to better compare hesim with related R packages

@@
 It is therefore no surprise that a number of related \proglang{R} packages have recently been developed, such as \pkg{BCEA} \citep{baio2017bayesian}, \pkg{SAVI} \citep{strong2014estimating}, \pkg{dampack} \citep{alarid2018dampack}, \pkg{EVSI} \citep{heath2018efficient}, \pkg{survHE} \citep{baio2020survhe}, and \pkg{heemod} \citep{filipovic2017markov}.
-Still, of the available packages, only \pkg{heemod} provides a general purpose framework for developing simulation models and it is limited to Markov cohort models.
+
+The functionality of existing packages can generally be separated into three categories: estimation of model parameters, simulation of model outcomes, and decision analysis.
+Nearly all packages mentioned above support decision analysis in some capacity: \pkg{BCEA}, \pkg{SAVI}, \pkg{dampack}, \pkg{EVSI}, and \pkg{heemod} all contain features for post-processing the output of simulation models in order to quantify decision uncertainty and conduct value of information analyses.
+On the other hand, there is, to our knowledge, only one package, \pkg{heemod}, that is designed for simulation of health economic models.
+\pkg{heemod} supports development of Markov cohort models and can incorporate parameter uncertainty for a PSA.
+Finally, it is important to note that one of, it not the primary advantages of using \proglang{R} for cost-effectiveness modeling is the rich set of packages available for statistical analysis.
+While any number of these can be used for model parameter estimation---as will be partially demonstrated in this paper---\pkg{survHE} is unique in that it was designed specifically for use in health economic evaluations and consequently contains features such as convenient comparisons of multiple parametric survival models and simulation of model parameters for PSA.
 
@@
 After simulating costs and quality-adjusted life-years (QALYs) from a PSA, decision analysis can be performed within a cost-effectiveness framework.
+
+Of the existing packages, \pkg{hesim} is most similar to \pkg{heemod} in that it is designed primarily for simulation of economic models, but can also be used to conduct a CEA.
+That said, there are some notable differences.
+First, \pkg{hesim} can be used for individual-level simulations, whereas \pkg{heemod} only supports cohort-level models.
+Second, all outcomes in \pkg{hesim} are simulated across treatment strategies and patient subgroups within a PSA in a vectorized fashion.
+This not only increases speed because outcomes are simulated at the \proglang{C++} level, but also makes it easier to account for patient heterogeneity and compare a large number of treatment strategies.
+Third, since all model types in \pkg{hesim} can be parameterized by fitting statistical models in \proglang{R}, it is easier to integrate the statistical analysis and economic model structure.
 
@@
 
-Analyses were performed using \pkg{hesim} version \code{0.5.0}. All code including a replication \proglang{R} script for this paper is available at \url{https://github.com/hesim-dev/hesim-manuscript}.
+Analyses were performed using \pkg{hesim} version \code{0.5.1}. All code including a replication \proglang{R} script for this paper is available at \url{https://github.com/hesim-dev/hesim-manuscript}.
 
@@
 As mentioned in the introduction, \pkg{heemod} is the only other general purpose \proglang{R} package for cost-effectiveness modeling.
-It differs from \pkg{hesim} in that it focuses solely on cDTSTMs (i.e., Markov cohort models) and is not designed to facilitate integration of economic models with the statistical models used for parameter estimation.
-Another important difference is that \pkg{hesim} is considerably faster since all simulations are vectorized (i.e, performed at the \proglang{C++} level).
-Performance comparisons are provided on the package website.
+Unlike \pkg{hesim}, simulations are not vectorized.
+To asses the significance of this, performance comparisons are provided on the package website.
 The time-inhomogeneous total hip replacement example from \citet{briggs2006decision} was run with both packages using 1,000 PSA iterations.

### 2021-08-01 manuscript.Rnw
More edits to better compare R packages

@@
 The functionality of existing packages can generally be separated into three categories: estimation of model parameters, simulation of model outcomes, and decision analysis.
-Nearly all packages mentioned above support decision analysis in some capacity: \pkg{BCEA}, \pkg{SAVI}, \pkg{dampack}, \pkg{EVSI}, and \pkg{heemod} all contain features for post-processing the output of simulation models in order to quantify decision uncertainty and conduct value of information analyses.
+Nearly all packages mentioned above can be used to perform decision analysis in some capacity: \pkg{BCEA}, \pkg{SAVI}, \pkg{dampack}, \pkg{EVSI}, and \pkg{heemod} all contain features for post-processing the output of simulation models in order to quantify decision uncertainty and conduct value of information analyses.
 On the other hand, there is, to our knowledge, only one package, \pkg{heemod}, that is designed for simulation of health economic models.
-\pkg{heemod} supports development of Markov cohort models and can incorporate parameter uncertainty for a PSA.
+\pkg{heemod} can be used to develop Markov cohort models and can incorporate parameter uncertainty for a PSA.
 Finally, it is important to note that one of, it not the primary advantages of using \proglang{R} for cost-effectiveness modeling is the rich set of packages available for statistical analysis.
@@
 That said, there are some notable differences.
-First, \pkg{hesim} can be used for individual-level simulations, whereas \pkg{heemod} only supports cohort-level models.
-Second, all outcomes in \pkg{hesim} are simulated across treatment strategies and patient subgroups within a PSA in a vectorized fashion.
+First, \pkg{hesim} can be used for individual-level simulations, whereas \pkg{heemod} can only simulate cohort-level models.
+Second, all outcomes in \pkg{hesim} are simulated across parameter samples for a PSA, treatment strategies, and patient subgroups in a vectorized fashion.
 This not only increases speed because outcomes are simulated at the \proglang{C++} level, but also makes it easier to account for patient heterogeneity and compare a large number of treatment strategies.
@@
 
-As mentioned in the introduction, \pkg{heemod} is the only other general purpose \proglang{R} package for cost-effectiveness modeling.
-Unlike \pkg{hesim}, simulations are not vectorized.
-To asses the significance of this, performance comparisons are provided on the package website.
-The time-inhomogeneous total hip replacement example from \citet{briggs2006decision} was run with both packages using 1,000 PSA iterations.
-At the time of writing, a cohort model in \pkg{heemod} completed in approximately 1.9 minutes while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \pkg{hesim} ran in 7.4 seconds.
-
 There are a number of multi-state modeling \proglang{R} packages that can both fit models and make predictions.
@@
 \pkg{mstate} is another package for multi-state modeling that is designed for estimation of non-parametric and semi-parametric (i.e., Cox proportional hazards) models \citep{de2011mstate}.
-It supports prediction of state probabilities and simulation of multi-state trajectories (to facilitate prediction for clock-reset models) from cumulative hazard functions evaluated at specific times.
+It can predict state probabilities and simulate multi-state trajectories (to facilitate prediction for clock-reset models) from cumulative hazard functions evaluated at specific times.
 \citet{williams2017cost} has adapted \pkg{mstate} for cost-effectiveness modeling applications.
@@
 Second, it is written completely in \proglang{R} and incorporating PSA, multiple treatment strategies, or heterogeneous patients results in code that is not vectorized.
-To compare computational performance, we ran 1,000 PSA iterations where 1 treatment strategy and 1,000 homogeneous patients were simulated through a 6-state Weibull model during each iteration.
-The model completed in 2.8 seconds in \pkg{hesim} and 3.1 hours in \pkg{mstate}, showing that computational considerations are not just an academic concern.
+To compare computational performance, we ran (see ``Benchmarks'' on the package website) 1,000 PSA iterations where 1 treatment strategy and 1,000 homogeneous patients were simulated through a 6-state Weibull model during each iteration.
+At the time of writing, the model completed in 2.8 seconds in \pkg{hesim} and 3.1 hours in \pkg{mstate}, showing that computational considerations are not just an academic concern.
+
+\pkg{heemod} is the only other general purpose \proglang{R} package for cost-effectiveness modeling.
+As noted in the introduction, there are some important differences between \pkg{hesim} and \pkg{heemod} related to the supported model types and integration of statistical and economic modeling.
+In addition, as with \pkg{mstate}, simulations are not vectorized which could result in significantly reduced performance.
+To test this, we simulated a time-inhomogeneous total hip replacement example from \citet{briggs2006decision}.
+A cohort model in \pkg{heemod} completed in approximately 1.9 minutes while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \pkg{hesim} ran in 7.4 seconds.
+
+There are of course features in other \proglang{R} packages that are not in the current version of \pkg{hesim}.
+For example, \pkg{heemod} has an explicit function, \code{calibrate_model()}, for calibrating model parameters that are difficult to measure.
+Furthermore, tunnel states can be automatically created (rather than needing to be created manually) with \pkg{heemod} using the \code{state_time} variable so that costs or transition probabilities can very as a function of time in a health state.
+\pkg{hesim} is also limited in its support for value of information analyses.
+For instance, neither the expected value of partial perfect information (EVPPI) or the expected value of sample information (EVSI) can be computed with \pkg{hesim}, whereas the EVPPI can be computed with \pkg{BCEA}, \pkg{dampack}, and \pkg{SAVI} and EVSI can be computed with \pkg{EVSI} and \pkg{dampack}.
+In this respect, these packages are complementary and simulation output from \pkg{hesim} could be used as inputs to their more advanced value of information functions.
 
@@
 Fifth, equity considerations could be incorporated by adding support for distributional CEA \citep{asaria2015distributional, asaria2016distributional, cookson2017using}.
+Sixth, functions could be added to convert \pkg{hesim} simulation output into formats required by other packages that can analyze the outputs of health economics models, such as those with more advanced value of information analysis capabilities.
 Finally, a larger aspiration is to provide support for parameterization via NMA, particularly for estimation of relative treatment effects in survival (e.g., for PSMs) and multi-state models.

### 2021-08-01 manuscript.Rnw
Added section in discussion about choice to use R6

@@
 Supported model types include cohort discrete time state transition models, individual continuous time state transition models, and partitioned survival models, encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes.
-A modular design based on \pkg{R6} and \code{S3} classes allows users to combine separate submodels for disease progression, costs, and utility in a flexible way.
+A modular design allows users to combine separate submodels for disease progression, costs, and utility in a flexible way.
 Probabilistic sensitivity analysis (PSA) is used to propagate uncertainty in model parameters to model outputs.
@@
 
-A third feature worth mentioning is that a PSM can be extended to $N$ health states.
+A third helpful feature that we did not mention is that a PSM can be extended to $N$ health states.
 The 3-state oncology case might, for instance, be extended to account for adverse events or multiple lines of therapy.
@@
 An example four-state model is provided on the package website.
+
+A design decision that is worth commenting on is the use of \pkg{R6}.
+While it may not be ``traditional'' \proglang{R}, we believe it fits our modular design nicely.
+It is consistent with our conceptualization of the economic modeling process (Figure~\ref{fig:econ-eval-process-hesim}) in that we can have separate submodels for disease progression, costs, and utility, which can be combined into an economic model.
+A particularly important benefit of \pkg{R6} is that methods can belong to objects.
+In our framework, this means that submodels can not only contain inputs for the economic model, but can also generate outputs.
+For example, the disease model of a PSM not only contains the parameters of the survival models and the input data required for prediction, but can also generate relevant outcomes such as survival curves or hazard functions.
+Likewise, the disease model of an iCTSTM not only stores parameters and input data, but can be used to simulate trajectories through a multi-state model and compute state probabilities.
+This makes it easier to check that the submodels produce desired outputs, compare submodels with different underlying modeling assumptions (e.g., different parametric survival distributions), and to use the package in a more modular way.
+The latter may be especially relevant to users who only want to use some features of the package; for, instance a user interested only in epidemiological outcomes could use the disease progression submodel without needing to use the utility or cost submodels.
+All that said, we utilize \code{S3} whenever it makes sense as it is more familiar to most \proglang{R} users.
 
@@
 As noted in the introduction, there are some important differences between \pkg{hesim} and \pkg{heemod} related to the supported model types and integration of statistical and economic modeling.
-In addition, as with \pkg{mstate}, simulations are not vectorized which could result in significantly reduced performance.
+In addition, as with \pkg{mstate}, simulations are not vectorized, which could result in significantly reduced performance.
 To test this, we simulated a time-inhomogeneous total hip replacement example from \citet{briggs2006decision}.

### 2021-08-02 manuscript.Rnw
Minor edits to prose; add references for R packages

@@
 \pkg{heemod} can be used to develop Markov cohort models and can incorporate parameter uncertainty for a PSA.
-Finally, it is important to note that one of, it not the primary advantages of using \proglang{R} for cost-effectiveness modeling is the rich set of packages available for statistical analysis.
-While any number of these can be used for model parameter estimation---as will be partially demonstrated in this paper---\pkg{survHE} is unique in that it was designed specifically for use in health economic evaluations and consequently contains features such as convenient comparisons of multiple parametric survival models and simulation of model parameters for PSA.
+Finally, it is important to note that one of, it not the primary advantage of using \proglang{R} for cost-effectiveness modeling is the rich set of packages available for statistical analysis.
+While any number of these can be used for parameter estimation, \pkg{survHE} is unique in that it was designed specifically for use in health economic evaluations and consequently contains specialized features such as convenient comparisons of multiple parametric survival models and simulation of parameters for PSA.
 
@@
 
-Of the existing packages, \pkg{hesim} is most similar to \pkg{heemod} in that it is designed primarily for simulation of economic models, but can also be used to conduct a CEA.
+Of the existing packages, \pkg{heemod} is most similar to \pkg{hesim} in that it is designed primarily for simulation of economic models, but can also be used to conduct a CEA.
 That said, there are some notable differences.
 First, \pkg{hesim} can be used for individual-level simulations, whereas \pkg{heemod} can only simulate cohort-level models.
-Second, all outcomes in \pkg{hesim} are simulated across parameter samples for a PSA, treatment strategies, and patient subgroups in a vectorized fashion.
+Second, all outcomes in \pkg{hesim} are simulated across parameter samples for a PSA, treatment strategies, and patients in a vectorized fashion.
 This not only increases speed because outcomes are simulated at the \proglang{C++} level, but also makes it easier to account for patient heterogeneity and compare a large number of treatment strategies.
@@
 
-The three types of economic models described in Section~\ref{sec:model-taxonomy} are implemented as the \pkg{R6} classes \code{CohortDtstm}, \code{IndivCtstm}, and \code{Psm}.
+The three types of economic models described in Section~\ref{sec:model-taxonomy} are implemented as the \pkg{R6} \citep{chang2020} classes \code{CohortDtstm}, \code{IndivCtstm}, and \code{Psm}.
 Each class contains methods for simulating disease progression, QALYs, and costs.
@@
 The latter is useful for storing the parameters of a multi-state model or the independent survival models required for a PSM.
-The parameters of a multinomial logistic regression are stored in a \code{params_mlogit} object and can be created by fitting a model for each row in a transition probability matrix with \code{nnet::multinom()}.
+The parameters of a multinomial logistic regression are stored in a \code{params_mlogit} object and can be created by fitting a model for each row in a transition probability matrix with \code{nnet::multinom()} \citep{venables2002}.
 
 \code{tparams_transprobs} objects store transition probability matrices that have been predicted for each PSA sample, treatment strategy, patient from the target population, and optionally time interval (to allow for time inhomogeneous Markov models).
-Transition probabilities can be predicted from a fitted multi-state model using the \pkg{msm} package or constructed ``by hand'' in a custom manner.
+Transition probabilities can be predicted from a fitted multi-state model using the \pkg{msm} package \citep{jackson2011} or constructed ``by hand'' in a custom manner.
 
@@
 The output is a generic cost-effectiveness object (of class \code{ce}) that can be used to perform CEA as demonstrated in Section~\ref{sec:cea}.
-Here, we use the \code{summary()} method to summarize the results and make use of the ``pipes'' from the \pkg{magrittr} to nicely format the output.
+Here, we use the \code{summary()} method to summarize the results and make use of the ``pipes'' from the \pkg{magrittr} \citep{bache2020} package to nicely format the output.
 
@@
 
-A third helpful feature that we did not mention is that a PSM can be extended to $N$ health states.
+A third feature that we did not mention is that a PSM can be extended to $N$ health states.
 The 3-state oncology case might, for instance, be extended to account for adverse events or multiple lines of therapy.

### 2021-08-02 manuscript.Rnw
Minor edits to prose and references

@@
 \section{Discussion} \label{sec:discussion}
-This paper formally describes the economic models and cost-effectiveness framework used by the \pkg{hesim} package.
+This paper formally describes the economic modeling and cost-effectiveness framework used by the \pkg{hesim} package.
 An illustrative example was provided for a three state model commonly used to evaluate treatments in oncology.
@@
 While it may not be ``traditional'' \proglang{R}, we believe it fits our modular design nicely.
-It is consistent with our conceptualization of the economic modeling process (Figure~\ref{fig:econ-eval-process-hesim}) in that we can have separate submodels for disease progression, costs, and utility, which can be combined into an economic model.
+It is consistent with our conceptualization of the economic modeling process (Figure~\ref{fig:econ-eval-process-hesim}) in that it is natural to have classes representing separate submodels for disease progression, costs, and utility, which can be combined into a single class representing an economic model.
 A particularly important benefit of \pkg{R6} is that methods can belong to objects.
@@
 \pkg{hesim} is also limited in its support for value of information analyses.
-For instance, neither the expected value of partial perfect information (EVPPI) or the expected value of sample information (EVSI) can be computed with \pkg{hesim}, whereas the EVPPI can be computed with \pkg{BCEA}, \pkg{dampack}, and \pkg{SAVI} and EVSI can be computed with \pkg{EVSI} and \pkg{dampack}.
+In particular, neither the expected value of partial perfect information (EVPPI) or the expected value of sample information (EVSI) can be computed with \pkg{hesim}, whereas the EVPPI can be computed with \pkg{BCEA}, \pkg{dampack}, and \pkg{SAVI} and EVSI can be computed with \pkg{EVSI} and \pkg{dampack}.
 In this respect, these packages are complementary and simulation output from \pkg{hesim} could be used as inputs to their more advanced value of information functions.
@@
 Fifth, equity considerations could be incorporated by adding support for distributional CEA \citep{asaria2015distributional, asaria2016distributional, cookson2017using}.
-Sixth, functions could be added to convert \pkg{hesim} simulation output into formats required by other packages that can analyze the outputs of health economics models, such as those with more advanced value of information analysis capabilities.
+Sixth, it would be convenient if \pkg{hesim} simulation output could be converted into formats required by other packages that can analyze the outputs of health economics models, such as those with more advanced value of information analysis capabilities.
 Finally, a larger aspiration is to provide support for parameterization via NMA, particularly for estimation of relative treatment effects in survival (e.g., for PSMs) and multi-state models.

### 2021-08-08 replication.R
Rerun replication script with updates packages + some minor edits:

@@ -31,10 +31,10 @@ opts_chunk$set(
   fig.height = 4, fig.width = 6
 )
 options(prompt = "> ", continue = "+  ", width = 120, useFancyQuotes = TRUE)
-ggplot2::theme_set(ggplot2::theme_bw())
+theme_set(theme_bw())
 set.seed(102)
 
-#' # Illustrative example
+#' # Simulation of model outcomes
 #' ## Individual continuous time state transition model
 #' ### Setup model
 #' #### Target population, treatment strategies, and model structure
@@ -45,7 +45,7 @@ tmat <- rbind(
   c(NA, NA, NA)
 )
 colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
-print(tmat)
+tmat
 
 # hesim data
 n_patients <- 1000
@@ -73,11 +73,11 @@ hesim_dat <- hesim_data(
   patients = patients,
   states = states
 )
-print(hesim_dat)
+hesim_dat
 
 #' #### Labels for ID variables
 labs_indiv <- get_labels(hesim_dat)
-print(labs_indiv)
+labs_indiv
 
 #' ### Parameterization
 #' #### Transition model
@@ -104,7 +104,7 @@ utility_tbl <- stateval_tbl(
              se = c(0.02, .05)
   ),
   dist = "beta")
-print(utility_tbl)
+utility_tbl
 
 #' #### Cost models
 # Medical costs
@@ -114,7 +114,7 @@ medcost_tbl <- stateval_tbl(
              se = c(2000, 9500)
   ),
   dist = "gamma")
-print(medcost_tbl)
+medcost_tbl
 
 # Drug costs
 n_times <- 2
@@ -133,7 +133,7 @@ drugcost_dt <-  data.table(
 drugcost_tbl <- stateval_tbl(
   drugcost_dt,
   dist = "fixed")
-print(drugcost_tbl)
+drugcost_tbl
 
 
 #' ### Simulation
@@ -268,7 +268,7 @@ head(coef_pfs_wei)
 drugcost_tbl <- stateval_tbl(
   drugcost_dt[time_start ==  0][, time_start := NULL],
   dist = "fixed")
-print(drugcost_tbl)
+drugcost_tbl
 
 #' ### Simulation
 #' #### Constructing the economic model

### 2021-08-08 manuscript.Rnw
Minor text edits to manuscript

@@
 There are a number of multi-state modeling \proglang{R} packages that can both fit models and make predictions.
-\pkg{hesim} leverages some of these for parameters estimation: \pkg{flexsurv} is used to fit parametric and flexible parametric survival and multi-state models when exact event times are available while \pkg{msm} is used to fit multi-state models when only panel data is available.
+\pkg{hesim} leverages some of these for parameter estimation: \pkg{flexsurv} is used to fit parametric and flexible parametric survival and multi-state models when exact event times are available while \pkg{msm} is used to fit multi-state models when only panel data is available.
 A key difference is that \pkg{hesim} is designed specifically for health economic modeling whereas the other packages are more general purpose statistical packages.
@@
 Second, it is written completely in \proglang{R} and incorporating PSA, multiple treatment strategies, or heterogeneous patients results in code that is not vectorized.
-To compare computational performance, we ran (see ``Benchmarks'' on the package website) 1,000 PSA iterations where 1 treatment strategy and 1,000 homogeneous patients were simulated through a 6-state Weibull model during each iteration.
+To compare computational performance (see ``Benchmarks'' on the package website), we ran 1,000 PSA iterations where 1 treatment strategy and 1,000 homogeneous patients were simulated through a 6-state Weibull model during each iteration.
 At the time of writing, the model completed in 2.8 seconds in \pkg{hesim} and 3.1 hours in \pkg{mstate}, showing that computational considerations are not just an academic concern.
@@
 In addition, as with \pkg{mstate}, simulations are not vectorized, which could result in significantly reduced performance.
-To test this, we simulated a time-inhomogeneous total hip replacement example from \citet{briggs2006decision}.
+To test this, we simulated the time-inhomogeneous total hip replacement example from \citet{briggs2006decision}.
 A cohort model in \pkg{heemod} completed in approximately 1.9 minutes while the same cohort model took approximately 1 second to run in \pkg{hesim} and an equivalent individual-level simulation in \pkg{hesim} ran in 7.4 seconds.
@@
 \pkg{hesim} is also limited in its support for value of information analyses.
-In particular, neither the expected value of partial perfect information (EVPPI) or the expected value of sample information (EVSI) can be computed with \pkg{hesim}, whereas the EVPPI can be computed with \pkg{BCEA}, \pkg{dampack}, and \pkg{SAVI} and EVSI can be computed with \pkg{EVSI} and \pkg{dampack}.
+In particular, neither the expected value of partial perfect information (EVPPI) or the expected value of sample information (EVSI) can be computed with \pkg{hesim}, whereas the EVPPI can be computed with \pkg{BCEA}, \pkg{dampack}, and \pkg{SAVI}, and EVSI can be computed with \pkg{EVSI} and \pkg{dampack}.
 In this respect, these packages are complementary and simulation output from \pkg{hesim} could be used as inputs to their more advanced value of information functions.
@@
 Third, more complex models of utility \citep{kharroubi2007modelling, hernandez2013relationship} and costs \citep{nixon2005methods} could help better capture heterogeneity in preferences and spending.
-Fourth, it would be useful to add functionality that automatically creates tunnel states for users in cDTSTMs.
+Fourth, it would be useful to add options so that tunnel states could be automatically created in cDTSTMs.
 Fifth, equity considerations could be incorporated by adding support for distributional CEA \citep{asaria2015distributional, asaria2016distributional, cookson2017using}.

### 2021-08-14 manuscript.Rnw
Minor manuscript edits

@@
 \section{Framework} \label{sec:framework}
-Economic models consist of a disease model, a utility model, and a set of cost models for each cost category.
+Economic models in \pkg{hesim} consist of a disease model, a utility model, and a set of cost models for each cost category.
 Model development proceeds in 4 steps.
@@
 Finally, the simulated QALYs and costs are used to perform decision analysis within a cost-effectiveness framework.
-While other approaches such as multi-criteria decision analysis (MCDA) could, in principle, be used as well, only CEA is currently supported.
+While other approaches such as multi-criteria decision analysis could, in principle, be used as well, only CEA is currently supported.
 All models are simulated using PSA so that decision uncertainty can be represented.
@@
 \node[draw, minimum height=4cm, minimum width=3.5cm, font=\footnotesize, text depth = 3.5cm] (C) at (5,0) {\textbf{3. Decision analysis}};
-\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (5, .3){CEA};
-\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (5, -.7){MCDA};
+\node[draw,fill=white, minimum height=.8cm, minimum width=2.5cm, font=\footnotesize] at (5, 0){CEA};
 
@@
 Section~\ref{sec:psm-example} uses a PSM in a case where PFS and OS are available but multi-state data is not, and Section~\ref{sec:cdtstm-example} develops a cDTSTM in a setting where multi-state data is available but it is not continuously observed.
-For the sake of illustration, in both Sections~\ref{sec:psm-example} and~\ref{sec:cdtstm-example}, patient-level data is available for the SOC, but not for estimation of relative treatment effects.
+For the sake of illustration, in both Sections~\ref{sec:psm-example} and~\ref{sec:cdtstm-example}, patient-level data is available for the standard of care (SOC), but not for estimation of relative treatment effects.
 
@@
 The target population, treatment strategies, and (non-death) health states are stored in a \code{hesim_data} object and identified with integer valued identification variables (\code{patient_id}, \code{strategy_id}, and \code{state_id}).
-There are 3 treatment strategies: standard of care (SOC) and two new interventions (New 1 and New 2).
+There are 3 treatment strategies: SOC and two new interventions (New 1 and New 2).
 The target population consists of a heterogeneous population of 1,000 patients, which is large enough to ensure that averages across patients in the individual simulation are reasonably stable.

### 2021-08-14 manuscript.Rnw
Typo fixes

@@
 
-Identifiers for the complete array of transitions matrices (inclusive of SOC and the new interventions) are generated with \code{tpmatrix_id()}.
+Identifiers for the complete array of transition matrices (inclusive of SOC and the new interventions) are generated with \code{tpmatrix_id()}.
 
@@
 This makes it easier to check that the submodels produce desired outputs, compare submodels with different underlying modeling assumptions (e.g., different parametric survival distributions), and to use the package in a more modular way.
-The latter may be especially relevant to users who only want to use some features of the package; for, instance a user interested only in epidemiological outcomes could use the disease progression submodel without needing to use the utility or cost submodels.
+The latter may be especially relevant to users who only want to use some features of the package; for instance, a user interested only in epidemiological outcomes could use the disease progression submodel without needing to use the utility or cost submodels.
 All that said, we utilize \code{S3} whenever it makes sense as it is more familiar to most \proglang{R} users.
@@
 \citet{williams2017cost} has adapted \pkg{mstate} for cost-effectiveness modeling applications.
-There are however, some disadvantages of using \pkg{mstate} for health economic modeling since it was not designed for it.
+There are, however, some disadvantages of using \pkg{mstate} for health economic modeling since it was not designed for it.
 \code{mssample()}---the core function for simulating multi-state models---does not natively support heterogeneous target populations, PSA, simultaneous simulation of multiple treatment strategies, or computation of QALYs and costs from the simulated output.
@@
 For example, \pkg{heemod} has an explicit function, \code{calibrate_model()}, for calibrating model parameters that are difficult to measure.
-Furthermore, tunnel states can be automatically created (rather than needing to be created manually) with \pkg{heemod} using the \code{state_time} variable so that costs or transition probabilities can very as a function of time in a health state.
+Furthermore, tunnel states can be automatically created (rather than needing to be created manually) with \pkg{heemod} using the \code{state_time} variable so that costs or transition probabilities can vary as a function of time in a health state.
 \pkg{hesim} is also limited in its support for value of information analyses.
